---
title: "KSPZV1 baseline high-dose PfSPZ multimodal ML- visualize xgboost ML results, JCI revision (Figure 6B)"
author: "Tuan M. Tran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document :
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r links discussing approaches, include=FALSE, eval=FALSE, echo=FALSE}
#Links discussing best approaches:
# https://stats.stackexchange.com/questions/264533/how-should-feature-selection-and-hyperparameter-optimization-be-ordered-in-the-m
```

# Objective

Visualization of xgboost ML with 4-fold cross-validation of full dataset.

# Approach

Plot ROC curves, confusion matrices, and SHAP plots from top n models using xgboost results derived from "KPSZV1 Baseline Highdose PfSPZ Multimodal ML xgb nested fs cv on full set JCI Revision.Rmd".

```{r libraries, message=FALSE}
library(xgboost)
library(dplyr)
library(tidyverse)
library(pROC)
library(caret)
library(doMC)
library(data.table)
library(xgboost)
library(mlr)
library(ggpubr)
library(tictoc)
library(Biobase)
library(googledrive)
library(parallel)
library(parallelMap)
parallelStartSocket(cpus = detectCores())
library(SHAPforxgboost)
```

```{r set local paths, eval=FALSE, include=FALSE, include=FALSE}
datadir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/kspzv1-systems-analysis/"
datadir2 <- "/Users/tuantran/Library/CloudStorage/GoogleDrive-tuantran@iu.edu/Shared drives/[Sec] IN-MED-INF-TRANLAB/Tran Lab Shared/Projects/Doris Duke PfSPZ Kenya/Tuan PfSPZ/KenyaPfSPZ/Final Data Files/"
plotdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Figures JCI resubmission/"
plot_dat_dir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Data for all Plots JCI Resubmission/"
```

```{r read in full dataset, eval=FALSE, include=FALSE, include=FALSE}
complete_baseline_eset <- read_rds(paste0(datadir, "KSPZV1 logCPM expression sets for visualization/PfSPZ_cpm_ExpressionSet_244x21815_AllGroups_bothBatches_0_rmBatchFX_06082021_TMT_logTRUE.rds"))
#high_dose_features <- read_rds(paste0(datadir,"Leetah ML code/kspzv1_hidose_imputed_ml_datasetfull_042123.RDS")) #original data from Leetah
high_dose_features <- read_rds(paste0(datadir2,"highdose_PfSPZ_baseline_correlation_ML_data_with_missing.rds")) #dataset derived from my own code
high_dose_baseline_eset <- complete_baseline_eset[,which(complete_baseline_eset$PATID %in% rownames(high_dose_features))]
colnames(high_dose_baseline_eset) <- gsub("_0", "", colnames(high_dose_baseline_eset))
```

```{r impute missing, eval=FALSE, include=FALSE, include=FALSE}
### Scale and impute missing data in high-dose features
library(missMDA)
library(FactoMineR)
#scale data first since we are using PCA to impute the data
high_dose_features_scaled <- scale(high_dose_features)
high_dose_features_scaled_imputed <- MIPCA(high_dose_features_scaled)
high_dose_features <- high_dose_features_scaled_imputed$res.imputePCA
high_dose_features <- t(high_dose_features)
```

```{r make highdose features, eval=FALSE, echo=FALSE, include=FALSE}
##before splitting, we want to join the classification variable and additional variables of interest to the features data
high_dose_features_class <- t(high_dose_features) %>%
  data.frame(check.names=FALSE) %>%
  rownames_to_column(var = "PATID") %>%
  left_join(., pData(high_dose_baseline_eset) %>%
              mutate(class = factor(ifelse(mal.atp.3 == 1, "infected", "protected"))) %>% #convert to class factor with reference as "infected"
              dplyr::select(PATID, site, SEX, age.vax1, mal.vax.1, class) %>%
              dplyr::rename(sex = "SEX",
                            age = "age.vax1",
                            pfbaseline = mal.vax.1),
            by = "PATID") %>%
  dplyr::select(class, sex, site, pfbaseline, age, everything()) %>%
  column_to_rownames(var = "PATID")

high_dose_features_class <- high_dose_features_class[colnames(high_dose_baseline_eset),]

if(all(rownames(high_dose_features_class) == colnames(high_dose_baseline_eset))){
  print("good to go!")
} else {
    print("please check to see if colnames match.")
}
```

```{r save high_dose_features_class as rds, eval=FALSE, echo=FALSE, include=FALSE}
# write_rds(high_dose_features_class, file = paste0(datadir2, "high_dose_baseline_class_features_scaled_imputed.rds"))
# write_rds(t(high_dose_features), file = paste0(datadir2, "high_dose_baseline_features_only_scaled_imputed.rds"))
```

```{r read high dose baseline features from local, eval=FALSE, include=FALSE, echo=FALSE}
#local
high_dose_features_class <- read_rds(file = paste0(datadir2, "high_dose_baseline_class_features_scaled_imputed.rds"))
train_features <- high_dose_features_class %>%
  dplyr::select(-c(sex, site))
```

### Load feature data

```{r read high dose baseline features from google drive, message=FALSE, warning=FALSE}
#from google drive
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("1QXCOZfeSizhJd53NDbnMU28xO14q1d2G"), path = temp, overwrite = TRUE)
high_dose_features_class <- readRDS(file = dl$local_path)
train_features <- high_dose_features_class %>%
  dplyr::select(-c(sex, site))
```

### Load ML results

This is a large file.

```{r plot ROC curve from cv results for best iteration local, eval=FALSE, include=FALSE, echo=FALSE, message=FALSE}
#readin data saved from above
datadir <- "/Users/tuantran/Library/CloudStorage/GoogleDrive-tuantran@iu.edu/Shared drives/[Sec] IN-MED-INF-TRANLAB/Tran Lab Shared/Projects/Doris Duke PfSPZ Kenya/Tuan PfSPZ/KenyaPfSPZ/Final Data Files/ML results Tuan/"
 
all_folds_2500_runs <- load(paste0(datadir,
                                  "full_high_dose_baseline_4fold_cv_2500_runs/",
                                  "all_folds_data_seed_3813_runs_2500_15_Oct_2023_full.RData"))
per_run_res_2500_runs <- readxl::read_xlsx(paste0(datadir,
                                  "full_high_dose_baseline_4fold_cv_2500_runs/",
                                  "cv_results_all_runs_all_folds_seed_3813_runs_2500_htune_iters_500_15_Oct_2023_full.xlsx"))
summarized_res_2500_runs <- readxl::read_xlsx(paste0(datadir,
                                  "full_high_dose_baseline_4fold_cv_2500_runs/",
                                  "cv_summarized_res_all_folds_seed_3813_runs_2500_htune_iters_500_15_Oct_2023_full.xlsx"))
```

```{r plot ROC curve from cv results for best iteration google drive, message=FALSE, warning=FALSE}
#from google drive
temp <- tempfile(fileext = ".Rdata")
dl <- drive_download(
  as_id("1R8FTKq6_DIKWWj-Z_gvsPtEXUfK_WlKe"), path = temp, overwrite = TRUE)
all_folds_2500_runs <- load(file = dl$local_path)

temp <- tempfile(fileext = ".xlsx")
dl <- drive_download(
  as_id("1RhGlVDH7Gbw8H-YUkO7ScX0NR96nTKlB"), path = temp, overwrite = TRUE)
per_run_res_2500_runs <- readxl::read_xlsx(dl$local_path)

temp <- tempfile(fileext = ".xlsx")
dl <- drive_download(
  as_id("1SM2KWomZ_uZaP8gpgdBwuI-3P810vP9v"), path = temp, overwrite = TRUE)
summarized_res_2500_runs <- readxl::read_xlsx(dl$local_path)
```

### Determine across-fold performance metrics (accuracy, AUC, Brier score, kappa, logless) select best performing high-dose PfSPZ baseline models

```{r determine across fold AUC and kappa, message=FALSE, warning=FALSE}
smooth <- FALSE
runs_to_test <- summarized_res_2500_runs[summarized_res_2500_runs$mean_cv_accuracy>.66,]$run #select only the high-dose PfSPZ models with accuracy >66%.
#runs_to_test <- c("run 211", "run 2222", "run 671")

best_run <- best_combo <- all_folds_response <- all_folds_predictor <- roc_dat <- auc <- confusion_mat <- response_dichotomous <- c()
across_fold_performance <- logloss_list <- brier_list <- c()
for(j in runs_to_test){
  best_run[[j]] <- per_run_res_2500_runs[per_run_res_2500_runs$run==j,] %>%
    group_by(run) %>%
    slice_max(AUC) %>%
    distinct(Features) %>%
    pull(var = run)
  best_combo[[j]] <- per_run_res_2500_runs[per_run_res_2500_runs$run==j,] %>%
    group_by(run) %>%
    slice_max(AUC) %>%
    distinct(Features) %>%
    pull(var = Features)
  #calculate predictive accuracy across all 4 folds
  all_folds_response[[j]] <- c(fold1[[best_run[[j]]]]$predictor_data$class,
                               fold2[[best_run[[j]]]]$predictor_data$class,
                               fold3[[best_run[[j]]]]$predictor_data$class,
                               fold4[[best_run[[j]]]]$predictor_data$class)
  all_folds_predictor[[j]] <- c(fold1[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold2[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold3[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold4[[best_run[[j]]]]$predictor_data$pred.prob)
  roc_dat[[j]] <- pROC::roc(response = all_folds_response[[j]],
                            predictor = all_folds_predictor[[j]],
                            smooth = smooth) 
  auc[[j]] <- roc_dat[[j]]$auc %>% as.numeric(.) %>% signif(.,digits=3)
  response_dichotomous[[j]]  <- ifelse(all_folds_predictor[[j]]  > 0.5,1,0)
  #remember that since we are using the high-dose model, we are going with the original high-dose designation where infected=0, protected=1
  response_dichotomous[[j]]   <- factor(ifelse(response_dichotomous[[j]]==1, "protected", "infected")) 
  confusion_mat[[j]] <- caret::confusionMatrix(response_dichotomous[[j]], all_folds_response[[j]], positive = "protected")
  logloss_list[[j]] <- MLmetrics::LogLoss(all_folds_predictor[[j]], as.numeric(all_folds_response[[j]])-1)
  brier_list[[j]] <- mlr::measureBrier(all_folds_predictor[[j]], all_folds_response[[j]], "infected", "protected")
  across_fold_performance[[j]] <- confusion_mat[[j]]$overall
} 

across_fold_performance_df <- bind_rows(across_fold_performance, .id = "run") %>%
  arrange(desc(Kappa))  %>%
  mutate(Kappa_rank = 1:nrow(.)) %>%
  arrange(desc(Accuracy)) %>%
  mutate(Accuracy_rank = 1:nrow(.))
auc_df_high_dose <- bind_rows(auc, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "AUC") %>%
  arrange(desc(AUC))  %>%
  mutate(AUC_rank = 1:nrow(.))
logloss_df <- bind_rows(logloss_list, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "LogLoss") %>%
  arrange(LogLoss)  %>%
  mutate(LogLoss_rank = 1:nrow(.))
brier_df <- bind_rows(brier_list, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "BrierScore") %>%
  mutate(BrierScore = signif(BrierScore, 3)) %>%
  arrange(BrierScore) %>%
  mutate(BrierScore_rank = 1:nrow(.))

across_fold_perform_high_dose <- auc_df_high_dose %>%
  left_join(., across_fold_performance_df, by = "run") %>%
  left_join(., summarized_res_2500_runs,
            by = "run") %>%
  left_join(., logloss_df,
            by = "run") %>%
  left_join(., brier_df,
            by = "run") %>%
  mutate(AverageRank = (Accuracy_rank + AUC_rank + BrierScore_rank + Kappa_rank + LogLoss_rank)/5) %>%
  dplyr::select(run, Features, AverageRank, BrierScore_rank, Kappa_rank, LogLoss_rank, AUC_rank, Accuracy_rank, AUC, BrierScore, LogLoss, Kappa, Accuracy:McnemarPValue, mean_cv_AUC:sd_cv_accuracy) %>%
  group_by(run) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  arrange(AverageRank)

#clean up feature names
  # across_fold_perform_high_dose <- across_fold_perform_high_dose %>%
  #   mutate(Features =  gsub("CSP_specific_IgG_", "CSP-specific IgG_", Features)) %>% 
  #   mutate(Features =  gsub("FACS_|_innate_BTM|_highBTM|_monaco|CytokineObsConc_", "", Features)) %>%   
  #   mutate(Features =  gsub("PfSPZ", "Pf", Features)) %>% 
  #   mutate(Features =  gsub("_of_live_leukocytes", "_of_live_PBMCs", Features)) %>% 
  #   mutate(Features =  gsub("_", " ", Features))
```


```{r save kappa values, include=FALSE, echo=FALSE, eval=FALSE}
writexl::write_xlsx(across_fold_perform_high_dose, paste0("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Supplementary Tables JCI resubmission/","Table S6 across folds xgb performance highdose baseline.xlsx"))
```

### Arrange data and plot AUROC curves 

```{r arrange data for roc plots, message=FALSE, warning=FALSE}
top_n <- 4
across_fold_perform_highdose_rank_by_AverageRank <- across_fold_perform_high_dose %>%
  arrange(AverageRank) %>%
  drop_na()
select_data <- across_fold_perform_highdose_rank_by_AverageRank[c(1:top_n),]
data.labels <- data.frame("name" = select_data[1:4,]$run,
                          "features" = select_data[1:4,]$Features,
                          "brier" = select_data[1:4,]$BrierScore,
                          "auc" = select_data[1:4,]$AUC,
                          "kappa" = select_data[1:4,]$Kappa,
                          "logloss" = select_data[1:4,]$LogLoss)
roc_plot_dat <- roc_dat[select_data[1:4,]$run] 
roc_plots <- ggroc(roc_plot_dat, legacy.axes = TRUE, size = 1.5) +
  theme_bw() +
  scale_color_brewer("", palette = "Set1", type = "qual") +
  scale_fill_brewer("", palette = "Set1", type = "qual") +
  theme(legend.position = "none") +
  geom_text(data = data.labels[1,], aes(0.875, 0.25, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[2,], aes(0.875, 0.2, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[3,], aes(0.875, 0.15, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[4,], aes(0.875, 0.1, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE) +
  #geom_text(data = data.labels[5,], aes(0.875, 0.05, label = paste("AUC =", auc)), hjust=1, show_guide=FALSE) +
  theme(plot.margin = unit(c(1,1,1,1), "cm")) +
  geom_abline(slope = 1, linetype = "dotted", color = "gray")
```

```{r plot roc, fig.align='center', fig.height=6, fig.width=6}
roc_plots
```

```{r save rocplot data, include=FALSE, message=FALSE, eval=FALSE, echo=FALSE}
writexl::write_xlsx(roc_plots$data, paste0(plot_dat_dir, "Fig 6B baseline high-dose PfSPZ ROC plot.xlsx"))
```

### Plot confusion matrices for high-dose baseline models

```{r confusion matrix plots web ref, echo=FALSE, include=FALSE, eval=FALSE}
#https://stackoverflow.com/questions/46063234/how-to-produce-a-confusion-matrix-and-find-the-misclassification-rate-of-the-naï
#https://stats.stackexchange.com/questions/82162/cohens-kappa-in-plain-english#:~:text=%3C0%20as%20indicating%20no%20agreement,1%20as%20almost%20perfect%20agreement.&text=%3E0.75%20as%20excellent%2C%200.40%20to,and%20%3C0.40%20as%20poor.%22
```

```{r make confusion matrix plots highdose baseline, message=FALSE, warning=FALSE}
library(ggplot2)
library(scales)

ggplotConfusionMatrix <- function(m){
  mytitle <- paste(
    "accuracy", percent_format()(m$overall[1]),
                   "kappa", percent_format()(m$overall[2]))
  mydata <- as.data.frame(m$table) %>%
    mutate(Prediction = ifelse(Prediction == "infected", "NP", "P")) %>%
    mutate(Reference = ifelse(Reference == "infected", "NP", "P"))
  p <- ggplot(data = mydata,
           aes(x = Reference, y = Prediction)) +
    geom_tile(aes(fill = log(Freq)), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue") +
    geom_text(aes(x = Reference, y = Prediction, label = Freq, size = 24)) +
    ggtitle(mytitle) +
    ylab("predicted") +
    xlab("actual outcome") +
    theme(legend.position = "none",
          panel.background = element_blank(),
          plot.background = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(color = "black", size = 10),
          axis.text = element_text(color = "black", size = 10),
          title = element_text(size = 9),) +
    theme(plot.margin = unit(c(0.5,0.45,0.01,0.4), "cm"))
  return(p)
}
```

```{r plot highdose confusion matrices, fig.align='center', fig.height=2, fig.width=8}
#my_models <- c("run 1727", "run 671", "run 808", "run 439") #ranked by Brier
my_models <- select_data[1:4,]$run #ranked by average rank

my_plots <- list(ggplotConfusionMatrix(confusion_mat[[my_models[1]]]),
              ggplotConfusionMatrix(confusion_mat[[my_models[2]]]),
              ggplotConfusionMatrix(confusion_mat[[my_models[3]]]),
              ggplotConfusionMatrix(confusion_mat[[my_models[4]]]))
names(my_plots) <- my_models

my_highdose_confus_mat_plot_top <- ggarrange(plotlist = my_plots[1:2],
                                         labels = names(my_plots[1:2]),
                                         font.label = list(size=14, face="plain"),
                                         ncol = 2,
                                         hjust = -0.1)
my_highdose_confus_mat_plot_bottom <- ggarrange(plotlist = my_plots[3:4],
                                         labels = names(my_plots[3:4]),
                                         font.label = list(size=14, face="plain"),
                                         ncol = 2,
                                         hjust = -0.1)
my_highdose_confus_mat_plot_all <- ggarrange(my_highdose_confus_mat_plot_top,
                                             my_highdose_confus_mat_plot_bottom,
                                             nrow=2)
```

```{r plot my_highdose_confus_mat_plot_all, fig.align='center', fig.height=6, fig.width=5}
my_highdose_confus_mat_plot_all
```

```{r save confusion matrix data, include=FALSE, message=FALSE, eval=FALSE, echo=FALSE}
foo_ids <- c(rep(names(my_plots[1]), 4),
             rep(names(my_plots[2]), 4),
             rep(names(my_plots[3]), 4),
             rep(names(my_plots[4]), 4))
foo <- bind_rows(my_plots[[1]]$data,
          my_plots[[2]]$data,
          my_plots[[3]]$data,
          my_plots[[4]]$data)
foo$model <- foo_ids
writexl::write_xlsx(foo, paste0(plot_dat_dir,"Fig 6B baseline high-dose PfSPZ confusion matrices.xlsx"))
```

### Feature importance by number of appearances in top %1 models

```{r feature importance by number of appearances in top n models, fig.align='center', fig.width=5, fig.height=3.5}
addSmallLegend <- function(myPlot, pointSize = 0.75, textSize = 6, spaceLegend = 0.75) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

#filter select data by performance metric
select_data <- across_fold_perform_highdose_rank_by_AverageRank %>%
  arrange(AverageRank) %>%
  slice_head(n=25)

#build shap data from results of each fold
shap_prep_fold1 <- shap_prep_fold2 <- shap_prep_fold3 <- shap_prep_fold4 <- shap_prep_rbind <- shap_plots <- c()
for(k in rev(select_data$run)){
  shap_prep_fold1[[k]] <- shap.prep(xgb_model = fold1[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold1[[k]]$predictor_data[,fold1[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold2[[k]] <- shap.prep(xgb_model = fold2[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold2[[k]]$predictor_data[,fold2[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold3[[k]] <- shap.prep(xgb_model = fold3[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold3[[k]]$predictor_data[,fold3[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold4[[k]] <- shap.prep(xgb_model = fold4[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold4[[k]]$predictor_data[,fold4[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_rbind[[k]] <- rbind(shap_prep_fold1[[k]],
                           shap_prep_fold2[[k]],
                           shap_prep_fold3[[k]],
                           shap_prep_fold4[[k]]) %>%
    group_by(variable) %>%
    mutate(ID = 1:n()) %>%
    ungroup()
  levels(shap_prep_rbind[[k]]$variable) <- gsub("FACS_|_innate_BTM|_highBTM|_monaco|CytokineObsConc_", "", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("PfSPZ", "Pf", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("_of_live_leukocytes", "_of_live_PBMCs", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("_", " ", levels(shap_prep_rbind[[k]]$variable))  
  shap_plots[[k]] <- shap.plot.summary(shap_prep_rbind[[k]], scientific = FALSE)
}
shap_prep_rbind_df <- bind_rows(shap_prep_rbind, .id = "model")

number_models <- sum(summary(factor(unique(shap_prep_rbind_df$model))))

var_appearances <- shap_prep_rbind_df %>%
  dplyr::select(model,variable) %>%
  distinct() %>%
  group_by(variable) %>%
  summarize(proportion=n()/number_models, n=n()) %>%
  ungroup() %>%
  arrange(desc(n))

foo <- shap_long_iris %>%
  group_by(variable) %>%
  summarize(mean = mean(value))


SHAP_feature_correlations <- shap_prep_rbind_df %>%
  dplyr::select(model,variable, value, rfvalue) %>%
  distinct() %>%
  group_by(model, variable) %>% 
  summarise(r = cor(value, rfvalue)) %>%
  ungroup() %>%
  group_by(variable) %>%
  summarize(mean_r=mean(r))

var_appearances <- var_appearances %>%
  left_join(., SHAP_feature_correlations,
            by = "variable")

feature_importance_plot <- var_appearances %>%
  filter(n>1) %>%
  ggplot(., aes(x= fct_reorder(variable, n, .desc = FALSE))) +
  geom_bar(aes(y=proportion, fill=mean_r), stat="identity") +
  scale_fill_gradient2(low = "darkblue", mid = "white",
                       high = "darkred",
                       midpoint = 0,limits = c(-1,1)) +
  theme_bw() +
  coord_flip() +
  xlab("features") +
  ylab("appearances") + 
  labs(fill = "R") +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) +
  theme(legend.position = "top",
        axis.text = element_text(colour = "black"))

feature_importance_plot <- addSmallLegend(feature_importance_plot)
feature_importance_plot
```

```{r save figure 5X, include=FALSE, message=FALSE, echo=FALSE, eval=FALSE}
pdf(paste0(plotdir, "Figure 6B highdose baseline feature by pct appearance barplot AverageRank.pdf"),
    width = 5.9, height = 3.4)
addSmallLegend(feature_importance_plot)
dev.off()
```

```{r save feature importance plot data, include=FALSE, message=FALSE, eval=FALSE, echo=FALSE}
writexl::write_xlsx(feature_importance_plot$data, paste0(plot_dat_dir,"Fig 6B baseline high-dose PfSPZ feature importance barplot.xlsx"))
```

```{r arrange shapley plot data from existing 4-fold cv data, fig.align='center', fig.width=8, fig.height=6}
#see vignettes in https://liuyanguu.github.io/post/2019/07/18/visualization-of-shap-for-xgboost/
#note that 0=infected (not protected) and 1 = protected

#filter select data by top n
select_data <- across_fold_perform_highdose_rank_by_AverageRank[c(1:top_n),]

#build shap data from results of each fold
shap_prep_fold1 <- shap_prep_fold2 <- shap_prep_fold3 <- shap_prep_fold4 <- shap_prep_rbind <- shap_plots <- c()
for(k in rev(select_data$run)){
  shap_prep_fold1[[k]] <- shap.prep(xgb_model = fold1[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold1[[k]]$predictor_data[,fold1[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold2[[k]] <- shap.prep(xgb_model = fold2[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold2[[k]]$predictor_data[,fold2[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold3[[k]] <- shap.prep(xgb_model = fold3[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold3[[k]]$predictor_data[,fold3[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold4[[k]] <- shap.prep(xgb_model = fold4[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold4[[k]]$predictor_data[,fold4[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_rbind[[k]] <- rbind(shap_prep_fold1[[k]],
                           shap_prep_fold2[[k]],
                           shap_prep_fold3[[k]],
                           shap_prep_fold4[[k]]) %>%
    group_by(variable) %>%
    mutate(ID = 1:n()) %>%
    ungroup()
  levels(shap_prep_rbind[[k]]$variable) <- gsub("FACS_|_innate_BTM|_highBTM|_monaco|CytokineObsConc_", "", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("PfSPZ", "Pf", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("_of_live_leukocytes", "_of_live_PBMCs", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("_", " ", levels(shap_prep_rbind[[k]]$variable))  
  shap_plots[[k]] <- shap.plot.summary(shap_prep_rbind[[k]], scientific = FALSE)
}

shap_plots_top <- ggarrange(plotlist = shap_plots[4:3], labels = names(shap_plots[4:3]),
                        heights = c(7,4.3),
                        nrow=2, align = "hv", common.legend = TRUE,
                        font.label = list(size = 10, face = "plain", color ="black"),
                        hjust = -0.5)

shap_plots_bottom <- ggarrange(plotlist = shap_plots[2:1], labels = names(shap_plots[2:1]),
                        heights = c(7,5.6),
                        nrow=2, align = "hv", common.legend = TRUE,
                        font.label = list(size = 10, face = "plain", color ="black"),
                        hjust = -0.5,
                        legend = "none")

shap_plots_all <- ggarrange(shap_plots_top, shap_plots_bottom,
                        heights = c(1.02,1),
                        ncol=1, align = "hv", common.legend = TRUE)

shap_plots_all
```

```{r save shap plot data, include=FALSE, message=FALSE, eval=FALSE, echo=FALSE}
foo_ids <- c(rep(names(shap_plots[1]), nrow(shap_plots[[1]]$data)),
             rep(names(shap_plots[2]), nrow(shap_plots[[2]]$data)),
             rep(names(shap_plots[3]), nrow(shap_plots[[3]]$data)),
             rep(names(shap_plots[4]), nrow(shap_plots[[4]]$data)))
foo <- bind_rows(shap_plots[[1]]$data,
          shap_plots[[2]]$data,
          shap_plots[[3]]$data,
          shap_plots[[4]]$data)
foo$model <- foo_ids
foo <- foo %>%
  dplyr::select(model, everything())
writexl::write_xlsx(foo, paste0(plot_dat_dir,"Fig 6B baseline high-dose PfSPZ SHAP plot data.xlsx"))
```

```{r make plot figure 5C, fig.align='center', fig.height=6, fig.width=16}
roc_plot_dat <- roc_dat[select_data[1:4,]$run] 
roc_plots <- ggroc(roc_plot_dat, legacy.axes = TRUE, size = 1) +
  theme_bw() +
  scale_color_brewer("", palette = "Set1", type = "qual") +
  scale_fill_brewer("", palette = "Set1", type = "qual") +
  theme(legend.position = "none") +
  geom_text(data = data.labels[1,], aes(1, 0.55, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE, size = 2.75) +
  geom_text(data = data.labels[2,], aes(1, 0.40, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE, size = 2.75) +
  geom_text(data = data.labels[3,], aes(1, 0.25, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE, size = 2.75) +
  geom_text(data = data.labels[4,], aes(1, 0.1, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE, size = 2.75) +
  theme(plot.margin = unit(c(0.1,3,0.1,3), "cm"))
```

```{r plot figure 6B, fig.align='center', fig.height=6, fig.width=16}
figure_6B_plot <- ggarrange(ggarrange(NULL, feature_importance_plot, NULL, nrow = 3,
                    heights = c(0.25,1,0.25)),
          shap_plots_all,
          ggarrange(roc_plots,
                    my_highdose_confus_mat_plot_all,
                    nrow=2,
                    heights = c(0.6,1)),
          widths = c(1.1,1.2,1),
          nrow=1)
figure_6B_plot
```

```{r save high dose xgb res and shap plots, include=FALSE, echo=FALSE, eval=FALSE}
ggsave(filename = paste0(plotdir, "Figure 6B highdose baseline AUC Confusion Matrix Shap plots for top 4 models JCI revision across folds AverageRank.pdf"), plot=figure_6B_plot, height=6, width=16, device = "pdf")
```

### Examine top n models

```{r examine top n models, fig.align='center', fig.width=13, fig.height=5, message=FALSE, warning=FALSE}
shap_prep_rbind_df <- bind_rows(shap_prep_rbind, .id = "model")

var_appearances <- shap_prep_rbind_df %>%
  group_by(variable) %>%
  mutate(n=n()/63) %>%
  distinct(variable, n) %>%
  arrange(desc(n))

shap_by_feature_dat <-shap_prep_rbind_df %>%
  left_join(., select_data %>%
              rename(model = "run") %>%
              dplyr::select(model, contains("rank")),
            by = "model") %>%
  left_join(., var_appearances,
            by = "variable") %>%
  mutate(model = fct_reorder(model, AverageRank)) %>%
  mutate(variable = fct_reorder2(variable, mean_value, n))

shap_by_feature <- shap_by_feature_dat %>%
  ggplot(., aes(x = value, y = rfvalue)) +
  geom_point(fill = "gray", pch = 21, alpha = 0.3) +
  theme_bw() + 
  ylab("feature value") +
  xlab("SHAP value (impact on model output)") +
  xlim(c(-2,2)) +
  geom_smooth(method='lm', formula= y~x, color = "salmon", alpha = 0.30) +
  stat_cor(method = "pearson", "cor.coef.name" = "R", size = 2.5, label.x.npc = "left", label.y.npc = "top",
           aes(label = paste(..r.label.., cut(..p.., 
                                          breaks = c(-Inf, 0.0001, 0.001, 0.01, 0.05, Inf),
                                          labels = c("'****'", "'***'", "'**'", "'*'", "'ns'")),
                         sep = "~"))) +
  ggh4x::facet_grid2(model~variable, scales = "free_x", labeller = label_wrap_gen(), switch = "y") +
  theme(strip.text.x.top = element_text(angle = 90, hjust = 0, vjust = 0.5),
        strip.clip = "off",
        axis.text = element_blank(),
        axis.ticks = element_blank())
  
shap_by_feature
```

```{r save shap_by_feature, eval=FALSE, echo=FALSE, include=FALSE}
pdf(paste0(plotdir, "Figure S5B highdose baseline feature values by SHAP values by variable by model AverageRank top ", top_n,".pdf"), width = 13, height = 5)
shap_by_feature
dev.off()
```

```{r arrange shapley plot data using de novo training, fig.align='center', fig.width=8, fig.height=20, eval=FALSE, echo=FALSE, include=FALSE}
#see vignettes in https://liuyanguu.github.io/post/2019/07/18/visualization-of-shap-for-xgboost/
library(SHAPforxgboost)
library(ggpubr)

### Run xgb on full data set using the downselected, cross-validated features
#This can be used to create new models de novo for Shapley plots.
source("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/kspzv1-systems-analysis/kspzv1 xgboost manual train and cv function.R")
xgb_res <- c()
set.seed(12345)
for(k in select_data$run){
  my_features <- fold4[[k]]$downselected_xgb_results$feature_names
  xgb_res[[k]] <- xgb_train_on_selected_features(train_features = train_features,
                                                 features_to_test = my_features,
                                                 maxiterations = 2000)
}
#see vignettes in https://liuyanguu.github.io/post/2019/07/18/visualization-of-shap-for-xgboost/
library(SHAPforxgboost)
library(ggpubr)


#note that 0=infected (not protected) and 1 = protected
shap_plot <- c()
for(k in rev(select_data$run)){
  my_features <- xgb_res[[k]]$xgb_results$feature_names
  shap_plot[[k]] <- shap.plot.summary.wrap1(xgb_res[[k]]$xgb_results,
                                           X = as.matrix(train_features[, my_features]))
}
                                           

shap_plots_top <- ggarrange(plotlist = shap_plot[4:3], labels = names(shap_plot[4:3]),
                        heights = c(6,3.3),
                        nrow=2, align = "hv", common.legend = TRUE,
                        font.label = list(size = 10, face = "plain", color ="black"),
                        hjust = -0.5)

shap_plots_bottom <- ggarrange(plotlist = shap_plot[2:1], labels = names(shap_plot[2:1]),
                        heights = c(7,5.7),
                        nrow=2, align = "hv", common.legend = TRUE,
                        font.label = list(size = 10, face = "plain", color ="black"),
                        hjust = -0.5,
                        legend = "none")

shap_plots <- ggarrange(shap_plots_top, shap_plots_bottom,
                        heights = c(1.1,1),
                        ncol=1, align = "hv", common.legend = TRUE)

shap_plots
```

```{r save xgb train results for shap plot, eval=FALSE, echo=FALSE, include=FALSE}
my_number <- 5
save(xgb_res, train_features, file = paste0(
  "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/ML results Tuan/full_high_dose_baseline_4fold_cv_2500_runs/",
  "highdose_baseline_top_4_models_by_AverageRank_xgbres_for_shap_plots_v", my_number,".RData"))
```

### Confirm differences in CD11c+ cells between NP and P in 1.8x10^6 PfSPZ group at baseline

```{r confirm differences in CD11c+ cells, fig.align='center', fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
library(ggpubr)
#load the baseline high dose dataset
#local filename: "highdose_PfSPZ_baseline_correlation_ML_data_with_missing.rds"
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("1PNBLTGLRlcXYuLhJxOWUy6YcLOgaJwDr"), path = temp, overwrite = TRUE)
baseline_highdose_features <- readRDS(file = dl$local_path) %>%
  rownames_to_column(var = "PATID") %>%
  dplyr::select(PATID, contains("CD11c"))

#local filename: "KSPZV1 PhenoData TTE 6 months Long Format.rds"
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("1v2l7G7AzXhgcqObzAabYy6_QJsdB5hFn"), path = temp, overwrite = TRUE)
pheno_data <- readRDS(file = dl$local_path) %>%
  filter(Dosegroup == "1.8 x 10^6 PfSPZ") %>%
  dplyr::select(PATID, tte.mal.atp.3, Outcome, Dosegroup) %>%
  distinct(PATID, tte.mal.atp.3, Outcome, Dosegroup)

facs_dat <- pheno_data %>%
  left_join(., baseline_highdose_features, by = "PATID") %>%
  drop_na(`FACS_CD11c+_of_live_PBMCs`) 

facs_dat_n <- facs_dat %>% 
  group_by(Outcome) %>%
  mutate(n=n()) %>%
  ungroup() %>%
  mutate(outcome = ifelse(Outcome == "infected", paste0("NP (", n, ")"),
                          paste0("P (", n, ")")))

CD11c_plot <- facs_dat_n %>% 
  ggplot(., aes(x = outcome, y = `FACS_CD11c+_of_live_PBMCs`, fill = outcome)) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2, width = 0.2) +
  ylab("CD11c+ (% of live PBMCs)") +
  xlab("outcome (n)") +
  scale_fill_manual(values = c("#A6CEE3", "#1F78B4")) +
  stat_compare_means(method = "wilcox", vjust = 1, hjust = 0.5) +
  theme_bw(base_family = "sans") +
  theme(legend.position = "none",
        axis.text = element_text(size = 12, colour = "black"),
        axis.title = element_text(size = 14))

CD11c_plot
```


```{r save cd11c plot, eval=FALSE, include=FALSE, echo=FALSE}
ggsave(filename = paste0(plotdir, "Figure S5B CD11c FACS NP vs P highdose PfSPZ baseline.pdf"),plot = CD11c_plot, device = "pdf",
       units = "in", width = 4, height = 3)
```

```{r read in data scale and impute, eval=FALSE, echo=FALSE, include=FALSE}
# Apply 1.8x10^6 PfSPZ full model to placebo group
placebo_features_class <- read_rds(file = paste0(datadir2, "placebo_baseline_class_features_scaled_imputed.rds"))
```

```{r make placebo features, eval=FALSE, echo=FALSE, include=FALSE}
set.seed(1234) #set seed for reproducibility
test_features <- placebo_features_class %>%
  dplyr::select(-c(sex, site))

test_dat <- t(test_features) %>%
  data.frame(check.names = FALSE) %>%
  t() %>%
  data.frame(check.names = FALSE) %>%
  mutate(class = factor(class)) %>%
  mutate_at(c(2:ncol(.)), as.numeric)

setDT(test_dat, keep.rownames = TRUE)
test_dat_samplenames <- test_dat$rn
test_dat <- test_dat[,-1]

if(all(colnames(test_dat) == colnames(test_features)) &
   all(test_dat_samplenames == rownames(test_features))){
  print("placebo (test) set good to go!")
  } else {
    print("please check to see if test samples and features match.")
    }
#check missing values
#table(is.na(test_dat))
#sapply(test_dat, function(x) sum(is.na(x))/length(x))*100
#see https://www.hackerearth.com/practice/machine-learning/machine-learning-algorithms/beginners-tutorial-on-xgboost-parameter-tuning-r/tutorial/ for         more data cleaning options

#assign data and labels using one hot encoding 
test_labels <- test_dat$class #Levels: infected protected --> 1 2
new_test <- model.matrix(~.+0, data = test_dat[,-c("class"), with=F])
colnames(new_test) <- gsub("\\`","",colnames(new_test))
#convert factor to numeric 
test_labels <- as.numeric(test_labels)-1 #note that 0=infected (not protected) and 1 = protected using one hot encoding 
#prepare matrix
dtest <- xgb.DMatrix(data = new_test, label = test_labels)

temp_df <- data.frame("class" = placebo_features_class$class, "labels" = test_labels)
table(temp_df$class, temp_df$label) #confirmed that infected = 0, and protected = 1
#here is tricky part. now we want to reverse the classes given that the "protected" group in placebo is hypothesized to resemble the "infected" or "not protected" group in the high-dose PfSPZ infants at baseline. (Remember that "infected/not protected" was originally labeled "0"i n the high-dose group)
temp_df <- temp_df %>%
  mutate(dummy_label = ifelse(labels==0, 1, 0))
placebo_dummy_label <- temp_df$dummy_label
table("actual" = temp_df$class, "dummy" = temp_df$dummy_label)
```

```{r get performance metrics of top highdose PfSPZ models evaluated using placebo as test, eval=FALSE, echo=FALSE, include=FALSE}
## Predict placebo group using high-dose PfSPZ model

#Determine if the best high-dose PfSPZ xgb model, as determined by highest AUC or lowest test mean error, can predict the opposite class in the placebo dataset

select_data <- across_fold_perform_high_dose[1:4,]

xgbDMat_test_placebo <- xgbpred_placebo_dichotomous <- xgbpred_placebo <- confusion_mat_placebo <- c()
roc_dat_placebo <- auc_placebo <- logloss_placebo <- performance_placebo <- brier_placebo <- c()
for(k in select_data$run){
  my_features <- xgb_res[[k]]$xgb_results$feature_names
  xgbDMat_test_placebo[[k]] <- xgb.DMatrix(data = new_test[,my_features],
                                   label = placebo_dummy_label,
                                   nthread = 8)
  xgbpred_placebo[[k]] <- predict(xgb_res[[k]]$xgb_results, xgbDMat_test_placebo[[k]])
  xgbpred_placebo_dichotomous[[k]] <- ifelse(xgbpred_placebo[[k]] > 0.5,1,0)
  xgbpred_placebo_dichotomous[[k]] <- factor(ifelse(xgbpred_placebo_dichotomous[[k]]==1, "protected", "not protected")) #remember that we are going with the original high-dose designation: infected=0, protected=1
  test_labels_for_confusion_matrix <- factor(ifelse(test_labels==1, "protected", "not protected"))
  confusion_mat_placebo[[k]] <- caret::confusionMatrix(xgbpred_placebo_dichotomous[[k]], test_labels_for_confusion_matrix)
  brier_placebo[[k]] <- mlr::measureBrier(xgbpred_placebo[[k]], test_labels_for_confusion_matrix,
                                          "infected", "protected")
  logloss_placebo[[k]] <- MLmetrics::LogLoss(xgbpred_placebo[[k]], as.numeric(test_labels_for_confusion_matrix)-1)
  performance_placebo[[k]] <- confusion_mat[[k]]$overall
  roc_dat_placebo[[k]] <- pROC::roc(response = test_labels,
                            predictor = xgbpred_placebo[[k]],
                            smooth = smooth)
  auc_placebo[[k]] <- roc_dat_placebo[[k]]$auc %>% as.numeric(.) %>% signif(.,digits=3)
}

across_fold_performance_df <- bind_rows(performance_placebo, .id = "run") %>%
  arrange(desc(Kappa))  %>%
  mutate(Kappa_rank = 1:nrow(.)) %>%
  arrange(desc(Accuracy)) %>%
  mutate(Accuracy_rank = 1:nrow(.))
auc_df_placebo <- bind_rows(auc_placebo, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "AUC") %>%
  arrange(desc(AUC))  %>%
  mutate(AUC_rank = 1:nrow(.))
logloss_df_placebo <- bind_rows(logloss_placebo, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "LogLoss") %>%
  arrange(LogLoss)  %>%
  mutate(LogLoss_rank = 1:nrow(.))
brier_df_placebo <- bind_rows(brier_placebo, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "BrierScore") %>%
  mutate(BrierScore = signif(BrierScore, 3)) %>%
  arrange(BrierScore) %>%
  mutate(BrierScore_rank = 1:nrow(.))

across_fold_perform_placebo <- auc_df_placebo %>%
  left_join(., across_fold_performance_df, by = "run") %>%
  left_join(., summarized_res_2500_runs %>%
              dplyr::select(run, Features),
            by = "run") %>%
  left_join(., logloss_df_placebo,
            by = "run") %>%
  left_join(., brier_df_placebo,
            by = "run") %>%
  mutate(AverageRank = (Accuracy_rank + AUC_rank + BrierScore_rank + Kappa_rank + LogLoss_rank)/5) %>%
  dplyr::select(run, Features, BrierScore_rank, Kappa_rank, LogLoss_rank, AUC_rank, AverageRank, Accuracy_rank, AUC, BrierScore, LogLoss, Kappa, Accuracy:McnemarPValue) %>%
  group_by(run) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  arrange(BrierScore_rank)
```

```{r save highdose models on placebo data performance metrics, eval=FALSE, echo=FALSE, include=FALSE}
resdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/ML results Tuan/"
writexl::write_xlsx(across_fold_perform_placebo,
                    path = paste0(
  "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Supplementary Tables JCI resubmission/",
                                                    "performrance_metrics_best_highdose_models_on_placebo",
                                                    gsub("\\:","-", format(Sys.time(),"%a-%b-%d-%X-%Y")),".xlsx"))
```

```{r xgb downselected features on full placebo dataset, eval=FALSE, echo=FALSE, include=FALSE}
## Train downselected features on placebo data

source("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/kspzv1-systems-analysis/kspzv1 xgboost manual train and cv function.R")
select_placebo_models <- across_fold_perform_placebo[1:4,]

xgb_placebo_res <- c()
for(k in select_placebo_models$run){
  my_features <- fold4[[k]]$downselected_xgb_results$feature_names
  xgb_placebo_res[[k]] <- xgb_train_on_selected_features(train_features = test_features,
                                                 features_to_test = my_features,
                                                 maxiterations = 1000)
}
```

```{r prepare shapley plots placebo, fig.align='center', fig.width=6, fig.height=5, eval=FALSE, echo=FALSE, include=FALSE}
## Plot SHAP for placebo

#see vignettes in https://liuyanguu.github.io/post/2019/07/18/visualization-of-shap-for-xgboost/
library(SHAPforxgboost)
library(ggpubr)


#note that 0=infected (not protected) and 1 = protected
shap_placebo_plot <- c()
for(k in select_placebo_models$run){
  my_features <- xgb_placebo_res[[k]]$xgb_results$feature_names
  shap_placebo_plot[[k]] <- shap.plot.summary.wrap1(xgb_placebo_res[[k]]$xgb_results,
                                           X = as.matrix(train_features[, my_features]))
}
                                           
shap_placebo_plots <- ggarrange(plotlist = shap_placebo_plot, labels = names(shap_placebo_plot),
                                heights = c(1,1,0.7,1,0.7),
                                nrow=5, align = "hv", common.legend = TRUE)
```

```{r plot shapley plots placebo, fig.align='center', fig.width=6, fig.height=8, eval=FALSE, echo=FALSE, include=FALSE}
shap_placebo_plots
```

```{r save placebo and shap plots, eval=FALSE, echo=FALSE, include=FALSE}
ggsave(filename = paste0(plotdir, "highdose_on_placebo_full_ds_shap_plots_Brier.pdf"), plot=shap_placebo_plots, height=5.5, width=5.85, device = "pdf")
```

```{r prepare roc data, eval=FALSE, echo=FALSE, include=FALSE}
### Make ROC curves for high-dose PfSPZ using models that work well for placebo

runs_to_test <- ml_res[ml_res$mean_cv_accuracy>.80,]$run #select only the high-dose PfSPZ models with accuracy >80%.
#runs_to_test <- c("run 211", "run 2222", "run 671")
smooth <- FALSE
best_run <- best_combo <- all_folds_response <- all_folds_predictor <- roc_dat <- auc <- confusion_mat <- response_dichotomous <- c()
for(j in runs_to_test){
  best_run[[j]] <- per_run_res_2500_runs[per_run_res_2500_runs$run==j,] %>%
    group_by(run) %>%
    slice_max(AUC) %>%
    distinct(Features) %>%
    pull(var = run)
  best_combo[[j]] <- per_run_res_2500_runs[per_run_res_2500_runs$run==j,] %>%
    group_by(run) %>%
    slice_max(AUC) %>%
    distinct(Features) %>%
    pull(var = Features)
  #calculate predictive accuracy across all 4 folds
  all_folds_response[[j]] <- c(fold1[[best_run[[j]]]]$predictor_data$class,
                               fold2[[best_run[[j]]]]$predictor_data$class,
                               fold3[[best_run[[j]]]]$predictor_data$class,
                               fold4[[best_run[[j]]]]$predictor_data$class)
  all_folds_predictor[[j]] <- c(fold1[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold2[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold3[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold4[[best_run[[j]]]]$predictor_data$pred.prob)
  roc_dat[[j]] <- pROC::roc(response = all_folds_response[[j]],
                            predictor = all_folds_predictor[[j]],
                            smooth = smooth) 
  auc[[j]] <- roc_dat[[j]]$auc %>% as.numeric(.) %>% signif(.,digits=3)
  response_dichotomous[[j]]  <- ifelse(all_folds_predictor[[j]]  > 0.5,1,0)
  #remember that we are going with the original high-dose designation where infected=0, protected=1.
  response_dichotomous[[j]]   <- factor(ifelse(response_dichotomous[[j]]==1, "protected", "infected")) 
  confusion_mat[[j]] <- caret::confusionMatrix(response_dichotomous[[j]], all_folds_response[[j]] )
}

auc_df_high_dose <- bind_rows(auc, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "AUC") %>%
  arrange(desc(AUC))
```

```{r plot ROC, eval=FALSE, echo=FALSE, include=FALSE}
data.labels <- data.frame("name" = unlist(best_run),
                          "features" = unlist(best_combo),
                          "auc" = unlist(auc))
#roc_dat[1:4]
roc_plots <- ggroc(roc_dat[1:3], legacy.axes = TRUE, size = 1.5) +
  theme_bw() +
  scale_color_brewer("", palette = 7, type = "qual") +
  scale_fill_brewer("", palette = 7, type = "qual") +
  theme(legend.position = "right") +
  geom_text(data = data.labels[1,], aes(0.875, 0.2, label = paste("AUC =", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[2,], aes(0.875, 0.15, label = paste("AUC =", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[3,], aes(0.875, 0.1, label = paste("AUC =", auc)), hjust=1, show_guide=FALSE) +
  #geom_text(data = data.labels[4,], aes(0.875, 0.05, label = paste("AUC =", auc)), hjust=1, show_guide=FALSE) +
  theme(plot.margin = unit(c(0.5,2,0.5,0.5), "cm")) +
  geom_abline(slope = 1, linetype = "dotted", color = "gray")

roc_plots
```

```{r plot confusion matrix, eval=FALSE, echo=FALSE, include=FALSE}
### Plot confusion matrix

#https://stackoverflow.com/questions/46063234/how-to-produce-a-confusion-matrix-and-find-the-misclassification-rate-of-the-naï
#https://stats.stackexchange.com/questions/82162/cohens-kappa-in-plain-english#:~:text=%3C0%20as%20indicating%20no%20agreement,1%20as%20almost%20perfect%20agreement.&text=%3E0.75%20as%20excellent%2C%200.40%20to,and%20%3C0.40%20as%20poor.%22

library(ggplot2)
library(scales)

ggplotConfusionMatrix <- function(m){
  mytitle <- paste("Accuracy", percent_format()(m$overall[1]),
                   "Kappa", percent_format()(m$overall[2]))
  p <-
    ggplot(data = as.data.frame(m$table) ,
           aes(x = Reference, y = Prediction)) +
    geom_tile(aes(fill = log(Freq)), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue") +
    geom_text(aes(x = Reference, y = Prediction, label = Freq)) +
    theme(legend.position = "none") +
    ggtitle(mytitle)
  return(p)
}

ggplotConfusionMatrix(confusion_mat$`run 671`)
```