---
title: "KSPZV1 delta high-dose PfSPZ multimodal ML- visualize xgboost ML results, JCI revision (Figure 6D)"
author: "Tuan M. Tran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document :
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r links discussing approaches, include=FALSE, eval=FALSE}
#Links discussing best approaches:
# https://stats.stackexchange.com/questions/264533/how-should-feature-selection-and-hyperparameter-optimization-be-ordered-in-the-m
```

# Objective

Visualization of xgboost ML with 4-fold cross-validation of full dataset.

# Approach

Plot ROC curves, confusion matrices, and SHAP plots from top n models using xgboost results derived from "KPSZV1 Delta Highdose PfSPZ Multimodal ML xgb nested fs cv on full set JCI Revision.Rmd".

```{r libraries, message=FALSE}
library(xgboost)
library(dplyr)
library(tidyverse)
library(pROC)
library(caret)
library(doMC)
library(data.table)
library(xgboost)
library(SHAPforxgboost)
library(mlr)
library(ggpubr)
library(tictoc)
library(Biobase)
library(googledrive)
library(parallel)
library(parallelMap)
parallelStartSocket(cpus = detectCores())
```

```{r set local paths, eval=FALSE, echo=FALSE, include=FALSE}
datadir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/kspzv1-systems-analysis/"
plotdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Figures JCI resubmission/"
datadir2 <- "/Users/tuantran/Library/CloudStorage/GoogleDrive-tuantran@iu.edu/Shared drives/[Sec] IN-MED-INF-TRANLAB/Tran Lab Shared/Projects/Doris Duke PfSPZ Kenya/Tuan PfSPZ/KenyaPfSPZ/Final Data Files/"
plot_dat_dir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Data for all Plots JCI Resubmission/"
```

```{r read in full dataset, eval=FALSE, echo=FALSE, include=FALSE}
high_dose_features_class <- readRDS(paste0(datadir2, "high_dose_delta_class_features_scaled_imputed.rds"))
train_features <- high_dose_features_class %>%
  dplyr::select(-c(sex, site))
```

### Load feature data

```{r read high dose postvax features from google drive, message=FALSE, warning=FALSE}
#from google drive
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("1QI4BIKcW-5oQfHl-o3t7Hj58OaKl8Mn8"), path = temp, overwrite = TRUE)
high_dose_features_class <- readRDS(file = dl$local_path)
train_features <- high_dose_features_class %>%
  dplyr::select(-c(sex, site))
```

### Load ML results

This is a large file.

```{r plot ROC curve from cv results for best iteration local, eval=FALSE, include=FALSE, echo=FALSE, message=FALSE}
#readin data saved from above
datadir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/ML results Tuan/"
 
all_folds_2500_runs <- load(paste0(datadir,
                                  "full_high_dose_delta_4fold_cv_2500_runs/",
                                  "all_folds_data_seed_highdose_delta_seed_2209_runs_2500_13_Oct_2023_full.RData"))
per_run_res_2500_runs <- readxl::read_xlsx(paste0(datadir,
                                  "full_high_dose_delta_4fold_cv_2500_runs/",
                                  "cv_results_all_runs_all_folds_highdose_delta_seed_2209_runs_2500_htune_iters_500_13_Oct_2023_full.xlsx"))
summarized_res_2500_runs <- readxl::read_xlsx(paste0(datadir,
                                  "full_high_dose_delta_4fold_cv_2500_runs/",
                                  "cv_summarized_res_all_folds_highdose_delta_seed_2209_runs_2500_htune_iters_500_13_Oct_2023_full.xlsx"))
```

```{r plot ROC curve from cv results for best iteration google drive, message=FALSE, warning=FALSE}
#from google drive
temp <- tempfile(fileext = ".Rdata")
dl <- drive_download(
  as_id("1R0xB2u_7nMdszP0q2glaouZjb3ihmYPU"), path = temp, overwrite = TRUE)
all_folds_2500_runs <- load(file = dl$local_path)

temp <- tempfile(fileext = ".xlsx")
dl <- drive_download(
  as_id("1SVbsvL-SK4zDjHle-LuGIh0qm4gAZf2x"), path = temp, overwrite = TRUE)
per_run_res_2500_runs <- readxl::read_xlsx(dl$local_path)

temp <- tempfile(fileext = ".xlsx")
dl <- drive_download(
  as_id("1R83aGAo2jfkf48MWyeF37DUbV2LRZDdt"), path = temp, overwrite = TRUE)
summarized_res_2500_runs <- readxl::read_xlsx(dl$local_path)
```

### Determine across-fold performance metrics (accuracy, AUC, Brier score, kappa, logless) select best performing high-dose PfSPZ delta models

```{r determine across fold AUC and kappa, message=FALSE, warning=FALSE}
smooth <- FALSE
runs_to_test <- summarized_res_2500_runs[summarized_res_2500_runs$mean_cv_accuracy>.66,]$run #select only the high-dose PfSPZ models with accuracy >66%.
#runs_to_test <- c("run 211", "run 2222", "run 671")

best_run <- best_combo <- all_folds_response <- all_folds_predictor <- roc_dat <- auc <- confusion_mat <- response_dichotomous <- c()
across_fold_performance <- logloss_list <- brier_list <- c()
for(j in runs_to_test){
  best_run[[j]] <- per_run_res_2500_runs[per_run_res_2500_runs$run==j,] %>%
    group_by(run) %>%
    slice_max(AUC) %>%
    distinct(Features) %>%
    pull(var = run)
  best_combo[[j]] <- per_run_res_2500_runs[per_run_res_2500_runs$run==j,] %>%
    group_by(run) %>%
    slice_max(AUC) %>%
    distinct(Features) %>%
    pull(var = Features)
  #calculate predictive accuracy across all 4 folds
  all_folds_response[[j]] <- c(fold1[[best_run[[j]]]]$predictor_data$class,
                               fold2[[best_run[[j]]]]$predictor_data$class,
                               fold3[[best_run[[j]]]]$predictor_data$class,
                               fold4[[best_run[[j]]]]$predictor_data$class)
  all_folds_predictor[[j]] <- c(fold1[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold2[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold3[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold4[[best_run[[j]]]]$predictor_data$pred.prob)
  roc_dat[[j]] <- pROC::roc(response = all_folds_response[[j]],
                            predictor = all_folds_predictor[[j]],
                            smooth = smooth) 
  auc[[j]] <- roc_dat[[j]]$auc %>% as.numeric(.) %>% signif(.,digits=3)
  response_dichotomous[[j]]  <- ifelse(all_folds_predictor[[j]]  > 0.5,1,0)
  #remember that since we are using the high-dose model, we are going with the original high-dose designation where infected=0, protected=1
  #but we are actually predicting the opposite for placebo, where 1 means infected(not protected), and infected=0
  response_dichotomous[[j]]   <- factor(ifelse(response_dichotomous[[j]]==1, "protected", "infected")) 
  confusion_mat[[j]] <- caret::confusionMatrix(response_dichotomous[[j]], all_folds_response[[j]], positive = "protected")
  logloss_list[[j]] <- MLmetrics::LogLoss(all_folds_predictor[[j]], as.numeric(all_folds_response[[j]])-1)
  brier_list[[j]] <- mlr::measureBrier(all_folds_predictor[[j]], all_folds_response[[j]], "infected", "protected")
  across_fold_performance[[j]] <- confusion_mat[[j]]$overall
} 

across_fold_performance_df <- bind_rows(across_fold_performance, .id = "run") %>%
  arrange(desc(Kappa))  %>%
  mutate(Kappa_rank = 1:nrow(.)) %>%
  arrange(desc(Accuracy)) %>%
  mutate(Accuracy_rank = 1:nrow(.))
auc_df_high_dose <- bind_rows(auc, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "AUC") %>%
  arrange(desc(AUC))  %>%
  mutate(AUC_rank = 1:nrow(.))
logloss_df <- bind_rows(logloss_list, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "LogLoss") %>%
  arrange(LogLoss)  %>%
  mutate(LogLoss_rank = 1:nrow(.))
brier_df <- bind_rows(brier_list, .id = "run") %>%
  pivot_longer(., cols = everything(), names_to = "run", values_to = "BrierScore") %>%
  mutate(BrierScore = signif(BrierScore, 3)) %>%
  arrange(BrierScore) %>%
  mutate(BrierScore_rank = 1:nrow(.))

across_fold_perform_high_dose <- auc_df_high_dose %>%
  left_join(., across_fold_performance_df, by = "run") %>%
  left_join(., summarized_res_2500_runs,
            by = "run") %>%
  left_join(., logloss_df,
            by = "run") %>%
  left_join(., brier_df,
            by = "run") %>%
  mutate(AverageRank = (Accuracy_rank + AUC_rank + BrierScore_rank + Kappa_rank + LogLoss_rank)/5) %>%
  dplyr::select(run, Features, AverageRank, BrierScore_rank, Kappa_rank, LogLoss_rank, AUC_rank, Accuracy_rank, AUC, BrierScore, LogLoss, Kappa, Accuracy:McnemarPValue, mean_cv_AUC:sd_cv_accuracy) %>%
  group_by(run) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  arrange(AverageRank)

#clean up feature names
  # across_fold_perform_high_dose <- across_fold_perform_high_dose %>%
  #   mutate(Features =  gsub("CSP_specific_IgG_", "CSP-specific IgG_", Features)) %>% 
  #   mutate(Features =  gsub("FACS_|_innate_BTM|_highBTM|_monaco|CytokineObsConc_", "", Features)) %>%   
  #   mutate(Features =  gsub("PfSPZ", "Pf", Features)) %>% 
  #   mutate(Features =  gsub("_of_live_leukocytes", "_of_live_PBMCs", Features)) %>% 
  #   mutate(Features =  gsub("_", " ", Features))
```


```{r save kappa values, include=FALSE, echo=FALSE, eval=FALSE}
writexl::write_xlsx(across_fold_perform_high_dose, paste0("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Supplementary Tables JCI resubmission/","Table S6 across folds xgb performance highdose delta.xlsx"))
```

### Arrange data and plot AUROC curves 

```{r arrange data for roc plots, message=FALSE, warning=FALSE}
top_n <- 4
across_fold_perform_highdose_rank_by_AverageRank <- across_fold_perform_high_dose %>%
  arrange(AverageRank) %>%
  drop_na()
select_data <- across_fold_perform_highdose_rank_by_AverageRank[c(1:top_n),]
data.labels <- data.frame("name" = select_data[1:4,]$run,
                          "features" = select_data[1:4,]$Features,
                          "brier" = select_data[1:4,]$BrierScore,
                          "auc" = select_data[1:4,]$AUC,
                          "kappa" = select_data[1:4,]$Kappa,
                          "logloss" = select_data[1:4,]$LogLoss)
roc_plot_dat <- roc_dat[select_data[1:4,]$run] 
roc_plots <- ggroc(roc_plot_dat, legacy.axes = TRUE, size = 1.5) +
  theme_bw() +
  scale_color_brewer("", palette = "Set1", type = "qual") +
  scale_fill_brewer("", palette = "Set1", type = "qual") +
  theme(legend.position = "none") +
  geom_text(data = data.labels[1,], aes(0.875, 0.25, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[2,], aes(0.875, 0.2, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[3,], aes(0.875, 0.15, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[4,], aes(0.875, 0.1, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE) +
  #geom_text(data = data.labels[5,], aes(0.875, 0.05, label = paste("AUC =", auc)), hjust=1, show_guide=FALSE) +
  theme(plot.margin = unit(c(1,1,1,1), "cm")) +
  geom_abline(slope = 1, linetype = "dotted", color = "gray")
```

```{r plot roc, fig.align='center', fig.height=6, fig.width=6}
roc_plots
```

```{r save rocplot data, include=FALSE, message=FALSE, eval=FALSE, echo=FALSE}
writexl::write_xlsx(roc_plots$data, paste0(plot_dat_dir, "Fig 6D delta high-dose PfSPZ ROC plot.xlsx"))
```

### Plot confusion matrices for high-dose delta models

```{r confusion matrix plots web ref, echo=FALSE, include=FALSE, eval=FALSE}
#https://stackoverflow.com/questions/46063234/how-to-produce-a-confusion-matrix-and-find-the-misclassification-rate-of-the-naï
#https://stats.stackexchange.com/questions/82162/cohens-kappa-in-plain-english#:~:text=%3C0%20as%20indicating%20no%20agreement,1%20as%20almost%20perfect%20agreement.&text=%3E0.75%20as%20excellent%2C%200.40%20to,and%20%3C0.40%20as%20poor.%22
```

```{r make confusion matrix plots highdose delta, message=FALSE, warning=FALSE}
library(ggplot2)
library(scales)

ggplotConfusionMatrix <- function(m){
  mytitle <- paste(
    "accuracy", percent_format()(m$overall[1]),
                   "kappa", percent_format()(m$overall[2]))
  mydata <- as.data.frame(m$table) %>%
    mutate(Prediction = ifelse(Prediction == "infected", "NP", "P")) %>%
    mutate(Reference = ifelse(Reference == "infected", "NP", "P"))
  p <- ggplot(data = mydata,
           aes(x = Reference, y = Prediction)) +
    geom_tile(aes(fill = log(Freq)), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue") +
    geom_text(aes(x = Reference, y = Prediction, label = Freq, size = 24)) +
    ggtitle(mytitle) +
    ylab("predicted") +
    xlab("actual outcome") +
    theme(legend.position = "none",
          panel.background = element_blank(),
          plot.background = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_text(color = "black", size = 10),
          axis.text = element_text(color = "black", size = 10),
          title = element_text(size = 9),) +
    theme(plot.margin = unit(c(0.5,0.45,0.01,0.4), "cm"))
  return(p)
}
```

```{r plot highdose confusion matrices, fig.align='center', fig.height=2, fig.width=8}
#my_models <- c("run 1727", "run 671", "run 808", "run 439") #ranked by Brier
my_models <- select_data[1:4,]$run #ranked by average rank

my_plots <- list(ggplotConfusionMatrix(confusion_mat[[my_models[1]]]),
              ggplotConfusionMatrix(confusion_mat[[my_models[2]]]),
              ggplotConfusionMatrix(confusion_mat[[my_models[3]]]),
              ggplotConfusionMatrix(confusion_mat[[my_models[4]]]))
names(my_plots) <- my_models

my_highdose_confus_mat_plot_top <- ggarrange(plotlist = my_plots[1:2],
                                         labels = names(my_plots[1:2]),
                                         font.label = list(size=14, face="plain"),
                                         ncol = 2,
                                         hjust = -0.1)
my_highdose_confus_mat_plot_bottom <- ggarrange(plotlist = my_plots[3:4],
                                         labels = names(my_plots[3:4]),
                                         font.label = list(size=14, face="plain"),
                                         ncol = 2,
                                         hjust = -0.1)
my_highdose_confus_mat_plot_all <- ggarrange(my_highdose_confus_mat_plot_top,
                                             my_highdose_confus_mat_plot_bottom,
                                             nrow=2)
```

```{r plot my_highdose_confus_mat_plot_all, fig.align='center', fig.height=6, fig.width=5}
my_highdose_confus_mat_plot_all
```

```{r save confusion matrix data, include=FALSE, message=FALSE, eval=FALSE, echo=FALSE}
foo_ids <- c(rep(names(my_plots[1]), 4),
             rep(names(my_plots[2]), 4),
             rep(names(my_plots[3]), 4),
             rep(names(my_plots[4]), 4))
foo <- bind_rows(my_plots[[1]]$data,
          my_plots[[2]]$data,
          my_plots[[3]]$data,
          my_plots[[4]]$data)
foo$model <- foo_ids
writexl::write_xlsx(foo, paste0(plot_dat_dir,"Fig 6D delta high-dose PfSPZ confusion matrices.xlsx"))
```

### Feature importance by number of appearances in top %1 models

```{r feature importance by number of appearances in top n models, fig.align='center', fig.width=5, fig.height=4.5}
addSmallLegend <- function(myPlot, pointSize = 0.75, textSize = 6, spaceLegend = 0.75) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

#filter select data by performance metric
select_data <- across_fold_perform_highdose_rank_by_AverageRank %>%
  arrange(AverageRank) %>%
  slice_head(n=25)

#build shap data from results of each fold
shap_prep_fold1 <- shap_prep_fold2 <- shap_prep_fold3 <- shap_prep_fold4 <- shap_prep_rbind <- shap_plots <- c()
for(k in rev(select_data$run)){
  shap_prep_fold1[[k]] <- shap.prep(xgb_model = fold1[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold1[[k]]$predictor_data[,fold1[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold2[[k]] <- shap.prep(xgb_model = fold2[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold2[[k]]$predictor_data[,fold2[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold3[[k]] <- shap.prep(xgb_model = fold3[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold3[[k]]$predictor_data[,fold3[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold4[[k]] <- shap.prep(xgb_model = fold4[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold4[[k]]$predictor_data[,fold4[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_rbind[[k]] <- rbind(shap_prep_fold1[[k]],
                           shap_prep_fold2[[k]],
                           shap_prep_fold3[[k]],
                           shap_prep_fold4[[k]]) %>%
    group_by(variable) %>%
    mutate(ID = 1:n()) %>%
    ungroup()
  levels(shap_prep_rbind[[k]]$variable) <- gsub("FACS_|_innate_BTM|_highBTM|_monaco|CytokineObsConc_", "", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("PfSPZ", "Pf", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("_of_live_leukocytes", "_of_live_PBMCs", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("_", " ", levels(shap_prep_rbind[[k]]$variable))  
  shap_plots[[k]] <- shap.plot.summary(shap_prep_rbind[[k]], scientific = FALSE)
}
shap_prep_rbind_df <- bind_rows(shap_prep_rbind, .id = "model")

number_models <- sum(summary(factor(unique(shap_prep_rbind_df$model))))

var_appearances <- shap_prep_rbind_df %>%
  dplyr::select(model,variable) %>%
  distinct() %>%
  group_by(variable) %>%
  summarize(proportion=n()/number_models, n=n()) %>%
  ungroup() %>%
  arrange(desc(n))

foo <- shap_long_iris %>%
  group_by(variable) %>%
  summarize(mean = mean(value))


SHAP_feature_correlations <- shap_prep_rbind_df %>%
  dplyr::select(model,variable, value, rfvalue) %>%
  distinct() %>%
  group_by(model, variable) %>% 
  summarise(r = cor(value, rfvalue)) %>%
  ungroup() %>%
  group_by(variable) %>%
  summarize(mean_r=mean(r))

var_appearances <- var_appearances %>%
  left_join(., SHAP_feature_correlations,
            by = "variable")

feature_importance_plot <- var_appearances %>%
  filter(n>1) %>%
  ggplot(., aes(x= fct_reorder(variable, n, .desc = FALSE))) +
  geom_bar(aes(y=proportion, fill=mean_r), stat="identity") +
  scale_fill_gradient2(low = "darkblue", mid = "white",
                       high = "darkred",
                       midpoint = 0,limits = c(-1,1)) +
  theme_bw() +
  coord_flip() +
  xlab("features") +
  ylab("appearances") + 
  labs(fill = "R") +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) +
  theme(legend.position = "top",
        axis.text = element_text(colour = "black"))

feature_importance_plot <- addSmallLegend(feature_importance_plot)
feature_importance_plot
```

```{r save figure 5X, include=FALSE, message=FALSE, echo=FALSE, eval=FALSE}
pdf(paste0(plotdir, "Figure 6D highdose delta feature by pct appearance barplot AverageRank.pdf"),
    width = 5.9, height = 3.4)
addSmallLegend(feature_importance_plot)
dev.off()
```

```{r save feature importance plot data, include=FALSE, message=FALSE, eval=FALSE, echo=FALSE}
writexl::write_xlsx(feature_importance_plot$data, paste0(plot_dat_dir,"Fig 6D delta high-dose PfSPZ feature importance barplot.xlsx"))
```

```{r arrange shapley plot data from existing 4-fold cv data, fig.align='center', fig.width=8, fig.height=6}
#see vignettes in https://liuyanguu.github.io/post/2019/07/18/visualization-of-shap-for-xgboost/
#note that 0=infected (not protected) and 1 = protected

#filter select data by top n
select_data <- across_fold_perform_highdose_rank_by_AverageRank[c(1:top_n),]

#build shap data from results of each fold
shap_prep_fold1 <- shap_prep_fold2 <- shap_prep_fold3 <- shap_prep_fold4 <- shap_prep_rbind <- shap_plots <- c()
for(k in rev(select_data$run)){
  shap_prep_fold1[[k]] <- shap.prep(xgb_model = fold1[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold1[[k]]$predictor_data[,fold1[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold2[[k]] <- shap.prep(xgb_model = fold2[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold2[[k]]$predictor_data[,fold2[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold3[[k]] <- shap.prep(xgb_model = fold3[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold3[[k]]$predictor_data[,fold3[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_fold4[[k]] <- shap.prep(xgb_model = fold4[[k]]$downselected_xgb_results,
                                    X_train = as.matrix(fold4[[k]]$predictor_data[,fold4[[k]]$downselected_xgb_results$feature_names]))
  shap_prep_rbind[[k]] <- rbind(shap_prep_fold1[[k]],
                           shap_prep_fold2[[k]],
                           shap_prep_fold3[[k]],
                           shap_prep_fold4[[k]]) %>%
    group_by(variable) %>%
    mutate(ID = 1:n()) %>%
    ungroup()
  levels(shap_prep_rbind[[k]]$variable) <- gsub("FACS_|_innate_BTM|_highBTM|_monaco|CytokineObsConc_", "", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("PfSPZ", "Pf", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("_of_live_leukocytes", "_of_live_PBMCs", levels(shap_prep_rbind[[k]]$variable))
  levels(shap_prep_rbind[[k]]$variable) <- gsub("_", " ", levels(shap_prep_rbind[[k]]$variable))  
  shap_plots[[k]] <- shap.plot.summary(shap_prep_rbind[[k]], scientific = FALSE)
}

shap_plots_top <- ggarrange(plotlist = shap_plots[4:3], labels = names(shap_plots[4:3]),
                        heights = c(4.3,7),
                        nrow=2, align = "hv", common.legend = TRUE,
                        font.label = list(size = 10, face = "plain", color ="black"),
                        hjust = -0.5)

shap_plots_bottom <- ggarrange(plotlist = shap_plots[2:1], labels = names(shap_plots[2:1]),
                        heights = c(6,6),
                        nrow=2, align = "hv", common.legend = TRUE,
                        font.label = list(size = 10, face = "plain", color ="black"),
                        hjust = -0.5,
                        legend = "none")

shap_plots_all <- ggarrange(shap_plots_top, shap_plots_bottom,
                        heights = c(1.025,1),
                        ncol=1, align = "hv", common.legend = TRUE)

shap_plots_all
```

```{r save shap plot data, include=FALSE, message=FALSE, eval=FALSE, echo=FALSE}
foo_ids <- c(rep(names(shap_plots[1]), nrow(shap_plots[[1]]$data)),
             rep(names(shap_plots[2]), nrow(shap_plots[[2]]$data)),
             rep(names(shap_plots[3]), nrow(shap_plots[[3]]$data)),
             rep(names(shap_plots[4]), nrow(shap_plots[[4]]$data)))
foo <- bind_rows(shap_plots[[1]]$data,
          shap_plots[[2]]$data,
          shap_plots[[3]]$data,
          shap_plots[[4]]$data)
foo$model <- foo_ids
foo <- foo %>%
  dplyr::select(model, everything())
writexl::write_xlsx(foo, paste0(plot_dat_dir,"Fig 6D delta high-dose PfSPZ SHAP plot data.xlsx"))
```

```{r make plot figure 5C, fig.align='center', fig.height=6, fig.width=16}
roc_plot_dat <- roc_dat[select_data[1:4,]$run] 
roc_plots <- ggroc(roc_plot_dat, legacy.axes = TRUE, size = 1) +
  theme_bw() +
  scale_color_brewer("", palette = "Set1", type = "qual") +
  scale_fill_brewer("", palette = "Set1", type = "qual") +
  theme(legend.position = "none") +
  geom_text(data = data.labels[1,], aes(1, 0.55, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE, size = 2.75) +
  geom_text(data = data.labels[2,], aes(1, 0.40, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE, size = 2.75) +
  geom_text(data = data.labels[3,], aes(1, 0.25, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE, size = 2.75) +
  geom_text(data = data.labels[4,], aes(1, 0.1, label = paste0(name, ", Brier=", brier, ", AUC=", auc)), hjust=1, show_guide=FALSE, size = 2.75) +
  theme(plot.margin = unit(c(0.1,3,0.1,3), "cm"))
```

```{r plot figure 5C, fig.align='center', fig.height=6, fig.width=16}
figure_5D_plot <- ggarrange(ggarrange(NULL, feature_importance_plot, NULL, nrow = 3,
                    heights = c(0.2,1,0.2)),
          shap_plots_all,
          ggarrange(roc_plots,
                    my_highdose_confus_mat_plot_all,
                    nrow=2,
                    heights = c(0.6,1)),
          widths = c(1.1,1.2,1),
          nrow=1)
figure_5D_plot
```

```{r save high dose xgb res and shap plots, include=FALSE, echo=FALSE, eval=FALSE}
ggsave(filename = paste0(plotdir, "Figure 6D highdose delta AUC Confusion Matrix Shap plots for top 4 models JCI revision across folds AverageRank.pdf"), plot=figure_5D_plot, height=6, width=16, device = "pdf")
```

### Examine top n models

```{r examine top n models, fig.align='center', fig.width=13, fig.height=5, message=FALSE, warning=FALSE}
shap_prep_rbind_df <- bind_rows(shap_prep_rbind, .id = "model")

var_appearances <- shap_prep_rbind_df %>%
  group_by(variable) %>%
  mutate(n=n()/63) %>%
  distinct(variable, n) %>%
  arrange(desc(n))

shap_by_feature_dat <-shap_prep_rbind_df %>%
  left_join(., select_data %>%
              rename(model = "run") %>%
              dplyr::select(model, contains("rank")),
            by = "model") %>%
  left_join(., var_appearances,
            by = "variable") %>%
  mutate(model = fct_reorder(model, AverageRank)) %>%
  mutate(variable = fct_reorder2(variable, mean_value, n))

shap_by_feature <- shap_by_feature_dat %>%
  ggplot(., aes(x = value, y = rfvalue)) +
  geom_point(fill = "gray", pch = 21, alpha = 0.3) +
  theme_bw() + 
  ylab("feature value") +
  xlab("SHAP value (impact on model output)") +
  xlim(c(-2,2)) +
  geom_smooth(method='lm', formula= y~x, color = "salmon", alpha = 0.30) +
  stat_cor(method = "pearson", "cor.coef.name" = "R", size = 2.5, label.x.npc = "left", label.y.npc = "top",
           aes(label = paste(..r.label.., cut(..p.., 
                                          breaks = c(-Inf, 0.0001, 0.001, 0.01, 0.05, Inf),
                                          labels = c("'****'", "'***'", "'**'", "'*'", "'ns'")),
                         sep = "~"))) +
  ggh4x::facet_grid2(model~variable, scales = "free_x", labeller = label_wrap_gen(), switch = "y") +
  theme(strip.text.x.top = element_text(angle = 90, hjust = 0, vjust = 0.5),
        strip.clip = "off",
        axis.text = element_blank(),
        axis.ticks = element_blank())
  
shap_by_feature
```

```{r save shap_by_feature, eval=FALSE, echo=FALSE, include=FALSE}
pdf(paste0(plotdir, "Figure S5E highdose postvax feature values by SHAP values by variable by model AverageRank top ", top_n,".pdf"), width = 13, height = 5)
shap_by_feature
dev.off()
```

```{r xgb downselected features on full dataset}
### Run xgb on full data set using the downselected, cross-validated features
#This can be used to create new models de novo for Shapley plots.
source("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/kspzv1-systems-analysis/kspzv1 xgboost manual train and cv function.R")
xgb_res <- c()
set.seed(12345)
#for(k in select_data$run){
for(k in my_models){
  my_features <- fold4[[k]]$downselected_xgb_results$feature_names
  xgb_res[[k]] <- xgb_train_on_selected_features(train_features = train_features,
                                                 features_to_test = my_features,
                                                 maxiterations = 2000)
}
#see vignettes in https://liuyanguu.github.io/post/2019/07/18/visualization-of-shap-for-xgboost/
#note that 0=infected (not protected) and 1 = protected

shap_plot <- c()
#for(k in rev(select_data$run)){
for(k in rev(my_models)){
  my_features <- xgb_res[[k]]$xgb_results$feature_names
  shap_plot[[k]] <- shap.plot.summary.wrap1(xgb_res[[k]]$xgb_results,
                                           X = as.matrix(train_features[, my_features]))
}
                                           

shap_plots_top <- ggarrange(plotlist = shap_plot[4:3], labels = names(shap_plot[4:3]),
                        heights = c(3,5),
                        nrow=2, align = "hv", common.legend = TRUE,
                        font.label = list(size = 10, face = "plain", color ="black"),
                        hjust = -0.5)

shap_plots_bottom <- ggarrange(plotlist = shap_plot[2:1], labels = names(shap_plot[2:1]),
                        heights = c(1,1),
                        nrow=2, align = "hv", common.legend = TRUE,
                        font.label = list(size = 10, face = "plain", color ="black"),
                        hjust = -0.5,
                        legend = "none")

shap_plots <- ggarrange(shap_plots_top, shap_plots_bottom,
                        heights = c(1,1.07),
                        ncol=1, align = "hv", common.legend = TRUE)

shap_plots
```


```{r save xgb train results for shap plot, include=FALSE, echo=FALSE, eval=FALSE}
my_number <- 5
save(xgb_res, train_features, file = paste0(
  "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/ML results Tuan/full_high_dose_delta_4fold_cv_2000_runs/",
  "highdose_pfspz_delta_top_4_models_by_AverageRank_xgb_res_for_shap_plots_v", my_number,".RData"))
```

### Confirm differences in CD56+CD16+ and Pf−specific CD3+CD4+ cells between ΔNP and ΔP in 1.8x10^6 PfSPZ group

Note that values are log2 fold-change post-vax over baseline.

```{r confirm differences in CD56+CD16+ and Pf−specific CD3+CD4+ cells, fig.align='center', fig.width=8, fig.height=8, message=FALSE, warning=FALSE}
library(ggpubr)
#load the delta high dose dataset
#local filename: "highdose_PfSPZ_delta_correlation_ML_data_with_missing.rds"
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("1Q-VUAyvGxZcmEV9hgshpuz87QifzpzSd"), path = temp, overwrite = TRUE)
delta_highdose_features <- readRDS(file = dl$local_path) %>%
  rownames_to_column(var="PATID") %>%
  dplyr::select(PATID, contains("CD3"), contains("CD56"), contains("atypical"), contains("Vd1/2"))

#local filename: "KSPZV1 PhenoData TTE 6 months Long Format.rds"
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("1v2l7G7AzXhgcqObzAabYy6_QJsdB5hFn"), path = temp, overwrite = TRUE)
pheno_data <- readRDS(file = dl$local_path) %>%
  filter(Dosegroup == "1.8 x 10^6 PfSPZ") %>%
  dplyr::select(PATID, tte.mal.atp.3, Outcome, Dosegroup) %>%
  distinct(PATID, tte.mal.atp.3, Outcome, Dosegroup)

facs_dat <- pheno_data %>%
  left_join(., delta_highdose_features, by = "PATID") %>%
  drop_na(`FACS_CD56+CD16+_of_live_PBMCs`) 
facs_dat_n <- facs_dat %>% 
  group_by(Outcome) %>%
  mutate(n=n()) %>%
  ungroup() %>%
  mutate(outcome = ifelse(Outcome == "infected", paste0("ΔNP (", n, ")"),
                          paste0("ΔP (", n, ")")))
CD56_plot <- facs_dat_n %>% 
  ggplot(., aes(x = outcome, y = `FACS_CD56+CD16+_of_live_PBMCs`, fill = outcome)) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2, width = 0.2) +
  ylab("CD56+CD16+ (% of live PBMCs)") +
  xlab("outcome (n)") +
  scale_fill_manual(values = c("#A6CEE3", "#1F78B4")) +
  stat_compare_means(method = "wilcox", vjust = 0, hjust = 0.5, label.y.npc = "bottom") +
  theme_bw(base_family = "sans") +
  theme(legend.position = "none",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12))

facs_dat <- pheno_data %>%
  left_join(., delta_highdose_features, by = "PATID") %>%
  drop_na(`FACS_PfSPZ-specific_CD3+CD4+_of_live_PBMCs`) 
facs_dat_n <- facs_dat %>% 
  group_by(Outcome) %>%
  mutate(n=n()) %>%
  ungroup() %>%
  mutate(outcome = ifelse(Outcome == "infected", paste0("ΔNP (", n, ")"),
                          paste0("ΔP (", n, ")")))
CD3CD4_plot <- facs_dat_n %>% 
  ggplot(., aes(x = outcome, y = `FACS_PfSPZ-specific_CD3+CD4+_of_live_PBMCs`, fill = outcome)) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2, width = 0.2) +
  ylab("Pf−specific CD3+CD4+ (% of live PBMCs)") +
  xlab("outcome (n)") +
  scale_fill_manual(values = c("#A6CEE3", "#1F78B4")) +
  stat_compare_means(method = "wilcox", vjust = 0, hjust = 0.5, label.y.npc = "bottom") +
  theme_bw(base_family = "sans") +
  theme(legend.position = "none",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12))

facs_dat <- pheno_data %>%
  left_join(., delta_highdose_features, by = "PATID") %>%
  drop_na(`FACS_atypical_memory_of_B_cells`) 
facs_dat_n <- facs_dat %>% 
  group_by(Outcome) %>%
  mutate(n=n()) %>%
  ungroup() %>%
  mutate(outcome = ifelse(Outcome == "infected", paste0("ΔNP (", n, ")"),
                          paste0("ΔP (", n, ")")))
atBcell_plot <- facs_dat_n %>% 
  ggplot(., aes(x = outcome, y = `FACS_atypical_memory_of_B_cells`, fill = outcome)) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2, width = 0.2) +
  ylab("atypical memory (% of B cells)") +
  xlab("outcome (n)") +
  scale_fill_manual(values = c("#A6CEE3", "#1F78B4")) +
  stat_compare_means(method = "wilcox", vjust = 0, hjust = 0.5, label.y.npc = "bottom") +
  theme_bw(base_family = "sans") +
  theme(legend.position = "none",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12))

facs_dat <- pheno_data %>%
  left_join(., delta_highdose_features, by = "PATID") %>%
  drop_na(`FACS_Vd1/2-_of_T_cells`) 
facs_dat_n <- facs_dat %>% 
  group_by(Outcome) %>%
  mutate(n=n()) %>%
  ungroup() %>%
  mutate(outcome = ifelse(Outcome == "infected", paste0("ΔNP (", n, ")"),
                          paste0("ΔP (", n, ")")))
vd12_plot <- facs_dat_n %>% 
  ggplot(., aes(x = outcome, y = `FACS_Vd1/2-_of_T_cells`, fill = outcome)) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2, width = 0.2) +
  ylab("Vδ1/2− (% of T cells)") +
  xlab("outcome (n)") +
  scale_fill_manual(values = c("#A6CEE3", "#1F78B4")) +
  stat_compare_means(method = "wilcox", vjust = 0, hjust = 0.5, label.y.npc = "bottom") +
  theme_bw(base_family = "sans") +
  theme(legend.position = "none",
        axis.text = element_text(size = 10, colour = "black"),
        axis.title = element_text(size = 12))

my_FACS_plot <- ggarrange(CD56_plot, CD3CD4_plot, atBcell_plot, vd12_plot)

my_FACS_plot
```


```{r save cd11c plot, eval=FALSE, include=FALSE, echo=FALSE}
cairo_pdf(filename = paste0(plotdir, "Figure SX multiple cells FACS NP vs P highdose PfSPZ delta.pdf"), width = 7.5, height = 7.5)
my_FACS_plot
dev.off()
```
