---
title: "KSPZV1 Baseline Multimodal ML revised - plot ROC curves and SHAP plots from CV data"
author: "Tuan M. Tran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document :
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r links discussing approaches, include=FALSE, eval=FALSE}
#Links discussing best approaches:
# https://stats.stackexchange.com/questions/264533/how-should-feature-selection-and-hyperparameter-optimization-be-ordered-in-the-m
```

# Objective

Performing training and cross-validation on full dataset

# Approach

1. Plot ROC curves and SHAP plots from xgb data derived from "KPSZV1 Baseline Multimodal ML Revised TMT xgb nested fs cv on full set.Rmd".


```{r libraries, include=FALSE}
library(xgboost)
library(dplyr)
library(tidyverse)
library(pROC)
library(caret)
library(doMC)
library(data.table)
library(xgboost)
library(mlr)
library(tictoc)
library(Biobase)
library(parallel)
library(parallelMap)
parallelStartSocket(cpus = detectCores())
```



```{r set local paths}
datadir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/kspzv1-systems-analysis/"
plotdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Figures JCI resubmission/"
```

```{r read in full dataset}
complete_baseline_eset <- read_rds(paste0(datadir, "KSPZV1 logCPM expression sets for visualization/PfSPZ_cpm_ExpressionSet_244x21815_AllGroups_bothBatches_0_rmBatchFX_06082021_TMT_logTRUE.rds"))

high_dose_features <- read_rds(paste0(datadir,"Leetah ML code/kspzv1_hidose_imputed_ml_datasetfull_042123.RDS"))
high_dose_baseline_eset <- complete_baseline_eset[,which(complete_baseline_eset$SAMPLEID %in% colnames(high_dose_features))]
```

```{r make highdose features}
##before splitting, we want to join the classification variable and additional variables of interest to the features data
high_dose_features_class <- t(high_dose_features) %>%
  data.frame(check.names=FALSE) %>%
  rownames_to_column(var = "SAMPLEID") %>%
  left_join(., pData(high_dose_baseline_eset) %>%
              mutate(class = factor(ifelse(mal.atp.3 == 1, "infected", "protected"))) %>% #convert to class factor with reference as "infected"
              dplyr::select(SAMPLEID, site, SEX, age.vax1, mal.vax.1, class) %>%
              dplyr::rename(sex = "SEX",
                            age = "age.vax1",
                            pfbaseline = mal.vax.1),
            by = "SAMPLEID") %>%
  dplyr::select(class, sex, site, pfbaseline, age, everything()) %>%
  column_to_rownames(var = "SAMPLEID") %>%
  dplyr::select(-c(sex, site))

if(all(rownames(high_dose_features_class) == colnames(high_dose_baseline_eset))){
  print("good to go!")
} else {
    print("please check to see if colnames match.")
}
train_features <- high_dose_features_class
```

### Plot ROC curve

```{r plot ROC curve from cv results for best iteration, fig.align='center', fig.width=6, fig.height=3.75}

#readin data saved from above
datadir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/ML results Tuan/"
 
all_folds_2500_runs <- load(paste0(datadir,
                                  "full_high_dose_4fold_cv_2500_runs/",
                                  "all_folds_data_seed_1943_runs_2500_19_Aug_2023_full.RData"))
per_run_res_2500_runs <- readxl::read_xlsx(paste0(datadir,
                                  "full_high_dose_4fold_cv_2500_runs/",
                                  "cv_results_all_runs_all_folds_seed_1943_runs_2500_htune_iters_500_19_Aug_2023_full.xlsx"))
summarized_res_2500_runs <- readxl::read_xlsx(paste0(datadir,
                                  "full_high_dose_4fold_cv_2500_runs/",
                                  "cv_summarized_res_all_folds_seed_1943_runs_2500_htune_iters_500_19_Aug_2023_full.xlsx"))
```

### Make ROC curvers for top n combinations

```{r roc curves for top n combinations}
smooth <- FALSE
n_top_features <- 3
best_run <- best_combo <- all_folds_response <- all_folds_predictor <- roc_dat <- auc <- c()
for(j in 1:n_top_features){
  best_run[[j]] <- summarized_res_2500_runs %>%
    arrange(desc(mean_cv_AUC)) %>%
    slice_max(mean_cv_AUC, n=j) %>%
    slice_min(mean_cv_AUC, n=1) %>%
    pull(var = run)
  best_combo[[j]] <- per_run_res_2500_runs[per_run_res_2500_runs$run==best_run[[j]],] %>%
    group_by(run) %>%
    slice_max(AUC) %>%
    distinct(Features) %>%
    pull(var = Features)
  #calculate predictive accuracy across all 4 folds
  all_folds_response[[j]] <- c(fold1[[best_run[[j]]]]$predictor_data$class,
                               fold2[[best_run[[j]]]]$predictor_data$class,
                               fold3[[best_run[[j]]]]$predictor_data$class,
                               fold4[[best_run[[j]]]]$predictor_data$class)
  all_folds_predictor[[j]] <- c(fold1[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold2[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold3[[best_run[[j]]]]$predictor_data$pred.prob,
                            fold4[[best_run[[j]]]]$predictor_data$pred.prob)
  roc_dat[[j]] <- pROC::roc(response = all_folds_response[[j]],
                            predictor = all_folds_predictor[[j]],
                            smooth = smooth) 
  auc[[j]] <- roc_dat[[j]]$auc %>% as.numeric(.) %>% signif(.,digits=3)
  }

data.labels <- data.frame("name" = as.character(1:n_top_features),
                          "run" = unlist(best_run),
                          "features" = unlist(best_combo),
                          "auc" = unlist(auc))

roc_plots <- ggroc(roc_dat, legacy.axes = TRUE, size = 1.5) +
  theme_bw() +
  scale_color_brewer("", palette = 7, type = "qual") +
  theme(legend.position = "right") +
  geom_text(data = data.labels[1,], aes(0.875, 0.2, label = paste("AUC =", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[2,], aes(0.875, 0.15, label = paste("AUC =", auc)), hjust=1, show_guide=FALSE) +
  geom_text(data = data.labels[3,], aes(0.875, 0.1, label = paste("AUC =", auc)), hjust=1, show_guide=FALSE) +
  theme(plot.margin = unit(c(0.5,2,0.5,0.5), "cm")) +
  geom_abline(slope = 1, linetype = "dotted", color = "gray")

roc_plots
```

### Run xgb on full data set using the downselected, cross-validated features

```{r xgb downselected features on full dataset}
source("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/kspzv1-systems-analysis/kspzv1 xgboost manual train and cv function.R")
xgb_res <- c()
set.seed(12345)
for(k in unlist(best_run)){
  my_features <- fold4[[k]]$downselected_xgb_results$feature_names
  xgb_res[[k]] <- xgb_train_on_selected_features(train_features = train_features,
                                                 features_to_test = my_features,
                                                 maxiterations = 500)
}
```

```{r plot shapley plots, fig.align='center', fig.width=8 fig.height=8}
#see vignettes in https://liuyanguu.github.io/post/2019/07/18/visualization-of-shap-for-xgboost/
library(SHAPforxgboost)
library(ggpubr)


#note that 0=infected (not protected) and 1 = protected
shap_plot <- c()
for(k in unlist(best_run)){
  my_features <- xgb_res[[k]]$xgb_results$feature_names
  shap_plot[[k]] <- shap.plot.summary.wrap1(xgb_res[[k]]$xgb_results,
                                           X = as.matrix(train_features[, my_features]))
}
                                           

shap_plots <- ggarrange(plotlist = shap_plot, nrow=3, align = "hv", common.legend = TRUE)

shap_plots
```

```{r save high dose xgb res and shap plots}
saveRDS(xgb_res, file = paste0(plotdir, "high_dose_pfspz_full_xgb_res_for_shap_cv_2500_runs_26.rds"))
ggsave(filename = paste0(plotdir, "high_dose_pfspz_full_ds_shap_plots_2500_runs_26.pdf"), plot=shap_plots, height=5.5, width=5.85, device = "pdf")
```

```{r save high dose auroc curves}
ggsave(filename = paste0(plotdir, "high_dose_pfspz_full_cv_auroc_2500_runs.pdf"), plot=roc_plots, height=3.5, width=5.75, device = "pdf")
```

# Apply 1.8x10^6 PfSPZ full model to placebo group

```{r readin placebo group}
datadir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/kspzv1-systems-analysis/"
placebo_features <- read_rds(paste0(datadir,"Leetah ML code/kspzv1_placebo_imputed_ml_datasetfull_042123.RDS"))
placebo_baseline_eset <- complete_baseline_eset[,which(complete_baseline_eset$SAMPLEID %in% colnames(placebo_features))]
```

```{r make placebo features}
##before splitting, we want to join the classification variable and additional variables of interest to the features data
placebo_features_class <- t(placebo_features) %>%
  data.frame(check.names=FALSE) %>%
  rownames_to_column(var = "SAMPLEID") %>%
  left_join(., pData(placebo_baseline_eset) %>%
              mutate(class = factor(ifelse(mal.atp.3 == 1, "infected", "protected"))) %>% #convert to class factor with reference as "infected"
              dplyr::select(SAMPLEID, site, SEX, age.vax1, mal.vax.1, class) %>%
              dplyr::rename(sex = "SEX",
                            age = "age.vax1",
                            pfbaseline = mal.vax.1),
            by = "SAMPLEID") %>%
  dplyr::select(class, sex, site, pfbaseline, age, everything()) %>%
  column_to_rownames(var = "SAMPLEID") %>%
  dplyr::select(-c(sex,site))

if(all(rownames(placebo_features_class) == colnames(placebo_baseline_eset))){
  print("placebo set is good to go!")
} else {
    print("please check to see if colnames match.")
}

set.seed(1234) #set seed for reproducibility
test_eset <- placebo_baseline_eset
test_features <- placebo_features_class

test_dat <- t(test_features) %>%
  data.frame(check.names = FALSE) %>%
  t() %>%
  data.frame(check.names = FALSE) %>%
  mutate(class = factor(class)) %>%
  mutate_at(c(2:ncol(.)), as.numeric)

setDT(test_dat, keep.rownames = TRUE)
test_dat_samplenames <- test_dat$rn
test_dat <- test_dat[,-1]

if(all(colnames(test_dat) == colnames(test_features)) &
   all(test_dat_samplenames == rownames(test_features))){
  print("placebo (test) set good to go!")
  } else {
    print("please check to see if test samples and features match.")
    }
#check missing values
#table(is.na(test_dat))
#sapply(test_dat, function(x) sum(is.na(x))/length(x))*100
#see https://www.hackerearth.com/practice/machine-learning/machine-learning-algorithms/beginners-tutorial-on-xgboost-parameter-tuning-r/tutorial/ for         more data cleaning options

#assign data and labels using one hot encoding 
test_labels <- test_dat$class #Levels: infected protected --> 1 2
new_test <- model.matrix(~.+0, data = test_dat[,-c("class"), with=F])
colnames(new_test) <- gsub("\\`","",colnames(new_test))
#convert factor to numeric 
test_labels <- as.numeric(test_labels)-1 #note that 0=infected (not protected) and 1 = protected using one hot encoding 
#prepare matrix
dtest <- xgb.DMatrix(data = new_test, label = test_labels)

temp_df <- data.frame("class" = placebo_features_class$class, "labels" = test_labels)
table(temp_df$class, temp_df$label) #confirmed that infected = 0, and protected = 1
#here is tricky part. now we want to reverse the classes given that the "protected" group in placebo is hypothesized to resemble the "infected" or "not protected" group in the high-dose PfSPZ infants at baseline. (Remember that "infected/not protected" was originally labeled "0"i n the high-dose group)
temp_df <- temp_df %>%
  mutate(dummy_label = ifelse(labels==0, 1, 0))
placebo_dummy_label <- temp_df$dummy_label
```

## Predict placebo group using high-dose PfSPZ model

Determine if the best high-dose PfSPZ xgb model, as determined by highest AUC or lowest test mean error, can predict the opposite class in the placebo dataset

```{r confusion matrix and importance}
my_features <- xgb_res[[k]]$xgb_results$feature_names
xgbDMat_test_dat1 <- xgb.DMatrix(data = new_test[,my_features],
                                 label = placebo_dummy_label,
                                 nthread = 8)
xgbpred1 <- predict(xgb_res[[k]]$xgb_results, xgbDMat_test_dat1)
xgbpred1 <- ifelse(xgbpred1 > 0.5,1,0)
xgbpred1 <- factor(ifelse(xgbpred1==1, "protected", "not protected")) #remember that we are going with the original high-dose designation: infected=0, protected=1
test_labels_for_confusion_matrix <- factor(ifelse(test_labels==1, "protected", "not protected"))
confusion_mat_best_run_error <- caret::confusionMatrix(xgbpred1, test_labels_for_confusion_matrix)
```

```{r save test features and results on downselected}
resdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/ML results Tuan/"
saveRDS(confusion_mat_best_run_error, file = paste0(resdir,
                                                    "full_high_dose_test_placebo/",
                                                    "confusion_mat_high_dose_best_run_error_model_test_on_placebo",
                                                    gsub("\\:","-", format(Sys.time(),"%a-%b-%d-%X-%Y")),".rds"))
```

## Train downselected features on placebo data

```{r xgb downselected features on full placebo dataset}
source("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/kspzv1-systems-analysis/kspzv1 xgboost manual train and cv function.R")
xgb_placebo_res <- c()
for(k in unlist(best_run)){
  my_features <- fold4[[k]]$downselected_xgb_results$feature_names
  xgb_placebo_res[[k]] <- xgb_train_on_selected_features(train_features = test_features,
                                                 features_to_test = my_features,
                                                 maxiterations = 500)
}
```

## Plot SHAP for placebo

```{r plot shapley plots placebo, fig.align='center', fig.width=6, fig.height=5}
#see vignettes in https://liuyanguu.github.io/post/2019/07/18/visualization-of-shap-for-xgboost/
library(SHAPforxgboost)
library(ggpubr)


#note that 0=infected (not protected) and 1 = protected
shap_placebo_plot <- c()
for(k in unlist(best_run)){
  my_features <- xgb_placebo_res[[k]]$xgb_results$feature_names
  shap_placebo_plot[[k]] <- shap.plot.summary.wrap1(xgb_placebo_res[[k]]$xgb_results,
                                           X = as.matrix(train_features[, my_features]))
}
                                           
shap_placebo_plots <- ggarrange(plotlist = shap_placebo_plot, nrow=3, align = "hv", common.legend = TRUE)

shap_placebo_plots
```

```{r save placebo and shap plots}
ggsave(filename = paste0(plotdir, "placebo_full_ds_shap_plots_2500_runs.pdf"), plot=shap_placebo_plots, height=5.5, width=5.85, device = "pdf")
```