---
title: "KSPZV1 WGCNA baseline"
author: "Tuan M. Tran"
date: "04/24/2023"
output: html_document
---

### Objective

Perform weighted gene correlation network analysis as originally described by Horvath et al. on pre-immunization baseline transcriptomic data from the KSPZV1 clinical trial.

References:

https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/
https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/JMiller/Tutorial%20document.pdf

Reviewer requested that we provide scatter plots for the correlations between modules and phenotypes.

### Load packages

```{r load packages, message=FALSE, warning=FALSE}
library(edgeR)
library(readxl)
library(EDASeq)
library(Biobase)
library(WGCNA)
library(ape)
library(Cairo)
library(CorLevelPlot)
library(tidyverse)
library(igraph)
library(remotes)
library(fgsea)
library(data.table)
library(ggplot2)
library(viridis)
library(ggpubr)
library(googledrive)
allowWGCNAThreads()
library(broom)
```

### Options and define variables

```{r options and define variables, message=FALSE, warning=FALSE}
myCor = "bicor"
power <- 12.5 #determined by evaluating previous plot using pickSoftThreshold
myMergeCutHeight <- 0.05
myDeepSplit <- 2
minModSize <- 20
enforceMMS <- FALSE
cor.pval <- 0.05
```

### Load ExpressionSet

```{r load cpm eset, message=FALSE, warning=FALSE}
#local file: "PfSPZ_cpm_ExpressionSet_244x21815_AllGroups_bothBatches_0_rmBatchFX_06082021_TMT_logTRUE.rds"

temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("1togaBlNIxiDj16FyXTC-r7Qw0A9cG2az"), path = temp, overwrite = TRUE)
x <- readRDS(file = dl$local_path)
dim(x)
```

### Make weighted gene correlation matrix based on full data set

```{r weighted gene correlation matrix based on full data set, message=FALSE, warning=FALSE}
WGCNA_matrix <- t(exprs(x)) #make correlations only with full eset
blockSize(ncol(WGCNA_matrix), rectangularBlocks = TRUE, maxMemoryAllocation = 4^31)
par(mfrow=c(1,1))
plotClusterTreeSamples(datExpr=WGCNA_matrix)
#used bicor for Timepoint 0
powers <- seq(4,20,by=0.5)
sft <- pickSoftThreshold(WGCNA_matrix, powerVector = powers, corFnc = myCor, verbose = 5, networkType ="signed", blockSize = ncol(WGCNA_matrix))
sft$powerEstimate
par(mfrow=c(1,1))
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab='Soft Threshold (power)',ylab='Scale Free Topology Model Fit,signed R²',
     type='n', main = paste('Scale independence'));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=1,col='red'); abline(h=0.90,col='red')
```

### Run an automated network analysis

```{r calculation of adjacency matrix, warning=FALSE, message=FALSE}
#use blockwiseModules given many genes
net <- blockwiseModules(WGCNA_matrix,
                        power=power,
                        deepSplit= myDeepSplit,
                        minModuleSize=minModSize,
                        TOMType="none", 
                        mergeCutHeight=myMergeCutHeight, 
                        TOMDenom="mean",
                        detectCutHeight=0.995,
                        corType=myCor,
                        networkType="signed",
                        pamStage=TRUE,
                        pamRespectsDendro=TRUE,
                        reassignThresh=0.05,
                        verbose=5,
                        saveTOMs=FALSE,
                        maxBlockSize=ncol(WGCNA_matrix), 
                        nThreads = 0)

nModules <- length(table(net$colors))-1
modules <- cbind(colnames(as.matrix(table(net$colors))),table(net$colors))
orderedModules <- cbind(Mnum=paste("M",seq(1:nModules),sep=""),Color=labels2colors(c(1:nModules)))
modules <- modules[match(as.character(orderedModules[,2]),rownames(modules)),]
tmpMEs <- MEs <- net$MEs
colnames(tmpMEs) <- paste("ME",colnames(MEs),sep="")
kMEdat <- signedKME(WGCNA_matrix, tmpMEs, corFnc=myCor) #calculate (signed) eigengene-based connectivity, also known as module membership
WGCNA_dat <- cbind(fData(x)$GeneSymbol, colnames(WGCNA_matrix),net$colors,kMEdat) %>%
  as.data.frame() %>%
  dplyr::rename(GeneSymbol = "fData(x)$GeneSymbol") %>%
  dplyr::rename(ENSEMBLID = "colnames(WGCNA_matrix)") %>%
  dplyr::rename(ModuleColors = "net$colors")
```

### Correlate modules with traits

```{r correlate modules with traits, message=FALSE, warning=FALSE}
# Define numbers of genes and samples
nGenes = ncol(WGCNA_matrix)
nSamples = nrow(WGCNA_matrix)

datvar <- pData(x) %>%
  dplyr::select(PATID, SEQBATCH, site, SEX, age.vax1, mal.vax.1, treat, mal.atp.3, tte.mal.atp.6, mal.dvax, mal.dvax.tot, pfcsp_pre, pfcsp_post, log2FC_CSPAb) %>%
  mutate('Gender (female)' = factor(ifelse(SEX == "F", 1, 0))) %>%
  mutate('Site (Siaya)' = factor(ifelse(site == "Siaya", 1, 0))) %>%
  dplyr::rename('Pf infection at first vax' = "mal.vax.1") %>%
  mutate(pfcsp_pre = ifelse(is.na(pfcsp_pre), median(pfcsp_pre, na.rm=TRUE), pfcsp_pre)) %>%
  mutate(pfcsp_post = ifelse(is.na(pfcsp_post), median(pfcsp_post, na.rm=TRUE), pfcsp_post)) %>%
  mutate('pre-vax anti-CSP IgG' = log10(pfcsp_pre+1)) %>%
  mutate('post-vax anti-CSP IgG' = log10(pfcsp_post+1)) %>%
  mutate('log2FC anti-CSP IgG' = log2((pfcsp_post+1)/(pfcsp_pre+1))) %>%
  mutate('1.8 x 10^6 PfSPZ' = factor(ifelse(treat == "1.8 x 10^6 PfSPZ", 1, 0))) %>%
  mutate('parasitemic events during vax period' = mal.dvax.tot) %>%
  mutate('uninfected, 3 months' = factor(ifelse(mal.atp.3 == 0, 1, 0))) %>%
  dplyr::rename('days to first parasitemia' = "tte.mal.atp.6") %>%
  mutate(Age = age.vax1) %>%
  dplyr::select(PATID, Age, site, SEX, treat,'pre-vax anti-CSP IgG', 'Pf infection at first vax', 'uninfected, 3 months', 'days to first parasitemia', 'log2FC anti-CSP IgG') %>% 
  as_tibble() %>%
  column_to_rownames(var = "PATID")

modTraitCor <- cor(orderMEs(net$MEs), datvar, use = "pairwise.complete.obs", method = "pearson")
modTraitP <- corPvalueStudent(modTraitCor, nSamples)
#Since we have a moderately large number of modules and traits, a suitable graphical representation will help in reading the table. We color code each association by the correlation value: Will display correlations and their p-values
#Select out only modules that have P<0.05 in Protection

modTraitP.temp <- modTraitP %>%
  as.data.frame() %>%
  rownames_to_column(var = "Module") %>%
  filter(.$'uninfected, 3 months' < cor.pval | .$'days to first parasitemia' < cor.pval)
modTraitCor.select <- modTraitCor[modTraitP.temp$Module,]
modTraitP.select <- modTraitP[modTraitP.temp$Module,]
textMatrix <- paste(signif(modTraitCor.select, 2), "\n(P=",
                   signif(modTraitP.select, 1), ")", sep = "")
dim(textMatrix) <- dim(modTraitCor.select)

topmodules <- chooseTopHubInEachModule(WGCNA_matrix, net$colors, omitColors = "grey", power = 12.5, type ="signed")
```

### Prepare data for differential module eigengene analysis

```{r prepare data for differential ME expression, fig.align='center', fig.height=8, fig.width=8}
diffME_dat <- orderMEs(net$MEs) %>% 
  rownames_to_column(var = "subject") %>%
  mutate(subject = gsub("\\_0", "", subject)) %>%
  right_join(., datvar %>%
               dplyr::select(-c(Age, "days to first parasitemia")) %>%
               mutate("Pf infection at first vax" = factor(.$"Pf infection at first vax", levels = c(0,1),
                                                           labels = c("Pf-","Pf+"))) %>%
               mutate("uninfected, 3 months" = factor(.$"uninfected, 3 months", levels = c(0,1),
                                                      labels = c("infected","never infected"))) %>%
               mutate("pre-vax anti-CSP IgG" = factor(ifelse(.$"pre-vax anti-CSP IgG" > 0, "present", "absent"),
                                                      levels = c("absent","present"))) %>%
               mutate("log2FC anti-CSP IgG" = factor(ifelse(.$"log2FC anti-CSP IgG" > 12.5, "high responder", "low responder"),
                                                      levels = c("low responder","high responder"))) %>%
               dplyr::rename("fold-change anti-CSP IgG" = "log2FC anti-CSP IgG") %>%
               rownames_to_column(var = "subject"),
             by = "subject") %>%
  pivot_longer(cols = contains("ME"), names_to = "module", values_to = "MEs") %>%
  pivot_longer(cols = c("pre-vax anti-CSP IgG":"fold-change anti-CSP IgG"), names_to = "trait", values_to = "trait_cat") %>%
  arrange(module, site, treat, trait, trait_cat) %>%
  drop_na(trait_cat)

diffME_dat_samplesize <- diffME_dat %>%
  group_by(module, treat, trait, trait_cat) %>%
  summarize(n=n()) %>% 
  ungroup()
```

### Limma method

```{r WGCNA to limma}
pheno_dat <- diffME_dat %>%
  dplyr::select(subject, site, treat, trait, trait_cat) %>%
  distinct(subject, site, treat, trait, trait_cat) %>%
  pivot_wider(names_from = trait, values_from = trait_cat) %>%
  arrange(subject) %>%
  column_to_rownames(var = "subject") 

feature_dat <- diffME_dat %>%
  dplyr::select(module, MEs) %>%
  distinct(module) %>%
  mutate(module_color = gsub("ME","", module)) %>%
  left_join(., topmodules %>%
  enframe() %>%
  dplyr::rename(EnsemblID = "value",
                module_color = "name") %>%
  left_join(., fData(x) %>%
              dplyr::select(EnsemblID, GeneSymbol), by = "EnsemblID")) %>%
  column_to_rownames(var = "module") %>%
  dplyr::rename(hub_gene = "GeneSymbol")

exprs_dat <- orderMEs(net$MEs) %>% 
  rownames_to_column(var = "subject") %>%
  mutate(subject = gsub("\\_0", "", subject)) %>%
  arrange(subject) %>%
  column_to_rownames(var = "subject") %>%
  dplyr::select(rownames(feature_dat)) %>%
  t() 
  
#check
ifelse(all(colnames(exprs_dat) == rownames(pheno_dat)) &
         all(rownames(exprs_dat) == rownames(feature_dat)),
       "Good to go!",
       "Check for matching names")

#make expression set
wgcna_eset <- ExpressionSet(exprs_dat, phenoData = AnnotatedDataFrame(pheno_dat), featureData = AnnotatedDataFrame(feature_dat))
```

Full dataset

```{r full dataset}
#Pf_baseline
pf_first_vax <- factor(wgcna_eset$`Pf infection at first vax`) #encodes a vector as a factor
treatment <- wgcna_eset$treat
site <- factor(wgcna_eset$site)
design <- model.matrix(~0+pf_first_vax+site+treatment) 
colnames(design)
colnames(design) <- c("uninfected_first_vax", "infected_first_vax","wagai", "lowdose","mediumdose","highdose")
fit <- lmFit(exprs(wgcna_eset), design) # Fit linear models for each gene given a series of arrays.
contrasts <- makeContrasts(infected_first_vax - uninfected_first_vax,
                           levels=design) 
contrasts
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2, trend=TRUE)

Pf_baseline <- topTable(fit2, coef="infected_first_vax - uninfected_first_vax", number = Inf) %>%
  rownames_to_column("module") %>%
  right_join(., feature_dat %>%
               rownames_to_column("module"),
             by = "module") %>%
  mutate(comparison = "infected_vs_uninfected_baseline") %>%
  dplyr::select(comparison, module, module_color, hub_gene, logFC, P.Value, adj.P.Val)
head(Pf_baseline)

#prevax_antiCSPIgG
prevax_antiCSPIgG <- factor(wgcna_eset$`pre-vax anti-CSP IgG`)
design <- model.matrix(~0+prevax_antiCSPIgG+site+treatment)
colnames(design)
colnames(design) <- c("noCSPAb_first_vax", "CSPAb_first_vax","wagai","lowdose","mediumdose","highdose")
fit <- lmFit(exprs(wgcna_eset), design) # Fit linear models for each gene given a series of arrays.
contrasts <- makeContrasts(CSPAb_first_vax - noCSPAb_first_vax,
                           levels=design) 
contrasts
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2, trend=TRUE)

CSPAb_baseline <- topTable(fit2, coef="CSPAb_first_vax - noCSPAb_first_vax", number = Inf) %>%
  rownames_to_column("module") %>%
  right_join(., feature_dat %>%
               rownames_to_column("module"),
             by = "module") %>%
  mutate(comparison = "present_vs_absent_baseline") %>%
  dplyr::select(comparison, module, module_color, hub_gene, logFC, P.Value, adj.P.Val)
head(CSPAb_baseline)

#protection
outcome_3mos <- factor(wgcna_eset$`uninfected, 3 months`)
design <- model.matrix(~0+outcome_3mos+site+treatment)
colnames(design)
colnames(design) <- c("not_protected","protected", "wagai", "lowdose","mediumdose","highdose")
fit <- lmFit(exprs(wgcna_eset), design) # Fit linear models for each gene given a series of arrays.
contrasts <- makeContrasts(protected - not_protected,
                           levels=design) 
contrasts
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2, trend=TRUE)

p_vs_np_3mos <- topTable(fit2, coef="protected - not_protected", number = Inf) %>%
  rownames_to_column("module") %>%
  right_join(., feature_dat %>%
               rownames_to_column("module"),
             by = "module") %>%
  mutate(comparison = "protected_vs_not_protected") %>%
  dplyr::select(comparison, module, module_color, hub_gene, logFC, P.Value, adj.P.Val)
head(p_vs_np_3mos)

#CSP Ab response 
FC_CSPAb <- factor(wgcna_eset$`fold-change anti-CSP IgG`)
design <- model.matrix(~0+FC_CSPAb+site+treatment) 
colnames(design)
colnames(design) <- c("low_responder","high_responder","wagai","lowdose","mediumdose","highdose")
fit <- lmFit(exprs_dat, design) # Fit linear models for each gene given a series of arrays.
contrasts <- makeContrasts(high_responder - low_responder,
                           levels=design) 
contrasts
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2, trend=TRUE)
hi_vs_low_CSP <- topTable(fit2, coef="high_responder - low_responder", number = Inf) %>%
  rownames_to_column("module") %>%
  right_join(., feature_dat %>%
               rownames_to_column("module"),
             by = "module") %>%
  mutate(comparison = "hi_vs_low_CSP") %>%
  dplyr::select(comparison, module, module_color, hub_gene, logFC, P.Value, adj.P.Val)
head(hi_vs_low_CSP)

all_module_trait_dat <- rbind(Pf_baseline, CSPAb_baseline, hi_vs_low_CSP, p_vs_np_3mos) %>%
  arrange(P.Value)

p_vs_np_3mos_select <- p_vs_np_3mos %>%
  filter(P.Value<0.05)

all_module_trait_hm_dat <- all_module_trait_dat %>%
  filter(hub_gene %in% p_vs_np_3mos_select$hub_gene) %>%
  dplyr::select(comparison, module_color, hub_gene, logFC) %>%
  pivot_wider(names_from = comparison, values_from = logFC) %>%
  column_to_rownames(var = "hub_gene")
```

### Heatamp for WGCNA Limma data

```{r heatmap wgcna limma, fig.align='center', fig.height=9, fig.width=9}
all_module_trait_hm_annot_dat <- all_module_trait_hm_dat %>%
  rownames_to_column(var= "hub_gene") %>%
  select(hub_gene)

all_module_trait_hm_mat <- all_module_trait_hm_dat %>%
  select(-module_color) %>%
  as.matrix()

row_ha = rowAnnotation(module = all_module_trait_hm_dat$module_color)

clab <- list(
  "hub_gene" = c("RIOK3" = "skyblue1", "SEC62" = "thistle3", CSDE1 = "mediumpurple1", EFHD2 = "lavenderblush2")
  )

clab <- rev(clab)
row_ha <- rowAnnotation(df = all_module_trait_hm_annot_dat, col = clab, na_col = "lavender", annotation_name_gp = gpar(fontsize = 7),
                        simple_anno_size = unit(0.2, "inches"))


paletteLength <- 50
# length(breaks) == length(paletteLength) + 1
# use floor and ceiling to deal with even/odd length pallettelengths
myBreaks <- c(seq(-1*max(abs(all_module_trait_hm_mat)), 0, length.out=ceiling(paletteLength/2) + 1), 
              seq(max(all_module_trait_hm_mat)/paletteLength, max(all_module_trait_hm_mat), length.out=floor(paletteLength/2)))
myColor <- colorRampPalette(c("dodgerblue", "white", "#b22222"))(length(myBreaks))

cairo_pdf(file = "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Figures JCI resubmission/Figure 2A Revised to WGCNA limma.pdf", width = 5, height = 5)
Heatmap(all_module_trait_hm_mat,
        col = circlize::colorRamp2(myBreaks, myColor),
        color_space = "LAB",
        row_names_side = "left",
        row_dend_side = "right",
        left_annotation = row_ha,
        border = TRUE)
dev.off()
```

### Identify hub genes from protection analysis

```{r id hub genes 2, message=FALSE, warning=FALSE}
myColors <- gsub("ME", "", p_vs_np_3mos_select$module_color)
mytopmodules <- topmodules[myColors] %>%
  as.data.frame() %>%
  dplyr::rename(EnsemblID = ".") %>%
  rownames_to_column("module_label") %>%
  as_tibble() %>%
  left_join(., fData(x) %>%
              dplyr::select(EnsemblID, GeneSymbol), by = "EnsemblID")
devtools::source_url("https://github.com/jtlovell/limmaDE2/blob/master/R/wgcna2igraph.R?raw=TRUE")
graph <- wgcna2igraph(net = net, WGCNA_matrix, modules2plot = myColors,
                      colors2plot =  myColors,
                      kME.threshold = 0.5, adjacency.threshold = 0.1,
                      adj.power = power, verbose = T,
                      node.size = 1.5, frame.color = NA, node.color = scales::muted("red"),
                      edge.alpha = .7, edge.width = 0.5)
hubscores <- hub_score(graph, scale = TRUE, weights = NULL,
  options = arpack_defaults)
```

### Plot individual scatter plots for significant module-trait relationships

```{r plot significant scatter plots, fig.align='center', fig.width=8, fig.height=8}
scatplot_dat <- orderMEs(net$MEs) %>% 
  dplyr::select(rownames(modTraitCor.select)) %>%
  rownames_to_column(var = "subject") %>%
  mutate(subject = gsub("\\_0", "", subject)) %>%
  right_join(., pData(wgcna_eset) %>%
               dplyr::select(-c(site, treat)) %>%
               rownames_to_column(var = "subject"),
             by = "subject") %>%
  pivot_longer(cols = contains("ME"), names_to = "module", values_to = "MEs") %>%
  mutate(module = gsub("MEskyblue1", "RIOK3", module)) %>%
  mutate(module = gsub("MEmediumpurple1", "CSDE1", module)) %>%
  mutate(module = gsub("MEthistle3", "SEC62", module)) %>%
  mutate(module = gsub("MElavenderblush2", "EFHD2", module))
```

```{r scatplot factor data, fig.align='center', fig.width=12, fig.height=5}
library(tidyverse)
library(broom)
scatplot_dat_factor <- scatplot_dat %>%
  pivot_longer(cols = c("Pf infection at first vax":"uninfected, 3 months"), names_to = "trait", values_to = "value")  %>%
  filter(module %in% p_vs_np_3mos_select$hub_gene) %>%
  mutate(trait = factor(trait, levels = c("Pf infection at first vax",
                                          "pre-vax anti-CSP IgG",
                                          "uninfected, 3 months",
                                          "fold-change anti-CSP IgG")))

scatplot_dat_factor_norm_test <- scatplot_dat_factor %>%
  group_by(module, trait, value)%>% 
  do(tidy(shapiro.test(.$MEs))) %>% 
  ungroup() %>%
  select(-method) %>%
  mutate(is_normal = ifelse(p.value<0.05, "no","yes"))
  

scatplot_dat_factor_n <- scatplot_dat_factor %>%
  group_by(module, trait, value) %>%
  summarise(n = n())

data_summary <- function(x) {
  m <- mean(x, na.rm=TRUE)
  sem <-sd(x, na.rm=TRUE)/sqrt(sum(!is.na(x)))
  ymin<-m-1.96*sem
  ymax<-m+1.96*sem
  return(c(y=m,ymin=ymin,ymax=ymax))
}

all_scat_plot <- scatplot_dat_factor %>%
  ggplot(., aes(x=value, y=MEs)) +
  ggbeeswarm::geom_beeswarm(alpha = 0.5, color = "lightblue") +
  stat_summary(fun="mean", geom="segment", mapping=aes(xend=..x.. - 0.1, yend=..y..), linewidth = 0.5)+
  stat_summary(fun="mean", geom="segment", mapping=aes(xend=..x.. + 0.1, yend=..y..), linewidth = 0.5) +
  stat_summary(fun.data=data_summary, geom="errorbar", width=0.1) +
  facet_grid(module~trait, scales = "free") + 
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  ylab("module eigengene") +
  xlab("") +
  theme_bw() +
  theme(axis.text = element_text(colour = "black"),
        strip.background = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        plot.title = element_text(size=8))

signif_hub_genes_only <- all_module_trait_dat %>%
  filter(hub_gene %in% p_vs_np_3mos_select$hub_gene)

ggsave("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Figures JCI resubmission/Fig SX JCI revised WGCNA beeswarm plots with mean 95CIs.pdf", all_scat_plot, device = "pdf", units = "in", height = 4, width = 5.5)
```


1.8x10^6 PfSPZ only

```{r 1.8x10^6 PfSPZ only dataset}
highdose_eset <- wgcna_eset[,wgcna_eset$treat == "1.8 x 10^6 PfSPZ"]

#Pf_baseline
pf_first_vax <- factor(highdose_eset$`Pf infection at first vax`) #encodes a vector as a factor
design <- model.matrix(~0+pf_first_vax) 
colnames(design)
colnames(design) <- c("uninfected","infected")
fit <- lmFit(exprs(highdose_eset), design) # Fit linear models for each gene given a series of arrays.
contrasts <- makeContrasts(infected - uninfected,
                           levels=design) 
contrasts
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2, trend=TRUE)

Pf_baseline_highdose <- topTable(fit2, coef="infected - uninfected", number = Inf) %>%
  rownames_to_column("module") %>%
  right_join(., feature_dat %>%
               rownames_to_column("module"),
             by = "module") %>%
  mutate(comparison = "infected_vs_uninfected_baseline") %>%
  dplyr::select(comparison, module, module_color, hub_gene, logFC, P.Value, adj.P.Val)

#prevax_antiCSPIgG
prevax_antiCSPIgG <- factor(highdose_eset$`pre-vax anti-CSP IgG`)
design <- model.matrix(~0+prevax_antiCSPIgG)
colnames(design)
colnames(design) <- c("absent","present")
fit <- lmFit(exprs(highdose_eset), design) # Fit linear models for each gene given a series of arrays.
contrasts <- makeContrasts(present - absent,
                           levels=design) 
contrasts
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2, trend=TRUE)

CSPAb_baseline_highdose <- topTable(fit2, coef="present - absent", number = Inf) %>%
  rownames_to_column("module") %>%
  right_join(., feature_dat %>%
               rownames_to_column("module"),
             by = "module") %>%
  mutate(comparison = "present_vs_absent_baseline") %>%
  dplyr::select(comparison, module, module_color, hub_gene, logFC, P.Value, adj.P.Val)

#protection
outcome_3mos <- factor(highdose_eset$`uninfected, 3 months`)
design <- model.matrix(~0+outcome_3mos)
colnames(design)
colnames(design) <- c("not_protected","protected")
fit <- lmFit(exprs(highdose_eset), design) # Fit linear models for each gene given a series of arrays.
contrasts <- makeContrasts(protected - not_protected,
                           levels=design) 
contrasts
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2, trend=TRUE)

p_vs_np_3mos_highdose <- topTable(fit2, coef="protected - not_protected", number = Inf) %>%
  rownames_to_column("module") %>%
  right_join(., feature_dat %>%
               rownames_to_column("module"),
             by = "module") %>%
  mutate(comparison = "protected_vs_not_protected") %>%
  dplyr::select(comparison, module, module_color, hub_gene, logFC, P.Value, adj.P.Val)

#CSP Ab response 
FC_CSPAb_highdose <- factor(highdose_eset$`fold-change anti-CSP IgG`)
design <- model.matrix(~0+FC_CSPAb_highdose) 
colnames(design)
colnames(design) <- c("low_responder","high_responder")
fit <- lmFit(exprs(highdose_eset), design) # Fit linear models for each gene given a series of arrays.
contrasts <- makeContrasts(high_responder - low_responder,
                           levels=design) 
contrasts
fit2 <- contrasts.fit(fit, contrasts)
fit2 <- eBayes(fit2, trend=TRUE)
hi_vs_low_CSP_highdose <- topTable(fit2, coef="high_responder - low_responder", number = Inf) %>%
  rownames_to_column("module") %>%
  right_join(., feature_dat %>%
               rownames_to_column("module"),
             by = "module") %>%
  mutate(comparison = "hi_vs_low_CSP") %>%
  dplyr::select(comparison, module, module_color, hub_gene, logFC, P.Value, adj.P.Val)
head(hi_vs_low_CSP_highdose)
```

```{r plot wgcna heatpmap, fig.align='center', fig.height=9, fig.width=4}
library(ComplexHeatmap)
row_ha = rowAnnotation(module = row_annot_dat$module_color)
col_fun = circlize::colorRamp2(c(-max(all_module_trait_hm_mat), 0, max(all_module_trait_hm_mat)), c("blue", "white", "red"))
Heatmap(all_module_trait_hm_mat, col = col_fun)
```

### Display the correlation values within a heatmap plot (Figure 2A of pre-print)

```{r plot module trait correlations heatmap, warning=FALSE, message=FALSE, fig.align='center', fig.width=8, fig.height=8}
par(mar = c(11, 9, 1, 1))
labeledHeatmap(Matrix = modTraitCor.select, xLabels = names(datvar),
               yLabels = rownames(modTraitCor.select), ySymbols = rownames(modTraitCor.select),
               colorLabels =FALSE,colors=blueWhiteRed(100),textMatrix=textMatrix,
               setStdMargins = FALSE, zlim = c(-1,1),
               main = paste("Module-trait relationships"),xLabelsAngle = 45) 
```


```{r scatplot continuous data}

scatplot_dat_uninfected_continuous <- scatplot_dat %>%
  pivot_longer(cols = c("uninfected, 3 months"), names_to = "trait", values_to = "value")  %>%
  ggplot(., aes(x=value, y=MEs)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~module, scales = "free", nrow = 2) +
  stat_cor(method="pearson")  +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +  ylab("module eigengene") +
  xlab("uninfected, 3 months") +
  theme_bw() +
  theme(axis.text = element_text(colour = "black"),
        strip.background = element_blank())
  
  
scatplot_dat_continuous <- scatplot_dat %>%
  pivot_longer(cols = c("pre-vax anti-CSP IgG", "days to first parasitemia", "log2FC anti-CSP IgG"), names_to = "trait", values_to = "value")  %>%
  filter(trait == "pre-vax anti-CSP IgG" |
         trait == "days to first parasitemia" & module %in% c("SEC62", "EFHD2") |
         trait == "log2FC anti-CSP IgG" & module %in% c("RIOK3", "SEC62", "EFHD2")) 

library(ggpubr)
log2FC_CSP_plot <- scatplot_dat_continuous %>%
  filter(trait ==  "log2FC anti-CSP IgG") %>%
  ggplot(., aes(x=value, y=MEs)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~module, scales = "free", nrow = 2) +
  stat_cor(method="pearson")  +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +  ylab("module eigengene") +
  xlab("log2(fold-change anti-CSP IgG)") +
  theme_bw() +
  theme(axis.text = element_text(colour = "black"),
        strip.background = element_blank())

days2parasitemia_plot <- scatplot_dat_continuous %>%
  filter(trait ==  "days to first parasitemia") %>%
  ggplot(., aes(x=value, y=MEs)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~module, scales = "free", nrow = 2) +
  stat_cor(method="pearson")  +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +  ylab("module eigengene") +
  xlab("days to first parasitemia") +
  theme_bw() +
  theme(axis.text = element_text(colour = "black"),
        strip.background = element_blank())

prevaxCSPAb_plot <- scatplot_dat_continuous %>%
  filter(trait ==  "pre-vax anti-CSP IgG") %>%
  ggplot(., aes(x=value, y=MEs)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~module, scales = "free", nrow = 2) +
  stat_cor(method="pearson")  +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  ylab("module eigengene") +
  xlab("pre-vax anti-CSP IgG") +
  theme_bw() +
  theme(axis.text = element_text(colour = "black"),
        strip.background = element_blank())

FigSX_WCGNA_continuous_corr <- ggarrange(prevaxCSPAb_plot, days2parasitemia_plot, log2FC_CSP_plot, widths = c(2,1.2), labels = "AUTO")

ggsave("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Figures JCI resubmission/Fig SX JCI revised WGCNA continuous corrs.pdf", FigSX_WCGNA_continuous_corr, device = "pdf", units = "in", height = 11, width = 9)
```

### Identify hub genes 2


```{r id hub genes 2, message=FALSE, warning=FALSE}
myColors <- gsub("ME", "", rownames(modTraitCor.select))
topmodules <- chooseTopHubInEachModule(WGCNA_matrix, net$colors, omitColors = "grey", power = 12.5, type ="signed")
mytopmodules <- topmodules[myColors] %>%
  as.data.frame() %>%
  dplyr::rename(EnsemblID = ".") %>%
  rownames_to_column("module_label") %>%
  as_tibble() %>%
  left_join(., fData(x) %>%
              dplyr::select(EnsemblID, GeneSymbol), by = "EnsemblID")
devtools::source_url("https://github.com/jtlovell/limmaDE2/blob/master/R/wgcna2igraph.R?raw=TRUE")
graph <- wgcna2igraph(net = net, WGCNA_matrix, modules2plot = myColors,
                      colors2plot =  myColors,
                      kME.threshold = 0.5, adjacency.threshold = 0.1,
                      adj.power = power, verbose = T,
                      node.size = 1.5, frame.color = NA, node.color = scales::muted("red"),
                      edge.alpha = .7, edge.width = 0.5)
hubscores <- hub_score(graph, scale = TRUE, weights = NULL,
  options = arpack_defaults)
```

### Top modules and hub genes (rownames for Figure 2A of pre-print)

```{r display my topmodules, echo=FALSE}
knitr::kable(mytopmodules)
```


### Plot network graph of significant modules (Figure 2B of pre-print)

Network graphs of significant modules containing nodes (red dots) and edges (lines) meeting minimum thresholds. Correlations between nodes in different modules are shown as black edges.

```{r plot networkd graph, message=FALSE, warning=FALSE, fig.align='center', fig.width=8, fig.height=8, echo=FALSE}
plot(graph)
```

### Identify modules with < 0.05

```{r id modules that are significant than pull out gene list, message=FALSE, warning=FALSE}
WGCNA_dat_select <- c()
for(i in modTraitP.temp$Module){
  module.color <- sub("ME","", i)
  module.colname <- paste0("kME", i)
  WGCNA_dat_select[[i]] <- WGCNA_dat %>% 
    filter(ModuleColors == module.color) %>%
    dplyr::select(GeneSymbol, all_of(module.colname))
}
lapply(WGCNA_dat_select, nrow)
```
### Output for Ingenuity Pathways Analysis

```{r output modules to csv for IPA}
write_csv(WGCNA_dat_select$MEskyblue1,
          "/Users/tuantran/OneDrive - Indiana University/FromGoogle/Manuscripts/KSPZV1 Manuscript/JCI Submission November 2022/MEskyblue1.csv")

write_csv(WGCNA_dat_select$MEmediumpurple1,
          "/Users/tuantran/OneDrive - Indiana University/FromGoogle/Manuscripts/KSPZV1 Manuscript/JCI Submission November 2022/MEmediumpurple1.csv")

write_csv(WGCNA_dat_select$MEthistle3,
          "/Users/tuantran/OneDrive - Indiana University/FromGoogle/Manuscripts/KSPZV1 Manuscript/JCI Submission November 2022/MEthistle3.csv")

write_csv(WGCNA_dat_select$MElavenderblush2,
          "/Users/tuantran/OneDrive - Indiana University/FromGoogle/Manuscripts/KSPZV1 Manuscript/JCI Submission November 2022/MElavenderblush2.csv")
```

### Overrepesenation Analysis

```{r WGCNA ora}
myuniverse <- unique(fData(x)[fData(x)$EnsemblID %in% colnames(WGCNA_matrix),]$GeneSymbol)
ORA_baseline_bound_df <- c()
minSize <- 15

#Combine similar modules skyblue1, mediumpurple1, thistle3 in to one
WGCNA_dat_select2 <- list()
WGCNA_dat_select2$EFHD2 <- WGCNA_dat_select$MElavenderblush2
WGCNA_dat_select2$RIOK3_CSDE1_SEC62 <- data.table::rbindlist(list(WGCNA_dat_select$MEskyblue1,
                                                WGCNA_dat_select$MEmediumpurple1,
                                                WGCNA_dat_select$MEthistle3)) %>%
  as.data.frame() %>%
  rename(kMEMEall = "kMEMEskyblue1")
for(k in names(WGCNA_dat_select2)){
  WGCNA_dat_select2[[k]] <- WGCNA_dat_select2[[k]][order(WGCNA_dat_select2[[k]][,2], decreasing = TRUE),]
  ORA_baseline_bound_df[[k]] <- ApplyORA2Genesets(genelist = WGCNA_dat_select2[[k]]$GeneSymbol,
                                                  geneset = "all",
                                                  universe = myuniverse,
                                                  output_directory = tempdir(),
                                                  filename_prefix = paste0("ORA_Mod_Corr_Protect_3_mos_", k,
                                                                           "_minSize", minSize),
                                                  minSize = minSize)
  }


RIOK3_ORA_baseline_bound <- bind_rows(ORA_baseline_bound_df$MEskyblue1, .id = "module_type") %>%
  mutate(module = "skyblue1") %>%
  mutate(module_size = nrow(WGCNA_dat_select$MEskyblue1)) %>%
  mutate(hub_gene = "RIOK3") 
CSDE1_ORA_baseline_bound <- bind_rows(ORA_baseline_bound_df$MEmediumpurple1, .id = "module_type") %>%
  mutate(module = "mediumpurple1") %>%
  mutate(module_size = nrow(WGCNA_dat_select$MEmediumpurple1)) %>%
  mutate(hub_gene = "CSDE1")
SEC62_ORA_baseline_bound <- bind_rows(ORA_baseline_bound_df$MEthistle3, .id = "module_type") %>%
  mutate(module = "thistle3") %>%
  mutate(module_size = nrow(WGCNA_dat_select$MEthistle3)) %>%
  mutate(hub_gene = "SEC62")
EFHD2_ORA_baseline_bound <- bind_rows(ORA_baseline_bound_df$MElavenderblush2, .id = "module_type") %>%
  mutate(module = "lavenderblush2") %>%
  mutate(module_size = nrow(WGCNA_dat_select$MElavenderblush2)) %>%
  mutate(hub_gene = "EFHD2")

EFHD2_ORA_baseline_bound <- bind_rows(ORA_baseline_bound_df$EFHD2, .id = "module_type") %>%
  mutate(module = "lavenderblush2") %>%
  mutate(module_size = nrow(WGCNA_dat_select$MElavenderblush2)) %>%
  mutate(hub_gene = "EFHD2")

RIOK3_CSDE1_SEC62_ORA_baseline_bound <- bind_rows(ORA_baseline_bound_df$RIOK3_CSDE1_SEC62, .id = "module_type") %>%
  mutate(module = "the_three_amigos") %>%
  mutate(module_size = nrow(WGCNA_dat_select$MElavenderblush2)) %>%
  mutate(hub_gene = "RIOK3_CSDE1_SEC62")


ORA_baseline_allbound <- bind_rows(RIOK3_ORA_baseline_bound,
                                  CSDE1_ORA_baseline_bound,
                                  SEC62_ORA_baseline_bound,
                                  EFHD2_ORA_baseline_bound,
                                  RIOK3_CSDE1_SEC62_ORA_baseline_bound) %>%
  mutate(neglogpadj = -log10(padj)) %>%
  mutate(pct_overlap = 100*(overlap/module_size)) %>%
  dplyr::select(module, module_size, hub_gene, module_type, pathway, overlap,
                pct_overlap, size, overlapGenes, pval, padj, neglogpadj)
```

### Plot WGCNA Over-represenation Results

baseline analysis

```{r plotdat wgcna ORA, warning=FALSE, message=FALSE}
myModuleTypes <- c("MSigDB_Hallmark_v7.4", "MSigDB_C2_kegg_v7.4", "highBTMs", "BloodGen3Module")
myORAClusterPlotDat <- ORA_baseline_allbound %>%
  mutate(pathway = gsub("VS", "v", pathway)) %>%
  mutate(pathway = gsub("Vd", "Vδ", pathway)) %>%
  mutate(pathway = gsub("gd", "γδ", pathway)) %>%
  mutate(pathway = sub(".*?\\_", "", pathway)) %>%
  mutate(pathway = fct_reorder(pathway, neglogpadj))  %>%
  arrange(desc(neglogpadj))%>%
  mutate(TextLabelColor = ifelse(module_type == "BloodGen3Module", scales::muted("red"),
                                 ifelse(module_type == "MSigDB_C2_kegg_v7.4", scales::muted("blue"),
                                        ifelse(module_type == "MSigDB_Hallmark_v7.4", "black","gray")))) %>%
  filter(padj < 0.01) %>%
  filter(pct_overlap >= 5) %>%
  filter(module_type %in% myModuleTypes) %>%
  group_by(hub_gene) %>%
  arrange(neglogpadj) %>%
  slice_tail(n = 4) %>%
  ungroup()
```

### Arrange plots

```{r make ggarranged plot, warning=FALSE, message=FALSE}
addSmallLegend <- function(myPlot, pointSize = 1.5, textSize = 3, spaceLegend = 0.3) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

scale_begin <- floor(min(myORAClusterPlotDat$pct_overlap, na.rm = TRUE))
scale_end <- ceiling(max(myORAClusterPlotDat$pct_overlap, na.rm = TRUE))

myArrangedPlot <- myORAClusterPlotDat %>%
  filter(hub_gene == "EFHD2" | hub_gene == "RIOK3_CSDE1_SEC62") %>%
  mutate(pathway = fct_reorder(pathway, neglogpadj))  %>%
  ggplot(., aes(x = neglogpadj, y = pathway, fill = pct_overlap)) +
  geom_bar(stat = 'identity') +
  geom_vline(xintercept = 2, color = "red", linetype = "dotted") +
  scale_fill_distiller(direction = 1, breaks = c(0, 10, 20, 30, 40, 50), limits = c(0,50)) +
  ylab("module") +
  theme_classic(base_family = "sans", base_size = 14) +
  theme(legend.position = "bottom",
        plot.margin = unit(c(0,0.5,0,0.5), "cm")) +
  facet_wrap(~hub_gene, scales ="free", nrow = 2) +
  theme(strip.background = element_blank())
```

### Baseline WGCNA Top Modules ORA bar plots (Figure 2C)
 
```{r print my arrange plot, fig.align='center', fig.width=6, fig.height=6, warning=FALSE, message=FALSE, echo=FALSE}
print(myArrangedPlot)
```


```{r output plot, echo=FALSE, eval=FALSE}
cairo_pdf(filename = "/Users/tuantran/OneDrive - Indiana University/FromGoogle/Manuscripts/KSPZV1 Manuscript/JCI Submission November 2022/Figures for JCI submission/Figure 2C Revised to ORA.pdf", width = 5, height = 4)
addSmallLegend(myArrangedPlot)
dev.off()
```



```{r plotdat wgcna gsea, warning=FALSE, message=FALSE, echo=FALSE, eval=FALSE}
#GSEA bar plots (Figure 2C of pre-print)

#GSEA of genes within modules that significant correlated with outcome using blood transcription modules, BloodGen3Modules, and KEGG pathways #as gene sets (BH-adjusted p<0.10). The SEC62-hubbed module did not show significant enrichment for any gene set.

#Plot WGCNA GSEA Results
#baseline analysis
addSmallLegend <- function(myPlot, pointSize = 1.5, textSize = 3, spaceLegend = 0.3) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}
myModuleTypes <- c("MSigDB_Hallmark_v7.4", "MSigDB_C2_kegg_v7.4", "highBTMs", "lowBTMs", "BloodGen3Module")
myGSEAClusterPlotDat <- GSEA_baseline_bound_df$MElavenderblush2 %>%
  mutate(module_hub = "EFHD2") %>%
  bind_rows(., GSEA_baseline_bound_df$MEthistle3 %>% mutate(module_hub = "SEC62") ) %>%
    bind_rows(., GSEA_baseline_bound_df$MEmediumpurple1  %>%
  mutate(module_hub = "CSDE1")) %>%
  bind_rows(., GSEA_baseline_bound_df$MEskyblue1  %>%
  mutate(module_hub = "RIOK3")) %>%
  mutate(neglogpadj = -log10(padj)) %>%
  mutate(pathway = gsub("VS", "v", pathway)) %>%
  mutate(pathway = gsub("Vd", "Vδ", pathway)) %>%
  mutate(pathway = gsub("gd", "γδ", pathway)) %>%
  mutate(pathway = sub(".*?\\_", "", pathway)) %>%
  mutate(pathway = fct_reorder(pathway, NES))  %>%
  arrange(desc(NES))%>%
  mutate(TextLabelColor = ifelse(module_type == "BloodGen3Module", scales::muted("red"),
                                 ifelse(module_type == "MSigDB_C2_kegg_v7.4", scales::muted("blue"),
                                        ifelse(module_type == "MSigDB_Hallmark_v7.4", "black","gray")))) %>%
  filter(padj < 0.10) %>%
  filter(module_type %in% myModuleTypes)
```

```{r FGSEA on WGCNA Results, warning=FALSE, message=FALSE, echo=FALSE, eval=FALSE}
#FGSEA on WGCNA Results
#rank by module membership

minSize <- 5
# Make rank list; ranked by ModuleMembership correlation
  ranks <- GSEA_baseline_bound_df <- c()
devtools::source_url("https://github.com/TranLab/ModuleLists/blob/main/NamedGeneRankList2GseaTable.R?raw=TRUE")
for(k in names(WGCNA_dat_select)){
  WGCNA_dat_select[[k]] <- WGCNA_dat_select[[k]][order(WGCNA_dat_select[[k]][,2], decreasing = TRUE),]
  ranks[[k]] <- WGCNA_dat_select[[k]][,2]
  names(ranks[[k]]) <- WGCNA_dat_select[[k]]$GeneSymbol
  GSEA_baseline_bound_df[[k]] <- NamedGeneRankList2GseaTable(rankedgenes = ranks[[k]], geneset = "all", output_directory = tempdir(),
                                                        filename_prefix = paste0("FGSEA_Mod_Corr_Protect_3_mos_", k,
                                                                                 "_minSize", minSize), scoreType = "pos", minSize = minSize, fixed_seed = TRUE)
  }
```


