---
title: "KSPZ1 Phagocytosis OR Forest Plots and Activation Markers"
author: "Tuan M. Tran"
date: "10/12/2023"
output: html_document
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(googledrive)
```

### Odds ratios of phagocytosis of various pre-treatments relative to controls (media only, isotype, mock treatment, etc.)

```{r load the data, eval=FALSE, include=FALSE, echo=FALSE}
#Notes: The tabulation of number of experiments (n_experiments) has a bug. It incorrectly calculates the number of experiments as 1 for treatment types.

myPath <- "/Users/tuantran/Library/CloudStorage/GoogleDrive-tuantran@iu.edu/Shared drives/[Sec] IN-MED-INF-TRANLAB/Tran Lab Shared/Projects/Doris Duke PfSPZ Kenya/Tuan PfSPZ/KenyaPfSPZ/Final Data Files/"
## read in the most recent count data
# temp <- tempfile(fileext = ".rds")
# dl <- drive_download(
#   as_id("1wdDk3JMQhxQTANKH2TtBMkKZXHJGogFz"), path = temp, overwrite = TRUE)
# 
# countdat <- readRDS(dl$local_path)
countdat <- readRDS(paste0(myPath,"PfSPZ phagocytosis Count dataframe Aditi 07122022.rds"))
```

```{r data processing, eval=FALSE, include=FALSE, echo=FALSE}
#exclude control and setup experiments
exclude <- c("Mono+cryo PfSPZ_PfCSP2A10_Unst Mono+Isotype-PfSpz", "Mono+cryo PfSPZ_PfCSP2A10_Unst Mono+L9LS-PfSpz","Mono+cryo PfSPZ_PfCSP2A10_LPS st_Mono+Isotype-Pfspz","Mono+cryo PfSPZ_PfCSP2A10_LPS st_Mono+L9LS-Pfspz","Mono+cryo PfSPZ_PfCSP2A10_LPS st_Mono+Pfspz","Mono+cryo PfSPZ_PfCSP2A10_Flagellin st_Mono+Pfspz")

propdat <- countdat %>%
  filter(Exp_type %in% c("Mono+cryo Pfspz_PfCSP2A10","Mono + Cryo Pfspz","THP-1 + Cryo PfSPZ")) %>%
  mutate(Exp_type = gsub("Pfspz","PfSPZ",Exp_type)) %>%
  mutate(Exp_type_group = paste0(Exp_type,"_",group)) %>%
  filter(!Exp_type_group %in% exclude) %>%
  mutate(Exp_type = ifelse(Exp_type == "Mono+cryo PfSPZ_PfCSP2A10","Mono + Cryo PfSPZ", Exp_type)) %>%
  mutate(Treatment =ifelse(Treatment == "Unstimulated", "vehicle", Treatment)) %>%
  mutate(Treatment =ifelse(group == "Unst Mono+PfSpz", "vehicle", Treatment)) %>%
  mutate(Treatment = ifelse(group == "Unst Mono+L9LS-PfSpz","unst + L9LS", Treatment)) %>%
  mutate(Treatment = ifelse(group == "Unst Mono+Isotype-PySpz","unst + Isotype", Treatment)) %>%
  mutate(Treatment = ifelse(group == "Unst Mono+Isotype-PfSpz","unst + Isotype", Treatment)) %>%
  mutate(Treatment = ifelse(group == "LPS st_Mono+Pfspz","LPS",Treatment)) %>%
  mutate(Treatment = ifelse(group =="LPS st_Mono+Isotype-Pfspz","LPS + Isotype",Treatment )) %>%
  mutate(Treatment = ifelse(group == "LPS st_Mono+L9LS-Pfspz" , "LPS + L9LS", Treatment)) %>%
 mutate(Treatment = gsub("Flagellin", "flagellin", Treatment)) %>%
  mutate(Treatment = gsub("Poly IC", "poly I:C", Treatment)) %>%
  mutate(Treatment = gsub("IMIQ", "imiquimod", Treatment)) %>%
  mutate(Treatment = gsub("B-glucan", "beta-glucan", Treatment)) %>%
  filter(Gate %in% c("Hoechst+Py+Total", "Hoechst+Py+Internal", "Hoechst+Py+External", "Hoechst+","Hoechst+Py-","Hoechst+Py-+External"))%>% 
  mutate(Count = as.numeric(Count)) %>%
   mutate(Treatment = ifelse(Exp_type_group == "Mono + Cryo PfSPZ_Unst Mono+PfSpz","vehicle6",Treatment)) %>%
   mutate(Treatment = ifelse(Exp_type_group == "Mono+cryo PfSPZ_PfCSP2A10_Unst Mono+PfSpz","vehicle4",Treatment)) %>%
  filter(Gate %in% c("Hoechst+Py+Total", "Hoechst+Py+Internal", "Hoechst+Py+External","Hoechst+","Hoechst+Py-","Hoechst+Py-+External"))  %>%
  mutate(Count = as.numeric(Count)) %>%
 #  dplyr::select(-c(group)) %>%
  group_by(Exp_type, Exp_no, Treatment, Gate) %>%
  summarise(Count = sum(Count)) %>%
  ungroup() %>%
  droplevels()
```

```{r fetdat,print forest plot, eval=FALSE, include=FALSE, echo=FALSE}
fetdat <- df <- temp <-  tempdat <- mydat <- mydat_bound <- fetres <- Exp_annot <- contigency_tab <- c()
#propdat$Exp_no <- as.numeric(propdat$Exp_no)
for(i in c("THP-1 + Cryo PfSPZ","Mono + Cryo PfSPZ")){
      tempdat[[i]] <- propdat[propdat$Exp_type == i,] %>%
      droplevels
  for(j in unique(tempdat[[i]]$Treatment)){
      mydat[[i]][[j]] <- propdat[propdat$Exp_type == i & propdat$Treatment == j,] %>%
      mutate(Experiments = paste(unique(Exp_no), collapse = ",")) %>%
      mutate(n_Experiments = length(unique(Exp_no)))
       mydat_bound[[i]] <- bind_rows(mydat[[i]]) %>%
      ungroup() %>%
      droplevels()
  }}


for(i in c("THP-1 + Cryo PfSPZ","Mono + Cryo PfSPZ")){
for(j in c("unst + L9LS","LPS","LPS + L9LS", "flagellin","beta-glucan", "poly I:C", "imiquimod","CpG")){
      
      df[[i]] <- mydat_bound[[i]] %>%
        as.data.frame() 
      
## following if-else sequences sets references         
        if (i == "Mono + Cryo PfSPZ" & j == "unst + L9LS")
                temp[[i]][[j]] <- df[[i]] %>%
                   as.data.frame() %>%
                    filter(`Treatment` %in% c("unst + Isotype",j)) %>%
                   group_by(Exp_type, Exp_no, Treatment, Gate, Experiments, n_Experiments) %>%
                  summarise(Count = sum(Count)) %>%
                   ungroup() %>%
                    #arrange(desc(Treatment))%>%
                  droplevels()
        
 ### LPS and Flagellin compairing with vehicle from 6th experiment       
        else if (i == "Mono + Cryo PfSPZ" & j %in% c("LPS","flagellin"))
           temp[[i]][[j]] <- df[[i]] %>%
              as.data.frame() %>%
              filter (Treatment %in% c( "vehicle6",j)) %>%
                   group_by(Exp_type, Exp_no, Treatment, Gate, Experiments, n_Experiments) %>%
                  summarise(Count = sum(Count)) %>%
                   ungroup() %>%
                   #arrange(desc(Treatment))%>%
                  droplevels()
        
  ### TLRs compairing with vehicle from 4th experiment         
        else if (i == "Mono + Cryo PfSPZ" & j %in% c("beta-glucan", "poly I:C", "imiquimod","CpG"))
           temp[[i]][[j]] <- df[[i]] %>%
              as.data.frame() %>%
              filter (Treatment %in% c( "vehicle4",j)) %>%
                   group_by(Exp_type, Exp_no, Treatment, Gate, Experiments, n_Experiments) %>%
                  summarise(Count = sum(Count)) %>%
                   ungroup() %>%
                   #arrange(desc(Treatment))%>%
                  droplevels()
        

         else if (i == "Mono + Cryo PfSPZ" & j == "LPS + L9LS")
           temp[[i]][[j]] <- df[[i]] %>%
              as.data.frame() %>%
              filter (Treatment %in% c( "LPS + Isotype",j)) %>%
                   group_by(Exp_type, Exp_no, Treatment, Gate, Experiments, n_Experiments) %>%
                  summarise(Count = sum(Count)) %>%
                   ungroup() %>%
                   #arrange(desc(Treatment))%>%
                  droplevels()

          else (
            temp[[i]][[j]] <- df[[i]] %>%
              as.data.frame() %>%
              filter(`Treatment` %in% c(j,"vehicle")) %>%
              group_by(Exp_type, Exp_no, Treatment, Gate, Experiments, n_Experiments)%>%
              summarise(Count = sum(Count)) %>%
              arrange(desc(Treatment))%>%
               ungroup() %>%
              droplevels())
    
    Exp_annot[[i]][[j]]  <- bind_rows(temp[[i]][[j]]) %>%
    dplyr::select(Treatment, Experiments, n_Experiments) %>%
    filter(Treatment == j) %>%
    distinct(Treatment, .keep_all = TRUE)
    
### following if statements removes unnecessary comparisons.
    
    if(i %in% c("THP-1 + Cryo PfSPZ") & j %in% c("unst + L9LS","LPS + L9LS","beta-glucan", "poly I:C", "imiquimod","CpG"))
    {
       temp [[i]] [[j]] = NULL
       Exp_annot[[i]][[j]] = NULL
    }   
  
    
}}
### fliping internal vs external cols-- a way to re-level vehicle and TLR agonists during the calculations

for (i in c("THP-1 + Cryo PfSPZ")){
  for (j in c("LPS", "flagellin")) {
   # contigency_tab[[i]][[j]] <- xtabs(Count ~ Treatment+Gate, data=temp[[i]][[j]])[,c("Hoechst+Py-+External","Hoechst+Py+Internal")]
    fetres[[i]][[j]] <- fisher.test(xtabs(Count ~ Treatment+Gate, data=temp[[i]][[j]])[,c("Hoechst+Py+Internal","Hoechst+Py-+External")])
fetdat[[i]][[j]] <- data.frame("OR" = fetres[[i]][[j]]$estimate, "LCI" = fetres[[i]][[j]]$conf.int[1], "UCI" = fetres[[i]][[j]]$conf.int[2], "p" = fetres[[i]][[j]]$p.value)
  }}

for (i in c("Mono + Cryo PfSPZ")){
  for (j in c("unst + L9LS","LPS + L9LS")) {
    #contigency_tab[[i]][[j]] <- xtabs(Count ~ Treatment+Gate, data=temp[[i]][[j]])[,c("Hoechst+Py-+External","Hoechst+Py+Internal")]
    fetres[[i]][[j]] <- fisher.test(xtabs(Count ~ Treatment+Gate, data=temp[[i]][[j]])[,c("Hoechst+Py-+External","Hoechst+Py+Internal")])
fetdat[[i]][[j]] <- data.frame("OR" = fetres[[i]][[j]]$estimate, "LCI" = fetres[[i]][[j]]$conf.int[1], "UCI" = fetres[[i]][[j]]$conf.int[2], "p" = fetres[[i]][[j]]$p.value)
  }}


for (i in c("Mono + Cryo PfSPZ")){
  for (j in c("beta-glucan","LPS", "poly I:C", "flagellin", "imiquimod","CpG")) {
    #contigency_tab[[i]][[j]] <- xtabs(Count ~ Treatment+Gate, data=temp[[i]][[j]])[,c("Hoechst+Py-+External","Hoechst+Py+Internal")]
    fetres[[i]][[j]] <- fisher.test(xtabs(Count ~ Treatment+Gate, data=temp[[i]][[j]])[,c("Hoechst+Py+Internal","Hoechst+Py-+External")])
fetdat[[i]][[j]] <- data.frame("OR" = fetres[[i]][[j]]$estimate, "LCI" = fetres[[i]][[j]]$conf.int[1], "UCI" = fetres[[i]][[j]]$conf.int[2], "p" = fetres[[i]][[j]]$p.value)
  }}

all_annots <- unlist(Exp_annot, recursive = FALSE) %>%
          do.call("rbind",.) %>%
          rownames_to_column(var = "Experiment.Treatment")%>%
          mutate(Experiment_type = gsub("\\..*","",Experiment.Treatment))

allfets <- unlist(fetdat, recursive = FALSE) %>%
                do.call("rbind",.) %>% 
                rownames_to_column(var = "Experiment.Treatment") %>%
                as_tibble() %>%
                mutate(BH.adj.p = p.adjust(.$p, n = nrow(.))) %>%
                right_join(., all_annots, by = "Experiment.Treatment") %>%
                mutate(Treatment = gsub("beta", "β", Treatment)) %>%
                mutate(Treatment = gsub("unst","vehicle", Treatment)) %>%
                mutate(label = Treatment) %>%
                mutate(label = ifelse(n_Experiments > 1,paste0(label," (",n_Experiments,")"),label))%>%
                mutate(plot_type = ifelse(Treatment == "vehicle + L9LS" ,"Mono + Cryo PfSPZ L9LS only", Experiment_type)) %>%
                mutate(plot_type = ifelse(Treatment == "LPS + L9LS" ,"Mono + Cryo PfSPZ L9LS only", plot_type)) %>%
                mutate(plot_type = ifelse(Treatment == "vehicle + RAM1","Mono + Cryo PySPZ RAM only",plot_type)) %>%
                 mutate(plot_type = ifelse(Treatment == "LPS + RAM1","Mono + Cryo PySPZ RAM only",plot_type)) %>%
                dplyr::select(-c(Experiment.Treatment)) %>%
                dplyr::select(c(9,6,1:5,7,8,10,11)) %>%
                as.data.frame()

allfets$plot_type <- factor(allfets$plot_type, levels= c("Mono + Cryo PfSPZ L9LS only","Mono + Cryo PfSPZ","THP-1 + Cryo PfSPZ"))


allfets$Treatment <- factor(allfets$Treatment, levels = c("vehicle + L9LS","LPS + L9LS","poly I:C","CpG","imiquimod","β-glucan","flagellin","LPS"))

allfets$label <- factor(allfets$label, levels = c("vehicle + L9LS","LPS + L9LS","poly I:C","CpG","imiquimod","β-glucan","flagellin","LPS"))

#cairo_pdf(paste0(myPath,"KSPZV1 phagocytosis analysis PfSPZ & PySPZ Py+Internal vs Py-+External 06292022.pdf"), width = 4, height = 4)
```

```{r load reviewer response data, eval=FALSE, include=FALSE, echo=FALSE}
## Add additional ORs from reviewer requested experiments
#Get data from experiments performed in response to JCI reviewer critiques

tempdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/KSPZV1_JCI revision_JB/Final data from all Mono and Pfspz experiments/"
newdat_exp1 <- readxl::read_excel(paste0(tempdir,"Table6_All cell type_phagocytosis_All exp.xlsx"), sheet = "Exp1") %>%
  mutate(experiment = 1)
newdat_exp2 <- readxl::read_excel(paste0(tempdir,"Table6_All cell type_phagocytosis_All exp.xlsx"), sheet = "Exp2") %>%
  mutate(experiment = 2)
newdat_exp3 <- readxl::read_excel(paste0(tempdir,"Table6_All cell type_phagocytosis_All exp.xlsx"), sheet = "Exp3") %>%
  mutate(experiment = 3)
newdat <- bind_rows(newdat_exp1, newdat_exp2, newdat_exp3)
```

```{r filter summarize and arrange, eval=FALSE, include=FALSE, echo=FALSE}
#Summarize experiments and calculate FET and odds ratios.
spz_pretreat_dat <- newdat %>%
  filter(Treatment %in% c("mock 1°, 2°Pfspz", "1° and 2°Pfspz")) %>%
  dplyr::select(experiment, Treatment,`Cell type`, `Cell count`, `Count(PfCSP2A10pos)`, `Count(PfCSP2A10pos CFSEneg)`) %>%
  mutate(PfSPZ_neg = `Cell count`-`Count(PfCSP2A10pos CFSEneg)`) %>%
  dplyr::rename(PfSPZ_pos = `Count(PfCSP2A10pos CFSEneg)`)  %>%
  dplyr::select(experiment, Treatment,`Cell type`, PfSPZ_neg, PfSPZ_pos) %>%
  dplyr::select(-experiment) %>%
  group_by(Treatment, `Cell type`) %>%
  dplyr::summarise(PfSPZ_neg = sum(PfSPZ_neg), PfSPZ_pos = sum(PfSPZ_pos)) %>%
  filter(`Cell type` == "Total CD14+") %>%
  dplyr::select(-c(`Cell type`)) %>%
  arrange(desc(Treatment)) %>% #make mock reference
  column_to_rownames(var = "Treatment")
spz_pretreat_fetres <- fisher.test(spz_pretreat_dat)
spz_pretreat_res_df <- data.frame("Experiment_type" = "fresh mono + Cryo PfSPZ",
                                  "Treatment" = "PfSPZ",
                                  "OR" = signif(spz_pretreat_fetres$estimate),
                                  "LCI" = spz_pretreat_fetres$conf.int[1],
                                  "UCI" = spz_pretreat_fetres$conf.int[2],
                                  "p" = spz_pretreat_fetres$p.value,
                                  "BH.adj.p" = spz_pretreat_fetres$p.value, #only one comparison
                                  "Experiments" = "3",
                                  "n_Experiments" = 1,
                                  "label" = "PfSPZ*",
                                  "plot_type" = "Mono + Cryo PfSPZ fresh cells")
```

```{r save propdat, include=FALSE, echo=FALSE, eval=FALSE}
writexl::write_xlsx(propdat, "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Data for all Plots JCI Resubmission/Figure 8F Data for Phagocytosis OR forest plots.xlsx")
```

```{r bind to allfests, eval=FALSE, include=FALSE, echo=FALSE}
#Bind new data to allfets
allfets_plus_new <- bind_rows(allfets, spz_pretreat_res_df) %>%
  dplyr::select(plot_type, Treatment, label, OR, LCI, UCI, p, BH.adj.p)
```

```{r save allfets_plus_new, eval=FALSE, include=FALSE, echo=FALSE}
saveRDS(allfets_plus_new, paste0(myPath, "KSPZV1_phagocytosis_allfets_plus_new.rds"))
```

```{r load data local, eval=FALSE, include=FALSE, echo=FALSE}
allfets_plus_new <- read_rds(paste0(myPath, "KSPZV1_phagocytosis_allfets_plus_new.rds"))
```

Download data from googledrive.

```{r load data googledrive, message=FALSE, warning=FALSE}
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("108OYZ9n9c70vWEM2Y9uZ_jvB1iSpgVPR"), path = temp, overwrite = TRUE)
allfets_plus_new <- readRDS(dl$local_path)
```

Make THP-1 OR forest plot from data

```{r plot forest plot thp, echo = TRUE, fig.align='center'}
thp.plot <- allfets_plus_new %>%
  filter(plot_type %in% c("THP-1 + Cryo PfSPZ"))%>%
  mutate(label = fct_relevel(label)) %>%
  mutate(Treatment = fct_relevel(Treatment)) %>%
  ggplot(aes(y=label, x=OR)) +
  geom_point(size=1, shape=19) +
  geom_errorbarh(aes(xmin=LCI, xmax=UCI), height=.3) +
  #coord_fixed(ratio= 0.8)+
  geom_vline(xintercept=1, linetype='longdash') +
  ylab("") +
  xlab("") +
  #ggtitle("THP1 monocytes + purified PfSPZ")+
  theme_bw() +
  theme(plot.title = element_text(size= 11), text = element_text(family = "Helvetica"),
        axis.text.x = element_blank(),
        axis.ticks = element_blank()) +
  theme(plot.margin = margin(0,0,0,0, 'cm')) +
  xlim(0,7)
```

Make TLR agonist and beta-glucan forest plot from data

```{r plot forest plot mono, echo = TRUE, fig.align='center', warning=FALSE, message=FALSE}
alltreat.plot <- allfets_plus_new %>%
  filter(plot_type %in% c("Mono + Cryo PfSPZ"))%>%
  mutate(label = fct_reorder(label, OR, .desc=TRUE)) %>%
  mutate(Treatment = fct_reorder(Treatment, OR, .desc=TRUE)) %>%
  ggplot(aes(y=label, x=OR)) +
  geom_point(size=1, shape=19) +
  geom_errorbarh(aes(xmin=LCI, xmax=UCI), height=.3) +
  #coord_fixed(ratio=0.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  ylab("pre-treatment") +
  xlab("")  +
  #ggtitle("primary monocytes + purified PfSPZ")+
  theme_bw() +
  theme(plot.title = element_text(size= 11), text = element_text(family = "Helvetica"),
        axis.text.x = element_blank(),
        axis.ticks = element_blank()) +
  theme(plot.margin = margin(0,0,0,0, 'cm')) +
  xlim(0,7)
```

Make L9LS control OR forest plot from data

```{r plot forest plot L9LS, echo = TRUE, fig.align='center', warning=FALSE, message=FALSE}
L9LS.plot <- allfets_plus_new %>%
  filter(plot_type %in% c("Mono + Cryo PfSPZ L9LS only"))%>%
  mutate(label = fct_relevel(label)) %>%
  mutate(Treatment = fct_relevel(Treatment)) %>%
  ggplot(aes(y=label, x=OR)) +
  geom_point(size=1, shape=19) +
  geom_errorbarh(aes(xmin=LCI, xmax=UCI), height=.3) +
  #coord_fixed(ratio= 0.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  ylab("") +
  xlab("Odds Ratio") +
 #ggtitle("H   Mono + Cryo PfSPZ L9LS only")+
  theme_bw() +
  theme(plot.title = element_text(size= 11), text = element_text(family = "Helvetica"))+
  theme(plot.margin = margin(0,0,0,0, 'cm')) +
  xlim(0,7)
```

Make PfSPZ pretreatment OR forest plot from data

```{r plot forest plot pfspz pretreat, echo = TRUE, fig.align='center', warning=FALSE, message=FALSE, fig.width=8, fig.height=2}
PfSPZ_pretreat_plot <- allfets_plus_new %>%
  filter(plot_type %in% c("Mono + Cryo PfSPZ fresh cells"))%>%
  mutate(label = fct_relevel(label)) %>%
  mutate(Treatment = fct_relevel(Treatment)) %>%
  ggplot(aes(y=label, x=OR)) +
  geom_point(size=1, shape=19) +
  geom_errorbarh(aes(xmin=LCI, xmax=UCI), height=.3) +
  #coord_fixed(ratio= 0.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  ylab("") +
  xlab("") +
 #ggtitle("H   Mono + Cryo PfSPZ PfSPZ_pretreat_plot")+
  theme_bw() +
  theme(plot.title = element_text(size= 11), text = element_text(family = "Helvetica"),
        axis.text.x = element_blank(),
        axis.ticks = element_blank()) +
  theme(plot.margin = margin(0,0,0,0, 'cm')) +
  xlim(0,7)
```

Use ggarrange to arrange all forest plots into one

```{r gg arrange the plots, echo = TRUE, warning=FALSE,message=FALSE, fig.align='center', fig.width=4, fig.height=5}
fig_7F <- ggarrange(thp.plot,
                   alltreat.plot,
                   PfSPZ_pretreat_plot,
                   L9LS.plot, align = "v",
                   nrow = 4,
                   heights = c(2.2,5,1.6,2.5))

fig_7F
```

```{r save Fig 7F, echo=FALSE, eval=FALSE, include=FALSE}
tempdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Figures JCI resubmission/"
cairo_pdf(paste0(tempdir,"Fig 7F Phagocytosis Odds Ratio Forest Plot with PfSPZ pretreatment JCI revision.pdf"), height = 3.2, width = 4)
fig_7F
dev.off()
```


```{r combine all plot types and save data, echo=FALSE, eval=FALSE, include=FALSE}
all_plot_types <- bind_rows(thp.plot$data, alltreat.plot$data, PfSPZ_pretreat_plot$data, L9LS.plot$data)

all_plot_types %>%
  writexl::write_xlsx("/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Data for all Plots JCI Resubmission/Figure 8F Phagocytosis Odds Ratios.xlsx")
```


### Effect of purified PfSPZ pre-treatment on PfSPZ phagocytosis by freshly isolate human monocytes

```{r read in local data}
tempdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/KSPZV1_JCI revision_JB/Final data from all Mono and Pfspz experiments/"
tab8_exp1 <- readxl::read_excel(paste0(tempdir,
                                       "Experiment 1/Table8_Pheno data selected gp and markers_Exp1.xlsx"),
                                sheet = "Total CD14+")
tab8_exp2 <- readxl::read_excel(paste0(tempdir,
                                       "Experiment 2/Table8_Pheno data_selected gp and markers_Exp2.xlsx"),
                                sheet = "Total CD14+")
tab8_exp3 <- readxl::read_excel(paste0(tempdir,
                                       "Experiment 3/Table8_Pheno data_selected gp and markers_Exp3.xlsx"),
                                sheet = "Total CD14+")
all_facs_tabs <- bind_rows(tab8_exp1, tab8_exp2, tab8_exp3) %>%
  drop_na(Treatment)
```

### Compare media only with single PfSPZ treatment (no secondary PfSPZ)

```{r media only with single PfSPZ, fig.align='center', fig.height=6, fig.width=8}
single_PfSPZ_MFI_plot_exp1 <- all_facs_tabs %>%
  filter(Experiment != "Exp3") %>%
  filter(Markers != "CD86") %>%
  filter(Treatment %in% c("media only", "1° Pfspz, no 2°Pfspz")) %>%
  mutate(Treatment = factor(Treatment, levels = c("media only", "1° Pfspz, no 2°Pfspz"))) %>%
  ggplot(., aes(x = Treatment, y = MFI)) +
  geom_point() +
  stat_summary(fun = mean, color = "red", position = position_dodge(0.75),
               geom = "pointrange",
               shape = 3,
               show.legend = FALSE) +
  stat_compare_means(aes(label = paste0("p = ", after_stat(p.format))),
                     method = "t.test",
                     label.x = 1.25) +
  theme_bw() +
  ggh4x::facet_grid2(c("Experiment", "Markers"),
              labeller = "label_value",
              scales = "free",
              independent = "all") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))

single_PfSPZ_MFI_plot_exp1 
```


```{r pretreat PfSPZ, fig.align='center', fig.height=6, fig.width=8}
pretreat_PfSPZ_MFI_plot_exp1 <- all_facs_tabs %>%
  filter(Experiment != "Exp3") %>%
  filter(Markers != "CD86") %>%
  filter(Treatment %in% c("mock 1°, 2°Pfspz", "1° and 2°Pfspz")) %>%
  mutate(Treatment = factor(Treatment, levels = c("mock 1°, 2°Pfspz", "1° and 2°Pfspz"))) %>%
  ggplot(., aes(x = Treatment, y = MFI)) +
  geom_point() +
  stat_summary(fun = mean, color = "red", position = position_dodge(0.75),
               geom = "pointrange",
               shape = 3,
               show.legend = FALSE) +
  stat_compare_means(aes(label = paste0("p = ", after_stat(p.format))),
                     method = "t.test",
                     label.x = 1.25) +
  theme_bw() +
  ggh4x::facet_grid2(c("Experiment", "Markers"),
              labeller = "label_value",
              scales = "free",
              independent = "all") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))

pretreat_PfSPZ_MFI_plot_exp1 
```