---
title: "KSPZV1 DGE Δ (postvax/baseline) limma voom - JCI revision"
author: "Tuan M. Tran"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document :
    theme: cerulean
---

### Objective

Perform differential gene expression using limma voom to assess differences between protected and not protected infants in the KSPZV1 malaria vaccine trial.
This analysis specifically evaluates differences in post-vax with adjustment for baseline.

This analysis also includes adjustments for:

1. Batch
2. Sex
3. Study site
4. log10(CSP Ab + 1)
5. number of malaria infections during vaccination period

```{r libraries, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(limma)
library(edgeR)
library(googledrive)
library(tidyverse)
library(fgsea)
library(data.table)
library(EDASeq)
library(gtools)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load ExpressionSet

```{r load SeqExpressionSet local, message=FALSE, warning=FALSE, include=FALSE, eval=FALSE, echo=FALSE}
#local path: "/Volumes/GoogleDrive/My Drive/Tran Lab Shared/Projects/Doris Duke PfSPZ Kenya/Tuan PfSPZ/KenyaPfSPZ/Final Data Files/KSPZV1 SeqExpressionSet Baseline Postvax Reduced Phenodat 35716x487.rds"
```

```{r load SeqExpressionSet googledrive, message=FALSE, warning=FALSE}
#from google drive
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("17P7RwDaPCwDXcrD82N4fYd3vGOcueFEH"), path = temp, overwrite = TRUE)
x <- readRDS(file = dl$local_path)
x$treat <- factor(x$treat, levels = c("Placebo", "4.5 x 10^5 PfSPZ", "9.0 x 10^5 PfSPZ", "1.8 x 10^6 PfSPZ"))
```

### Set options

```{r set options, message=FALSE, warning=FALSE}
myGroups <- unique(x$treat)
```

```{r more reduction, echo=FALSE}
#xgroup <- setNames(as.list(c(1:4)), c("Placebo", "4.5 x 10^5 PfSPZ", "9.0 x 10^5 PfSPZ", "1.8 x 10^6 PfSPZ")) #if running for all dose groups
xgroup <- setNames(as.list(c(1:1)), c("1.8 x 10^6 PfSPZ"))
for(i in names(xgroup)){
  xgroup[[i]] <- x[, which(x$treat == i)]
  xgroup[[i]] <- xgroup[[i]][,xgroup[[i]]$Timepoint == 0 | xgroup[[i]]$Timepoint == 25]
  xgroup[[i]]$treat <- droplevels(xgroup[[i]]$treat)
  print(i)
  print(table(Outcome = xgroup[[i]]$Outcome, Timepoint = xgroup[[i]]$Timepoint))
  print(table(Outcome = xgroup[[i]]$Outcome, Dosegroup = xgroup[[i]]$treat))
  print(dim(xgroup[[i]]))
}
```

## Build DGEList Object and make Expression Sets

### Build DGEList Object

```{r make DGEList object, message=FALSE, warning=FALSE}
#Define group by dose, timepoint and outcome #if running for all dose groups
#ygroup <- setNames(as.list(c(1:4)), c("Placebo", "4.5 x 10^5 PfSPZ", "9.0 x 10^5 PfSPZ", "1.8 x 10^6 PfSPZ"))
ygroup <- setNames(as.list(c(1:1)), c("1.8 x 10^6 PfSPZ"))
for(i in names(ygroup)){
  ygroup[[i]]  <- DGEList(counts=counts(xgroup[[i]]), genes=fData(xgroup[[i]]), group= factor(paste(xgroup[[i]]$treat, xgroup[[i]]$Timepoint,
                                                                                                    xgroup[[i]]$Outcome, sep = "_")),
                          remove.zeros=T)
  ygroup[[i]]$samples$PATID <- gsub("_.*", "", rownames(ygroup[[i]]$samples))
  }
```

### Filter out low expression features

```{r low expression filters, message=FALSE, warning=FALSE}
for(i in names(ygroup)){
  keep <- filterByExpr(ygroup[[i]])
  ygroup[[i]] <- ygroup[[i]][keep, , keep.lib.sizes=FALSE]
  ygroup[[i]]$samples$lib.size <- colSums(ygroup[[i]]$counts)
  ygroup[[i]] <- calcNormFactors(ygroup[[i]])	#Normalization
}
```

## Remove unpaired samples and then arrange by patient and timepoint

This is necessary for paired analyses

```{r remove unpaired samples then arrange}
for(i in names(ygroup)){
  ygroup[[i]] <- ygroup[[i]][,!is.na(ygroup[[i]]$samples$group)]
  ygroup[[i]] <- ygroup[[i]][,duplicated(ygroup[[i]]$samples$PATID) | duplicated(ygroup[[i]]$samples$PATID, fromLast = TRUE)]
  ygroup[[i]] <- ygroup[[i]][,order(ygroup[[i]]$samples$PATID, ygroup[[i]]$samples$group)]
  print(i)
  print(ifelse(all(gsub('_.*', '', colnames(ygroup[[i]][,grepl("_0", colnames(ygroup[[i]]))])) ==
                     gsub('_.*', '', colnames(ygroup[[i]][,grepl("_25", colnames(ygroup[[i]]))]))) &
                 all(ygroup[[i]]$samples$Timepoint == rep(c(0,25),length(ygroup[[i]]$samples$Timepoint)/2)), 
               "Samples are paired. All is good.","Stop and check order of samples."))
}
```

## Remake ExpressionSet after sample and feature (gene) filtering steps.

```{r remake SeqExpressionSet after gene filtering and sample reduction, echo = FALSE}
for(i in names(ygroup)){
  myintersection <- intersect(rownames(pData(xgroup[[i]])), colnames(ygroup[[i]]$counts))
  xgroup[[i]] <- newSeqExpressionSet(counts = ygroup[[i]]$counts[,myintersection],
                                     featureData = AnnotatedDataFrame(fData(xgroup[[i]])[which(rownames(fData(xgroup[[i]])) %in% rownames(ygroup[[i]]$counts)),]),
                                     phenoData = AnnotatedDataFrame(pData(xgroup[[i]])[myintersection,]))
}
```

## Design Matrix - Compare Protected vs Not Protected

Analysis between Protected and Not Protected within a Dose Group.
Subject as random effect using duplicate correlation. Batch, sex, study site, CSP-specific IgG pre-immunization, and malaria during vaccination period as fixed effects.
Limma voom can handle this. See following:

https://bioconductor.riken.jp/packages/3.9/bioc/vignettes/variancePartition/inst/doc/dream.html

```{r build design matrix  protected vs susceptible by group}
DeltaDelta <- DeltaNotProtected <- DeltaProtected <- fit <- design <- xgroup
for(i in names(xgroup)){
  Subject <- factor(xgroup[[i]]$PATID)
  Outcome <- factor(xgroup[[i]]$Outcome, levels = c(1,0), labels = c("NotProtected","Protected"))
  Timepoint <- factor(xgroup[[i]]$Timepoint, levels = c(0,25), labels = c("baseline","postvax"))
  Age <- xgroup[[i]]$age.vax1
  Sex <- as.factor(xgroup[[i]]$SEX)
  Batch <- factor(xgroup[[i]]$SEQBATCH, levels = c("Aug2019","Nov2019"))
  Site <- factor(xgroup[[i]]$site)
  MALdVax <- factor(xgroup[[i]]$mal.dvax, levels = c(0,1), labels = c("noMALdVAX","yesMALdVAX"))
  MALdVaxTotal <- as.numeric(ifelse(is.na(xgroup[[i]]$mal.dvax.tot), 0, xgroup[[i]]$mal.dvax.tot))
  CSPAb_baseline <- as.numeric(xgroup[[i]]$pfcsp_pre)
  CSPAb_baseline[is.na(CSPAb_baseline)] <- median(CSPAb_baseline, na.rm = TRUE) #Impute 2 missing values with median of all samples
  #Define protective outcome-specific PfSPZ Vaccination effects and append them to the design matrix
  Protected.Postvax <- Outcome == "Protected" & Timepoint=="postvax"
  NotProtected.Postvax <- Outcome =="NotProtected" & Timepoint=="postvax"
  design[[i]] <- model.matrix(~Batch+Sex+Site+CSPAb_baseline+MALdVaxTotal)
  design[[i]] <- cbind(design[[i]], Protected.Postvax, NotProtected.Postvax) #make design matrix according to edgeR 3.5 (p42)
  #https://www.bioconductor.org/packages/release/bioc/vignettes/edgeR/inst/doc/edgeRUsersGuide.pdf
  rownames(design[[i]]) <- xgroup[[i]]$PATID
  print(paste0("running model fit for ", i))
  print(colnames(design[[i]])) #check colnames

  #Use voom() to convert the read counts to log2-cpm, with associated weights, ready for linear modeling
  v <- voom(ygroup[[i]], design[[i]])
  cor <- duplicateCorrelation(v, design[[i]], block = Subject)
  print(cor$consensus)
  v <- voom(ygroup[[i]], design[[i]], plot = TRUE, block = Subject, correlation = cor$consensus)
  cor <- duplicateCorrelation(v, design[[i]], block = Subject)
  print(cor$consensus)
}
```

```{r fit}
for(i in names(xgroup)){
  fit <- lmFit(v, design[[i]], block = Subject, correlation = cor$consensus)
  fit <- eBayes(fit)

  DeltaProtected[[i]] <- topTable(fit,n=Inf, coef="Protected.Postvax")
  DeltaNotProtected[[i]] <-  topTable(fit,n=Inf, coef="NotProtected.Postvax")
  fit2 <- contrasts.fit(fit, contrasts = c(rep(0,(ncol(design[[i]])-2)),1,-1))
  fit2 <- eBayes(fit2)
  DeltaDelta[[i]] <- topTable(fit2,n=Inf)
}
```

```{r output Table S3 DEGS, echo=FALSE, eval=FALSE, include=FALSE}
resdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Tables JCI resubmission/"
all_delta_degtabs <- list("ΔP" = DeltaProtected$`1.8 x 10^6 PfSPZ`[,c(7,9,2:5,8,10:15)],
                          "ΔNP" = DeltaNotProtected$`1.8 x 10^6 PfSPZ`[,c(7,9,2:5,8,10:15)],
                          "ΔP_v_ΔNP" = DeltaDelta$`1.8 x 10^6 PfSPZ`[,c(7,9,2:5,8,10:15)])

writexl::write_xlsx(all_delta_degtabs, paste0(resdir, "Table S3 Delta Post-vax DEGs Revised Adj Batch Sex Site log CSP Ab baseline MALdVaxTotal.xlsx"))
```

### Examine top 15 DEGs in 1.8x10^6 PfSPZ dose group

Δ Protected 

```{r DEG table placebo, echo = FALSE}
DeltaProtected$`1.8 x 10^6 PfSPZ`[1:15,c(7,9,8,10:15)] %>%
  mutate_if(is.numeric, signif, digits = 3) %>%
  knitr::kable(., row.names = FALSE)
```

Δ Not Protected

```{r DEG table low dose, echo = FALSE}
DeltaNotProtected$`1.8 x 10^6 PfSPZ`[1:15,c(7,9,8,10:15)] %>%
  mutate_if(is.numeric, signif, digits = 3) %>%
  knitr::kable(., row.names = FALSE)
```

Δ Protected vs. Δ Not Protected

```{r DEG table med dose, echo = FALSE}
DeltaDelta$`1.8 x 10^6 PfSPZ`[1:15,c(7,9,8,10:15)] %>%
  mutate_if(is.numeric, signif, digits = 3) %>%
  knitr::kable(., row.names = FALSE)
```

### Apply GSEA

Rank genes by -log10(P.Value)*sign(logFC). Run fgsea from fgsea package using NamedGeneRankList2GseaTable helper function.

```{r apply fgsea, warning=FALSE, message=FALSE}
set.seed(23)
#restructure dataframes
degtabs <- setNames(as.list(c(1:4)), c("Placebo", "4.5 x 10^5 PfSPZ", "9.0 x 10^5 PfSPZ", "1.8 x 10^6 PfSPZ"))
for(i in names(ygroup)){
    degtabs[[i]] <- setNames(as.list(c(1:3)), c("DeltaDelta", "DeltaP", "DeltaNP"))
    degtabs[[i]]$DeltaDelta <- DeltaDelta[[i]]
    degtabs[[i]]$DeltaP <- DeltaProtected[[i]]
    degtabs[[i]]$DeltaNP <- DeltaNotProtected[[i]]
}
ranks <- degtabs
for(i in names(degtabs)){
  for(j in names(degtabs[[i]])){
    ranks[[i]][[j]] <-  degtabs[[i]][[j]] %>%
      mutate(rankmetric = -log10(.$P.Value)*sign(.$logFC)) %>%
      dplyr::select(GeneSymbol,rankmetric) %>%
      na.omit() %>% 
      distinct() %>% 
      group_by(GeneSymbol) %>%
      summarize(rankmetric = mean(rankmetric)) %>%
      arrange(desc(rankmetric))
  }
}
#Run fgsea from fgsea package using NamedGeneRankList2GseaTable helper function
devtools::source_url("https://github.com/TranLab/ModuleLists/blob/main/NamedGeneRankList2GseaTable.R?raw=TRUE")
GSEAtab <- ranks
for(i in names(ranks)){
  for(j in names(ranks[[i]])){
    print(paste0("starting ", j, " in ", i))
    GSEAtab[[i]][[j]] <- NamedGeneRankList2GseaTable(rankedgenes = deframe(ranks[[i]][[j]]),
                                                     geneset = "all",
                                                     output_directory = tempdir(),
                                                     filename_prefix = "GSEA",
                                                     sampleSize = 101,
                                                     minSize = 20,
                                                     maxSize = Inf,
                                                     scoreType = "std") %>%
      as_tibble() %>%
      arrange(desc(NES)) %>% 
      dplyr::select(module_type, pathway, ES, NES, size, leadingEdge, pval, padj) %>% 
      mutate(leadingEdge = gsub("^c\\(|\\)$", "", leadingEdge)) %>%
      mutate(leadingEdge = gsub('"', "", leadingEdge)) %>%
      arrange(padj)
    #closeAllConnections() #this line prevents using up all url connections
    }
}
```

### Visualize GSEA data as bubble plot

Filter based on FDR<5%. 

```{r plot bubble plot, echo = FALSE}
GSEAtab_temp <- GSEAtab
#for(i in names(GSEAtab)){
for(i in c("1.8 x 10^6 PfSPZ")){
  GSEAtab_temp[[i]] <- GSEAtab[[i]] %>%
  bind_rows(., .id = "comparison") %>%
  mutate(comparison = factor(comparison, levels = c("DeltaP", "DeltaNP", "DeltaDelta"), labels = c("ΔP", "ΔNP", "ΔP vs ΔNP")))
}
plotDat <- GSEAtab_temp$`1.8 x 10^6 PfSPZ` %>%
  #bind_rows(., .id = "treatment")  %>%
  #mutate(treatment = factor(treatment, levels = c("Placebo", "4.5 x 10^5 PfSPZ", "9.0 x 10^5 PfSPZ", "1.8 x 10^6 PfSPZ"))) %>%
  filter(padj < 0.05) %>%
  filter(module_type %in% c("highBTMs", "lowBTMs", "MonacoModules", "BloodGen3Module", "MSigDB_Hallmark_v7.4"))  %>%
  filter(!grepl("TBA", pathway)) %>%
  #dplyr::select(treatment, comparison, module_type, pathway, leadingEdge, size, NES, padj) %>%
  dplyr::select(comparison, module_type, pathway, leadingEdge, size, NES, padj) %>%
  mutate(neglogpadj = -log10(padj)) %>%
  mutate(pathway = gsub("gd", "γδ", pathway)) %>%
  mutate(pathway = gsub("Vd2", "Vδ2", pathway)) %>%
  mutate(pathway = gsub("Vg", "Vγ", pathway)) %>%
  mutate(pathway = gsub("HALLMARK_", "", pathway)) %>%
  mutate(pathway = gsub("_", " ", pathway)) %>%
  mutate(pathway = sub(".*?\\_", "", pathway)) %>%
  #group_by(treatment, module_type) %>%
  group_by(module_type) %>%
  mutate(pathway = fct_reorder(pathway, NES, .desc = TRUE)) %>%
  ungroup() %>%
  filter(!grepl("TBD", pathway)) %>%
  mutate(module_type = factor(module_type, levels = c("highBTMs", "MonacoModules", "lowBTMs", "MSigDB_Hallmark_v7.4", "BloodGen3Module"))) %>%
  arrange(desc(neglogpadj)) %>% 
  droplevels()

#plotting options
basetextsize <- 9  
myfont <- "sans"
bubble_max_size <- 7

MyPlot <- plotDat %>%
  filter(module_type %in% c("highBTMs", "MonacoModules"))  %>%
  ggplot(., aes(x = comparison, y = pathway)) +
  geom_point(aes(size=neglogpadj, fill = NES), alpha = 0.65, shape=21, stroke = 0.25) +
  scale_size_area(name = expression(-log[10]~adj.~p~value), max_size = bubble_max_size, breaks = c(1.3, 5, 10, 20, 40), limits = c(1,50)) +
      scale_fill_gradient2(low = "blue",
                           mid = "white",
                           high = "red") +
  hrbrthemes::theme_ipsum_es(base_family = myfont, base_size = basetextsize) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank(),
        legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#facet_wrap(treatment~module_type, scales = "free_y", ncol = 2)
facet_wrap(~module_type, scales = "free_y", ncol = 2)
```

### Plot GSEA bubbleplots (Figure 4D)

Red: goes up post-vax relative to baseline
Blue: goes down post-vax relative to baseline

```{r plot gsea bubbleplot, fig.align='center', fig.width=7, fig.height=4, echo=FALSE, message=FALSE, warning=FALSE}
addSmallLegend <- function(myPlot, pointSize = 4, textSize = 6, spaceLegend = 0.6) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(strip.text = element_text(size=8),
              legend.direction = "vertical",
              legend.position = "right",
              legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"),
              axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, color = "black"),
              axis.text.y = element_text(color = "black"))
}
print(addSmallLegend(MyPlot))
```

```{r output to pdf, echo=FALSE, include=FALSE, eval=FALSE}
figdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Figures JCI resubmission/Figure Panels/"
cairo_pdf(paste0(figdir, "Figure 4D voom version high-dose PfSPZ includes sex adj FDR5.pdf"), width = 6.5, height = 3.75)
addSmallLegend(MyPlot)
dev.off()
```

```{r save figure 4d plot data, include=FALSE, eval=FALSE, echo=FALSE}
plotDat_4D <- plotDat %>%
  filter(module_type %in% c("highBTMs", "MonacoModules"))
writexl::write_xlsx(plotDat_4D, path = "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Data for all Plots JCI Resubmission/Figure 4D Delta high-dose PfSPZ GSEA Bubble Plots HighBTMs and MonacoMods.xlsx")
```

### Gene-level analysis

1. Identify genes that were differentially induced in uninfected and infected children receiving 1.8x10^6 PfSPZ Vaccine

2. Merge ΔP vs ΔNP, ΔP, and ΔNP tables.

3. Criteria:
    ΔP vs ΔNP P val < 0.005
    ΔP logFC > 0 and ΔP P val < 0.05
    ΔNP logFC < 0 and ΔNP P val < 0.10

```{r determine genes of interest, message=FALSE, warning=FALSE}
AllDegs1 <- bind_rows(DeltaDelta, .id = "Treatment") %>%
  mutate(Comparison = "DeltaDelta")
AllDegs2 <- bind_rows(DeltaProtected, .id = "Treatment") %>%
  mutate(Comparison = "DeltaProtected")
AllDegs3 <- bind_rows(DeltaNotProtected, .id = "Treatment") %>%
  mutate(Comparison = "DeltaNotProtected")
AllDegs <- bind_rows(AllDegs1, AllDegs2, AllDegs3) %>%
  dplyr::select(Treatment, Comparison, EnsemblID, GeneSymbol, descripton_new, logFC, AveExpr, "t", P.Value, adj.P.Val) %>%
  dplyr::rename(PValue = P.Value) %>%
  dplyr::rename(FDR = adj.P.Val) %>%
  dplyr::rename(Description = "descripton_new")

gois <- AllDegs %>%
  filter(Treatment == "1.8 x 10^6 PfSPZ") %>%
  droplevels() %>%
  dplyr::select(Treatment, Comparison, EnsemblID, GeneSymbol, Description, logFC, PValue, FDR) %>%
  pivot_wider(., names_from = Comparison, names_glue = "{Comparison}_{.value}", values_from = c(logFC, PValue, FDR)) %>%
  filter((DeltaDelta_PValue < 0.005 & DeltaDelta_logFC > 2) &
           (DeltaProtected_logFC > 0 & DeltaProtected_PValue < 0.05)) %>%
  arrange(DeltaDelta_PValue)

gois_ms <- gois %>%
  dplyr::select(GeneSymbol, DeltaDelta_logFC, DeltaDelta_PValue, DeltaProtected_logFC, DeltaProtected_PValue, DeltaNotProtected_logFC, DeltaNotProtected_PValue)
```

### Figure 4E (table)

```{r table of goi, message=FALSE, warning=FALSE}
gois_ms %>%
  mutate_if(is.numeric, signif,3) %>%
  knitr::kable()
```

```{r output Fig 4E table, echo=FALSE, eval=FALSE, include=FALSE}
plot_dat <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Data for all Plots JCI Resubmission/"
gois_ms %>%
  mutate_if(is.numeric, signif,3) %>%
  writexl::write_xlsx(paste0(plot_dat, "Figure 4E Table Genes Significantly Induced delta P vs delta NP.xlsx"))
```

### Reshape and merge for survival analysis

```{r reshape and merge, echo = TRUE, message=FALSE, warning=FALSE}
library(gtools)
#from google drive
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("17xx0YaggLiyWdJc9rqA1nkPSE7dj05p0"), path = temp, overwrite = TRUE)
cpm_delta <- readRDS(file = dl$local_path)
cpm_delta <- cpm_delta[fData(cpm_delta)$GeneSymbol %in% gois_ms$GeneSymbol, ]
cpm_delta <- cpm_delta[,cpm_delta$treat != "Placebo"] #remove placebo---we want to just look at all PfSPZ doses
pData(cpm_delta) <- droplevels(pData(cpm_delta))
survdat <- Biobase::exprs(cpm_delta) %>%
  as.data.frame() %>%
  rownames_to_column(var = "EnsemblID") %>%
  left_join(., fData(cpm_delta) %>%
              dplyr::select(EnsemblID, GeneSymbol),
            by = "EnsemblID") %>%
  dplyr::select(-c(EnsemblID)) %>%
  dplyr::select(GeneSymbol, everything()) %>%
  pivot_longer(2:ncol(.), names_to = "PATID", values_to = "expression") %>%
  group_by(GeneSymbol) %>%
  mutate(exprs_up_down = factor(ifelse(expression>log2(1), "induced", "not induced"), levels = c("not induced", "induced"))) %>%
  mutate(exprs_terciles = factor(quantcut(expression, q=3, na.rm=TRUE), labels = c("bottom_tercile", "middle_tercile", "top_tercile"))) %>%
  mutate(exprs_median = factor(quantcut(expression, q=2, na.rm=TRUE), labels = c("bottom", "top"))) %>%
  left_join(., fData(cpm_delta) %>%
              dplyr::select(c(GeneSymbol, description)), by = "GeneSymbol") %>%
  left_join(., pData(cpm_delta)%>%
              dplyr::select(c(PATID, Timepoint, treat, site, SEX, pfcsp_pre, age.vax1, mal.vax.1, mal.dvax, mal.dvax.tot, contains("atp"))), by = "PATID") %>%
  droplevels()
survdat$tte.mal.atp.6 <- survdat$tte.mal.atp.6+1
survdat[is.na(survdat$pfcsp_pre),]$pfcsp_pre <- median(survdat$pfcsp_pre, na.rm = TRUE) #Impute 2 missing values with median of all samples
```

### Kaplan-Meier plots

```{r km plots, echo=TRUE, message=FALSE, warning=FALSE}
library(survival)
library(survminer)
myFollowup <- 6 #3 or 6
N <- length(unique(survdat$PATID))
survdat.filtered <- fit <- mySurvPlot <- c()
for(i in gois_ms$GeneSymbol){
  survdat.filtered[[i]] <- survdat %>%
    filter(GeneSymbol == i)
  fit[[i]] <- survfit(Surv(tte.mal.atp.6, mal.atp.6) ~ exprs_up_down, data= survdat.filtered[[i]])
  mySurvPlot[[i]] <- ggsurvplot(fit[[i]], risk.table = TRUE, pval = TRUE,
                                break.time.by = 30, title = unique(survdat.filtered[[i]]$GeneSymbol),
                                xlab = "Days since vaccination",
                                ylab = "% free of parasitemia",
                                censor = TRUE,
                                palette = "lancet",
                                conf.int = TRUE,
                                conf.int.alpha = 0.1,
                                font.family = "Arial")
}
```


```{r print km plots, fig.align='center', fig.height = 5, fig.width = 6, echo = FALSE, eval = FALSE, include=FALSE}
for(i in gois_ms$GeneSymbol){
  plot.new()
  print(mySurvPlot[[i]], newpage = FALSE)
}
```

```{r print invidividual plots, echo=FALSE, eval=FALSE}
# figdir <- "/Volumes/GoogleDrive/My Drive/Manuscripts/KSPZV1 Manuscript/Med Revised Submission 2022/Figures/"
# cairo_pdf(filename = paste0(figdir,"KSPZV1 Fig 4I Delta KM FSTL4 Revision.pdf"), width = 5.5, height = 5.5)
# print(mySurvPlot$FSTL4)
# dev.off()
```

### Kaplan-Meier plot for *FSTL4* (Figure 4F)

```{r km plot 1, echo = TRUE, fig.align='center', fig.height=6, fig.width=6}
print(mySurvPlot$FSTL4, newpage = FALSE)
```

```{r save km plot data, include=FALSE, eval=FALSE, echo=FALSE}
plot_dat <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Data for all Plots JCI Resubmission/"
writexl::write_xlsx(list("FSTL4_KM_plot_dat" = mySurvPlot$FSTL4$data.survplot,
                         "FSTL4_surv_tab_dat" = mySurvPlot$FSTL4$data.survtable),
                    path = paste0(plot_dat, "Figure 4F FSTL4 Survplot Data.xlsx"))
```

### Cox Regression (Table S4)

```{r do cox regression to separate dose groups, echo=TRUE, message=FALSE, warning=FALSE}
cox.res <- c()
for(i in names(survdat.filtered)){
  survdat.cp <- survdat.filtered[[i]]
  survdat.cp$tte.mal.atp.6 <- survdat.cp$tte.mal.atp.6 + 1 #add one day to each value given zero values
  survdat.cp$start		<-		0
  survdat.cp$stop		<-		survdat.cp$tte.mal.atp.6
  subj.pcrp <- rep(survdat.cp$start, as.vector(table(survdat.cp$PATID)))
  survdat.cp$start1 <- survdat.cp$start-subj.pcrp
  survdat.cp$stop1 <- survdat.cp$stop-subj.pcrp
  survdat.cp <- survdat.cp %>%
    droplevels()
  fit <- coxph(Surv(start1, stop1, mal.atp.6) ~ exprs_up_down + factor(SEX) + factor(site) + log10(pfcsp_pre+1) + mal.dvax.tot + factor(treat), data = survdat.cp)
  #vfit <- cox.zph(fit)
  #par(mfrow=c(3,2))
  #plot(vfit)
  #cox.zph(fit)
  cox.res[[i]] <- as.data.frame(cbind(fit$n, fit$nevent, summary(fit)$conf.int[,c(1,3:4)], summary(fit)$coefficient[,5]))
  colnames(cox.res[[i]]) <- c("n","number of events","HR", "LCI", "UCI", "P value")
  cox.res[[i]] <- cox.res[[i]] %>%
    as.matrix() %>%
    signif(., digits = 3) %>%
    as.data.frame()
  cox.res[[i]]$Gene <- i
  rownames(cox.res[[i]]) <- c("upregulated post-vax (ref:downregulated post-vax)",
                              "gender (ref:female)",
                              "Wagai (ref:Siaya)",
                              "CSP-specific IgG baseline",
                              "number of Pf infections during vaccination period",
                              "9.0x10^5 PfSPZ (ref:4.5x10^5 PfSPZ)",
                              "1.8x10^6 PfSPZ (ref:4.5x10^5 PfSPZ)")
  cox.res[[i]] <- cox.res[[i]] %>%
    rownames_to_column(var = "Covariate") %>%
    dplyr::select(Gene, everything())
  }
  cox.res.bound <- bind_rows(cox.res, .id = "Gene") %>%
    mutate(Significant = ifelse(`P value`<0.001, "***",
                              ifelse(`P value`<0.01, "**",
                                     ifelse(`P value`<0.05, "*", ""))))
```

Only display FSTL4 here.

```{r show table, echo=TRUE}
cox.res.bound %>%
  filter(Gene == "FSTL4") %>%
  knitr::kable()
```


```{r write table to file, echo=FALSE, eval=FALSE, include=FALSE}
cox.res.bound %>%
  writexl::write_xlsx(., paste0(resdir, "Table S4 Cox PH for FSTL4 n169 postvax 6 mos adj Batch Sex Site CSPAb MALdVAXTotal revised.xlsx"))
```

