---
title: "PfSPZ correlation chord diagrams post-vax"
author: "Prasida Holla, Aditi Upadhye"
date: "6/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective:
To make linear correlations between the PfSPZ vaccine transcriptomic data and non-trasncriptomic data, the latter of which includes select features from the phenoData, flow cytometric parameters and cytokine levels as assessed by luminex cytokine array. The transcriptomic data from RNA-seq is collapsed into high level blood transcription modules (BTMs) or into Monaco et al. modules. The correlations are visualized as a chord diagram or circos plot. 

Steps:
1. Filter and keep select features from non-transcriptomic data (Phil Swanson's FACS data, post-vaccination CSP-specific IgG antibody levels and cytokine levels)
2. Load the expression set and filter the phenoData for Time point (0 for baseline, 25 for post-vaccination), and time to parasitemia at 6 months. 
3. Merge the selected phenoData features with the filtered non-transcritomic data. We will call this "alldata"
4. Collapse transcriptomic data (eset) into blood transcription modules (BTMs) or Monaco et al. modules
5. Merge the "alldata" with the BTMs or Monaco modules using patient ID
6. Run correlations using ggcorrplot, which should give a table of Pearson's r values and pvalues for each correlation
7. Convert pvalues to FDR and use it to keep only significant correlations
8. Keeping only the top triangle of the correlation matrix, use the package 'Circlize' to visualize the correlations as a chord diagram (aka circos plot). 
9. This script is specifically for baseline features

Notes: 
- Post-vaccination CSP-specific IgG was used for correlations at baseline
- Cytokine levels were only included in the baseline, and not in the correlations of post-vaccination and delta

Per reviewer request, now visualize only correlations with FDR<5%.

## Load required packages

```{r load libraries, include = TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tibble)
library(viridis)
library(ggpubr)
library(readxl)
library(ggcorrplot)
library(Biobase)
library(miscTools)
library(circlize)
library(RColorBrewer)
library(googledrive)
library(Hmisc)
library(tmod)
```

```{r set local directory, include=FALSE, eval=FALSE, echo=FALSE}
datadir <- "/Users/tuantran/Library/CloudStorage/GoogleDrive-tuantran@iu.edu/Shared drives/[Sec] IN-MED-INF-TRANLAB/Tran Lab Shared/Projects/Doris Duke PfSPZ Kenya/Tuan PfSPZ/KenyaPfSPZ/Final Data Files/"
figdir <- "/Users/tuantran/Library/CloudStorage/OneDrive-IndianaUniversity/Manuscripts/KSPZV1 Manuscript/JCI Resubmission April 2023/Figures JCI resubmission/Working Figures/"
```

# Import the expression set and non-transcriptomic data (FACS, cytokine, ICS)

```{r load data,  message=FALSE, warning=FALSE}
##import all non-transcriptomic data google drive
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("1acv8D-gex3ps8o9SCWZT5ueP7XuPeHs3"), path = temp, overwrite = TRUE)
alldata <- readRDS(file = dl$local_path)

#load the post-vax expression dataset
temp <- tempfile(fileext = ".rds")
dl <- drive_download(
  as_id("1w5Wzta8XISNHZ6KL-x9_qO98m3LxaOBa"), path = temp, overwrite = TRUE)
eset <- readRDS(file = dl$local_path)
```

```{r load data local, include=FALSE, eval=FALSE, echo=FALSE}
#local path
#alldata <- readRDS("/Users/tuantran/Library/CloudStorage/GoogleDrive-tuantran@iu.edu/Shared drives/[Sec] IN-MED-INF-TRANLAB/Tran Lab Shared/Projects/Doris Duke PfSPZ Kenya/Tuan PfSPZ/KenyaPfSPZ/Final Data Files/PfSPZ_FACS_PlasmaCytokine_MonocyteActivation_05012021.Rds")
#alldata <- readRDS(paste0(wdir,"Forbaseline_non_transcriptomic_data.rds")) 

#aditi's file
# alldata <- readRDS("/Users/tuantran/Library/CloudStorage/GoogleDrive-tuantran@iu.edu/Shared drives/[Sec] IN-MED-INF-TRANLAB/Tran Lab Shared/Projects/Doris Duke PfSPZ Kenya/Tuan PfSPZ/KenyaPfSPZ/Final Data Files/all_non_transcriptomic_data.rds")

# eset <- readRDS("/Users/tuantran/Library/CloudStorage/GoogleDrive-tuantran@iu.edu/Shared drives/[Sec] IN-MED-INF-TRANLAB/Tran Lab Shared/Projects/Doris Duke PfSPZ Kenya/Tuan PfSPZ/KenyaPfSPZ/Final Data Files/KSPZV1 logCPM expression sets for visualization/PfSPZ_cpm_ExpressionSet_243x21576_AllGroups_bothBatches_25_rmBatchFX_06082021_TMT_logTRUE.rds")
```

# Convert gene expression data to BTM expression scores and join to other data

```{r collapse eset to modules,  message=FALSE, warning=FALSE, results='hide'}
source("https://raw.githubusercontent.com/TranLab/ModuleLists/main/Gene2ModuleExpressionScores.R")

pas_hibtm <- Gene2ModuleExpressionScores(eset, module_list = "highBTMs", summary_stat = median)
rownames(pas_hibtm) <- paste(rownames(pas_hibtm), "highBTM")
rownames(pas_hibtm) <- gsub(" ", "_", rownames(pas_hibtm))

pas_monaco <- Gene2ModuleExpressionScores(eset, module_list = "MonacoModules", summary_stat = median)
rownames(pas_monaco) <- paste(rownames(pas_monaco), "monaco")
rownames(pas_monaco) <- gsub(" ", "_", rownames(pas_monaco))

pas_lowbtm <- Gene2ModuleExpressionScores(eset, module_list = "lowBTMs", summary_stat = median)
rownames(pas_lowbtm) <- paste(rownames(pas_lowbtm), "lowBTM")
rownames(pas_lowbtm) <- gsub(" ", "_", rownames(pas_lowbtm))
```

```{r collapse eset to modules continued,  message=FALSE, warning=FALSE, results='hide'}
#test innate only
data(tmod)
innate <-  tmod[unique(c(grep("[Ii]nterferon", tmod$gs$Title),
                         grep("interferon", tmod$gs$Title),
                         grep("[Ii]nnate", tmod$gs$Title),
                         grep("LPS", tmod$gs$Title),
                         grep("[Ee]ndotoxin", tmod$gs$Title),
                         grep("[Aa]ctivated", tmod$gs$Title),
                         grep("[Mm]onocytes", tmod$gs$Title),
                         grep("DC", tmod$gs$Title),
                         grep("[Dd]endritic", tmod$gs$Title),
                         grep("[Nn]eutrophils", tmod$gs$Title),
                         grep("[Mm]yeloid", tmod$gs$Title),
                         grep("[Vv]iral", tmod$gs$Title),
                         grep("[Pp]rostanoids", tmod$gs$Title),
                         grep("[Pp]rostaglandins", tmod$gs$Title),
                         grep("IRF", tmod$gs$Title),
                         grep("IFN", tmod$gs$Title),
                         grep("IL", tmod$gs$Title),
                         grep("CXC", tmod$gs$Title),
                         grep("TLR", tmod$gs$Title),
                         grep("JAK", tmod$gs$Title),
                         grep("STAT", tmod$gs$Title),
                         grep("RIG", tmod$gs$Title),
                         grep("TNF", tmod$gs$Title),
                         grep("[Ii]nflam", tmod$gs$Title),
                         grep("I\\:C", tmod$gs$Title)))]
eigengene_LI_DC <- eigengene(x = exprs(eset), g = fData(eset)$GeneSymbol, mset = tmod, k = 1)

pas_all <- t(rbind(pas_hibtm, pas_monaco, pas_lowbtm, eigengene_LI_DC)) %>%
  as.data.frame() %>%
  rownames_to_column(var = "PATID") %>%
  mutate(PATID = gsub("_25", "", PATID))

clin_dat <- pData(eset) %>%
  dplyr::select(PATID, treat, tte.mal.atp.6, pfcsp_post,log2FC_CSPAb)

alldata_select <- alldata %>%
  dplyr::select(PATID.OG, Timepoint, contains("FACS"), contains("CytokineObsConc")) %>%
  dplyr::rename(PATID = PATID.OG) %>%
  mutate(PATID = gsub("\\_.*", "", PATID)) %>%
  pivot_wider(., id_cols = PATID, names_from = Timepoint, values_from = 3:ncol(.),names_sep = "_") %>%
  select_if(~sum(!is.na(.)) > 0) %>% #remove columns with all NAs
  dplyr::select(PATID, contains("FACS_PfSPZ"), matches("(FACS.*25)"), matches("(FACS_PfSPZ-specific_CD3.*25)"),
                matches("(FACS_CSP-specific.*25)"), contains("CytokineObsConc")) %>%
  dplyr::select(-c(contains("Vd2_of_TCRgd"))) %>% #remove redundant features
  dplyr::select(-c(contains("of_live_lymphocytes"))) %>% #remove redundant features
  dplyr::select(-c(contains("of_live_monocytes"))) #remove redundant features

all_data_btm <- clin_dat %>%
                left_join(., alldata_select,  by = "PATID") %>%
                left_join(., pas_all,  by = "PATID") %>%
  dplyr::select(-treat)
```

The following is applicable to baseline data only: columns that have a large percentage of zero values and a few high leverage points tend to show up as significant correlations downstream. To avoid this, remove columns that contain ~>75% zero values. Only certain cytokines fell into this category, and by thresholding, we were able to remove them and prevent 'false' correlations. 

```{r remove columns with >75% zeros,  message=FALSE, warning=FALSE}
# For high-level BTMs
ncol(all_data_btm)
i <- colSums(all_data_btm == 0, na.rm=TRUE) < 0.75*nrow(all_data_btm)  
all_data_btm <- all_data_btm[i]
ncol(all_data_btm) 
```

# Run correlations for all BTMs

```{r correlations of the combined all_data,  message=FALSE, warning=FALSE}
#remove Monaco
all_data_allBTMs <- all_data_btm %>%
  dplyr::select(-c(PATID))
cor_res_allBTMs <- rcorr(as.matrix(all_data_allBTMs))

#Flatten correlation matrix using nifty function
#http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

flat_cormat_allBTMs <- flattenCorrMatrix(cor_res_allBTMs$r, cor_res_allBTMs$P)

#filter only correlations that include CSP IgG, post-vax, and time-to-parasitemia
flat_cormat_allBTMs_filtered <- flat_cormat_allBTMs %>%
  filter(grepl("tte.mal.atp.6", .$row) | grepl("tte.mal.atp.6", .$col) |
           grepl("pfcsp_post", .$row) | grepl("pfcsp_post", .$col) |
           grepl("log2FC_CSPAb", .$row) | grepl("log2FC_CSPAb", .$col) |
           (grepl("_25", .$row) & grepl("PfSPZ-specific_CD3", .$row)) | (grepl("_25", .$col) & grepl("PfSPZ-specific_CD3", .$col))) %>%
  filter(!grepl("_of_live_lymphocytes", .$row)) %>% #remove redundant features (of live lymphocytes = of live PBMCs)
  filter(!grepl("_of_live_lymphocytes", .$col)) %>%
  mutate(row_dummy = gsub("_25|_0", "", .$row)) %>%
  mutate(col_dummy = gsub("_25|_0", "", .$col)) %>%
  filter(row_dummy != col_dummy) %>% #remove features that correlate with itself overtime
  mutate(FDR = p.adjust(p, method = "BH")) %>%
  filter(FDR<0.05)
  arrange(FDR) 

select_features <- unique(c(flat_cormat_allBTMs_filtered$row, flat_cormat_allBTMs_filtered$column))
```


```{r plot networks allBTMs}
library(corrr)

cormat_allBTMs_filtered <- cor_res_allBTMs$r[select_features,select_features]

rownames(cormat_allBTMs_filtered) <- gsub("FACS_|_lowBTM|_highBTM|_monaco|CytokineObsConc_", "", rownames(cormat_allBTMs_filtered))
colnames(cormat_allBTMs_filtered) <- gsub("FACS_|_lowBTM|_highBTM|_monaco|CytokineObsConc_", "", colnames(cormat_allBTMs_filtered))
rownames(cormat_allBTMs_filtered) <- gsub("PfSPZ", "Pf", rownames(cormat_allBTMs_filtered))
colnames(cormat_allBTMs_filtered) <- gsub("PfSPZ", "Pf", colnames(cormat_allBTMs_filtered))
rownames(cormat_allBTMs_filtered) <- gsub("_of_live_PBMCs", "", rownames(cormat_allBTMs_filtered))
colnames(cormat_allBTMs_filtered) <- gsub("_of_live_PBMCs", "", colnames(cormat_allBTMs_filtered))
rownames(cormat_allBTMs_filtered) <- gsub("_of_live_leukocytes", "", rownames(cormat_allBTMs_filtered))
colnames(cormat_allBTMs_filtered) <- gsub("_of_live_leukocytes", "", colnames(cormat_allBTMs_filtered))
rownames(cormat_allBTMs_filtered) <- gsub("25", "post-vax", rownames(cormat_allBTMs_filtered))
colnames(cormat_allBTMs_filtered) <- gsub("25", "post-vax", colnames(cormat_allBTMs_filtered))
rownames(cormat_allBTMs_filtered) <- gsub("0", "baseline", rownames(cormat_allBTMs_filtered))
colnames(cormat_allBTMs_filtered) <- gsub("0", "baseline", colnames(cormat_allBTMs_filtered))
rownames(cormat_allBTMs_filtered) <- gsub("_", " ", rownames(cormat_allBTMs_filtered))
colnames(cormat_allBTMs_filtered) <- gsub("_", " ", colnames(cormat_allBTMs_filtered))


my_network_plot <- network_plot(cormat_allBTMs_filtered, 
                                legend = "full",
                                colors = c("blue", "white", "red"),
                                min_cor = .10) 
my_network_plot
#ggsave(filename = paste0(resdir, "network_plot_mincor20_spearman.pdf"), plot = my_network_plot, device = "pdf", width = 8, height = 3)
```

# Run correlations for high BTMs

```{r correlations of the combined all_data hiBTMs,  message=FALSE, warning=FALSE}
#remove Monaco
all_data_highBTMs <- all_data_btm %>%
  dplyr::select(-c(PATID, contains("_monaco"), contains("_lowBTM")))
cor_res_highBTMs <- rcorr(as.matrix(all_data_highBTMs))

#Flatten correlation matrix using nifty function
#http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

flat_cormat_highBTMs <- flattenCorrMatrix(cor_res_highBTMs$r, cor_res_highBTMs$P)

#filter only correlations that include CSP IgG, cytokines, and time-to-parasitemia

flat_cormat_highBTMs_filtered <- flat_cormat_highBTMs %>%
  filter(grepl("tte", .$row) | grepl("tte", .$col) |
           grepl("pfcsp_post", .$row) | grepl("pfcsp_post", .$col) |
           grepl("log2FC_CSPAb", .$row) | grepl("log2FC_CSPAb", .$col) |
           (grepl("_25", .$row) & grepl("PfSPZ-specific_CD3", .$row)) | (grepl("_25", .$col) & grepl("PfSPZ-specific_CD3", .$col)) |
           (grepl("_25", .$row) & grepl("CSP-spec", .$row)) | (grepl("_25", .$col) & grepl("CSP-spec", .$col))) %>%
  filter(grepl("tte", .$row) | grepl("tte", .$col)) %>%
  mutate(FDR = p.adjust(p, method = "BH")) %>%
  filter(FDR<0.20) %>%
  arrange(FDR)

select_features <- unique(c(flat_cormat_highBTMs_filtered$row, flat_cormat_highBTMs_filtered$column))
```

```{r plot networks highBTMs}
library(corrr)

cormat_highBTMs_filtered <- cor_res_highBTMs$r[select_features,select_features]

rownames(cormat_highBTMs_filtered) <- gsub("FACS_|_lowBTM|_highBTM|_monaco|CytokineObsConc_", "", rownames(cormat_highBTMs_filtered))
colnames(cormat_highBTMs_filtered) <- gsub("FACS_|_lowBTM|_highBTM|_monaco|CytokineObsConc_", "", colnames(cormat_highBTMs_filtered))
rownames(cormat_highBTMs_filtered) <- gsub("PfSPZ", "Pf", rownames(cormat_highBTMs_filtered))
colnames(cormat_highBTMs_filtered) <- gsub("PfSPZ", "Pf", colnames(cormat_highBTMs_filtered))
rownames(cormat_highBTMs_filtered) <- gsub("_of_live_PBMCs", "", rownames(cormat_highBTMs_filtered))
colnames(cormat_highBTMs_filtered) <- gsub("_of_live_PBMCs", "", colnames(cormat_highBTMs_filtered))
rownames(cormat_highBTMs_filtered) <- gsub("_of_live_leukocytes", "", rownames(cormat_highBTMs_filtered))
colnames(cormat_highBTMs_filtered) <- gsub("_of_live_leukocytes", "", colnames(cormat_highBTMs_filtered))
rownames(cormat_highBTMs_filtered) <- gsub("25", "post-vax", rownames(cormat_highBTMs_filtered))
colnames(cormat_highBTMs_filtered) <- gsub("25", "post-vax", colnames(cormat_highBTMs_filtered))
rownames(cormat_highBTMs_filtered) <- gsub("0", "baseline", rownames(cormat_highBTMs_filtered))
colnames(cormat_highBTMs_filtered) <- gsub("0", "baseline", colnames(cormat_highBTMs_filtered))
rownames(cormat_highBTMs_filtered) <- gsub("_", " ", rownames(cormat_highBTMs_filtered))
colnames(cormat_highBTMs_filtered) <- gsub("_", " ", colnames(cormat_highBTMs_filtered))


my_highBTM_network_plot <- network_plot(cormat_highBTMs_filtered, 
                                legend = "full",
                                colors = c("blue", "white", "red"),
                                min_cor = .20)
```

```{r plot highBTM network plot, fig.align='center', fig.height=10, fig.width=10}
my_highBTM_network_plot
#ggsave(filename = paste0(resdir, "network_plot_mincor20_spearman.pdf"), plot = my_network_plot, device = "pdf", width = 8, height = 3)
```


# Run correlations for Monaco modules

```{r correlations of the combined all_data monaco,  message=FALSE, warning=FALSE}
#remove highBTM
all_data_monaco <- all_data_btm %>%
  dplyr::select(-c(PATID, contains("highBTM"), contains("lowBTM")))
cor_res_monaco <- rcorr(as.matrix(all_data_monaco), type = "pearson")

#Flatten correlation matrix using nifty function
#http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

flat_cormat_monaco <- flattenCorrMatrix(cor_res_monaco$r, cor_res_monaco$P)

#filter only correlations that include CSP IgG, cytokines, and time-to-parasitemia
flat_cormat_monaco_filtered <- flat_cormat_monaco %>%
  filter(grepl("tte", .$row) | grepl("tte", .$col) |
           grepl("pfcsp_post", .$row) | grepl("pfcsp_post", .$col) |
           grepl("log2FC_CSPAb", .$row) | grepl("log2FC_CSPAb", .$col) |
           (grepl("_25", .$row) & grepl("PfSPZ-specific_CD3", .$row)) | (grepl("_25", .$col) & grepl("PfSPZ-specific_CD3", .$col)) |
           (grepl("_25", .$row) & grepl("CSP-spec", .$row)) | (grepl("_25", .$col) & grepl("CSP-spec", .$col))) %>%
    mutate(FDR = p.adjust(p, method = "BH")) %>%
  filter(FDR<0.2) %>%
  arrange(FDR)

select_features <- unique(c(flat_cormat_monaco_filtered$row, flat_cormat_monaco_filtered$column))
```

```{r plot networks monaco}
library(corrr)

cormat_monaco_filtered <- cor_res_monaco$r[select_features,select_features]

rownames(cormat_monaco_filtered) <- gsub("FACS_|_lowBTM|_highBTM|_monaco|CytokineObsConc_", "", rownames(cormat_monaco_filtered))
colnames(cormat_monaco_filtered) <- gsub("FACS_|_lowBTM|_highBTM|_monaco|CytokineObsConc_", "", colnames(cormat_monaco_filtered))
rownames(cormat_monaco_filtered) <- gsub("PfSPZ", "Pf", rownames(cormat_monaco_filtered))
colnames(cormat_monaco_filtered) <- gsub("PfSPZ", "Pf", colnames(cormat_monaco_filtered))
rownames(cormat_monaco_filtered) <- gsub("_of_live_PBMCs", "", rownames(cormat_monaco_filtered))
colnames(cormat_monaco_filtered) <- gsub("_of_live_PBMCs", "", colnames(cormat_monaco_filtered))
rownames(cormat_monaco_filtered) <- gsub("_of_live_leukocytes", "", rownames(cormat_monaco_filtered))
colnames(cormat_monaco_filtered) <- gsub("_of_live_leukocytes", "", colnames(cormat_monaco_filtered))
rownames(cormat_monaco_filtered) <- gsub("25", "post-vax", rownames(cormat_monaco_filtered))
colnames(cormat_monaco_filtered) <- gsub("25", "post-vax", colnames(cormat_monaco_filtered))
rownames(cormat_monaco_filtered) <- gsub("0", "baseline", rownames(cormat_monaco_filtered))
colnames(cormat_monaco_filtered) <- gsub("0", "baseline", colnames(cormat_monaco_filtered))
rownames(cormat_monaco_filtered) <- gsub("_", " ", rownames(cormat_monaco_filtered))
colnames(cormat_monaco_filtered) <- gsub("_", " ", colnames(cormat_monaco_filtered))


my_monaco_network_plot <- network_plot(cormat_monaco_filtered, 
                                legend = "full",
                                colors = c("blue", "white", "red"),
                                min_cor = .20) 
```

```{r plot monaco network plot, fig.align='center', fig.height=10, fig.width=10}
my_monaco_network_plot
#ggsave(filename = paste0(resdir, "network_plot_mincor20_spearman.pdf"), plot = my_network_plot, device = "pdf", width = 8, height = 3)
```


# Run correlations for low BTMs

```{r correlations of the combined all_data hiBTMs,  message=FALSE, warning=FALSE}
#remove high BTM and monaco
all_data_lowBTMs <- all_data_btm %>%
  dplyr::select(-c(PATID, contains("_highBTM"), contains("_monaco")))

#Flatten correlation matrix using nifty function
#http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

#find correlations Hmisc method
cor_res_lowBTMs <- rcorr(as.matrix(all_data_lowBTMs), type = "spearman")
flat_cormat_lowBTMs <- flattenCorrMatrix(cor_res_lowBTMs$r, cor_res_lowBTMs$P)

#find correlations psych method
# cor_res_lowBTMs <- all_data_lowBTMs %>%
#   psych::corr.test(., use = "pairwise.complete.obs", method = "spearman", adjust = "fdr")
# flat_cormat_lowBTMs <- flattenCorrMatrix(cor_res_lowBTMs$r, cor_res_lowBTMs$p)
```

Filter only correlations that include CSP IgG, Pf-specific post-vax, and time-to-parasitemia

```{r limit to relevant correlations lowBTMs}
flat_cormat_lowBTMs_filtered <- flat_cormat_lowBTMs %>%
  filter(grepl("tte", .$row) | grepl("tte", .$col) |
           grepl("pfcsp_post", .$row) | grepl("pfcsp_post", .$col) |
           grepl("log2FC_CSPAb", .$row) | grepl("log2FC_CSPAb", .$col) |
           (grepl("_25", .$row) & grepl("PfSPZ-specific_CD3", .$row)) | (grepl("_25", .$col) &
                                                                           grepl("PfSPZ-specific_CD3", .$col)) |
           (grepl("_25", .$row) & grepl("CSP-specific", .$row)) | (grepl("_25", .$col) & grepl("CSP-specific", .$col))) %>%
  mutate(FDR = p.adjust(p, method = "BH")) %>%
  filter(FDR<0.2) %>%
  arrange(FDR)

select_features <- unique(c(flat_cormat_lowBTMs_filtered$row, flat_cormat_lowBTMs_filtered$column))
```

```{r plot networks lowBTMs}
library(corrr)

cor_r_lowBTMs_filtered <- cor_res_lowBTMs$r[select_features,select_features]
cor_n_lowBTMs_filtered <- cor_res_lowBTMs$n[select_features,select_features]
cor_p_lowBTMs_filtered <- cor_res_lowBTMs$p[select_features,select_features]
cor_fdr_res_lowBTMs_filtered <- psych::corr.p(cor_r_lowBTMs_filtered,
                                           cor_n_lowBTMs_filtered,
                                           adjust="fdr",alpha=.2,minlength=5,ci=TRUE)

rownames(cor_r_lowBTMs_filtered) <- gsub("FACS_|_lowBTM|_highBTM|_monaco|CytokineObsConc_", "", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("FACS_|_lowBTM|_highBTM|_monaco|CytokineObsConc_", "", colnames(cor_r_lowBTMs_filtered))
rownames(cor_r_lowBTMs_filtered) <- gsub("PfSPZ", "Pf", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("PfSPZ", "Pf", colnames(cor_r_lowBTMs_filtered))
rownames(cor_r_lowBTMs_filtered) <- gsub("_of_live_PBMCs", "", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("_of_live_PBMCs", "", colnames(cor_r_lowBTMs_filtered))
rownames(cor_r_lowBTMs_filtered) <- gsub("_of_live_leukocytes", "", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("_of_live_leukocytes", "", colnames(cor_r_lowBTMs_filtered))
rownames(cor_r_lowBTMs_filtered) <- gsub("25", "post-vax", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("25", "post-vax", colnames(cor_r_lowBTMs_filtered))
rownames(cor_r_lowBTMs_filtered) <- gsub("_0", " baseline", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("_0", " baseline", colnames(cor_r_lowBTMs_filtered))
rownames(cor_r_lowBTMs_filtered) <- gsub("_", " ", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("_", " ", colnames(cor_r_lowBTMs_filtered))
rownames(cor_r_lowBTMs_filtered) <- gsub("tte.mal.atp.6", "time to parasitemia", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("tte.mal.atp.6", "time to parasitemia", colnames(cor_r_lowBTMs_filtered))
rownames(cor_r_lowBTMs_filtered) <- gsub("log2FC CSPAb", "ΔCSP-specific IgG", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("log2FC CSPAb", "ΔCSP-specific IgG", rownames(cor_r_lowBTMs_filtered), colnames(cor_r_lowBTMs_filtered))
rownames(cor_r_lowBTMs_filtered) <- gsub("pfcsp post", "CSP-specific IgG post-vax", rownames(cor_r_lowBTMs_filtered))
colnames(cor_r_lowBTMs_filtered) <- gsub("pfcsp post", "CSP-specific IgG post-vax", colnames(cor_r_lowBTMs_filtered))

my_lowBTM_corplot <- ggcorrplot(corr = cor_r_lowBTMs_filtered,
                               hc.order = TRUE,
                               hc.method = "complete",
                               outline.col = "white",
                               ggtheme = ggplot2::theme_minimal(),
                               colors = c("darkblue", "white", "darkred"),
                               type = "upper",
                               sig.level = 0.05,
                               p.mat = cor_p_lowBTMs_filtered,
                               insig = "blank",
                               lab = "true",
                               tl.cex = 10,
                               lab_size = 2)

my_lowBTM_network_plot <- network_plot(cor_r_lowBTMs_filtered, 
                                legend = "full",
                                colors = c("darkblue", "white", "darkred"),
                                min_cor = .20,
                                curved = TRUE)

```

```{r plot network plot lowBTM, fig.align='center', fig.height=4, fig.width=8}
my_lowBTM_network_plot
#ggsave(filename = paste0(resdir, "all_timepoints_corplot_cognitive_cytokines_spearman_fdr10_with_maternal_education.pdf"), plot = baseline_corplot, device = "pdf", width = 8, height = 8)
```

```{r plot corrmat lowBTM, fig.align='center', fig.height=6, fig.width=6}
my_lowBTM_corplot
#ggsave(filename = paste0(resdir, "all_timepoints_corplot_cognitive_cytokines_spearman_fdr10_with_maternal_education.pdf"), plot = baseline_corplot, device = "pdf", width = 8, height = 8)
```


# Run correlations for eigenvalue BTMs

```{r correlations of the combined innate BTMs,  message=FALSE, warning=FALSE}
#remove high BTM and monaco
all_data_innate_BTMs <- all_data_btm %>%
  dplyr::select(-c(PATID, contains("_highBTM"), contains("_monaco"), contains("_lowBTM"), contains("DC.")))

#Flatten correlation matrix using nifty function
#http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

#find correlations Hmisc method
cor_res_innate_BTMs <- rcorr(as.matrix(all_data_innate_BTMs), type = "pearson")
flat_cormat_innate_BTMs <- flattenCorrMatrix(cor_res_innate_BTMs$r, cor_res_innate_BTMs$P)

#find correlations psych method
# cor_res_innate_BTMs <- all_data_innate_BTMs %>%
#   psych::corr.test(., use = "pairwise.complete.obs", method = "spearman", adjust = "fdr")
# flat_cormat_innate_BTMs <- flattenCorrMatrix(cor_res_innate_BTMs$r, cor_res_innate_BTMs$p)

#find correlations using group dplyr and cor
# all_data_innate_BTMs_long <- all_data_innate_BTMs %>%
#   pivot_longer(., cols = )
# 
# summarized_correlations <- FACS_TTE_longdat %>%
#   group_by(feature) %>%
#   summarise(baseline_spearman_r_tte6mos = stats::cor.test(Baseline, tte.mal.atp.6, method = myCorMeth)$estimate,
#             baseline_spearman_pval_tte6mos = stats::cor.test(Baseline, tte.mal.atp.6, method = myCorMeth)$p.val,
#             postvax_spearman_r_tte6mos = stats::cor.test(Post_Vaccination, tte.mal.atp.6, method = myCorMeth)$estimate,
#             postvax_spearman_pval_tte6mos = stats::cor.test(Post_Vaccination, tte.mal.atp.6, method = myCorMeth)$p.val,
#             LFC_spearman_r_tte6mos = stats::cor.test(LFC, tte.mal.atp.6, method = myCorMeth)$estimate,
#             LFC_spearman_pval_tte6mos = stats::cor.test(LFC, tte.mal.atp.6, method = myCorMeth)$p.val) %>%

```

Filter only correlations that include CSP IgG, Pf-specific post-vax, and time-to-parasitemia

```{r limit to relevant correlations innate_BTMs}
# flat_cormat_innate_BTMs_filtered <- flat_cormat_innate_BTMs %>%
#   filter(grepl("tte", .$row) | grepl("tte", .$col) |
#            grepl("pfcsp_post", .$row) | grepl("pfcsp_post", .$col) |
#            grepl("log2FC_CSPAb", .$row) | grepl("log2FC_CSPAb", .$col) |
#            (grepl("_25", .$row) & grepl("PfSPZ-specific_CD3", .$row)) | (grepl("_25", .$col) &
#                                                                            grepl("PfSPZ-specific_CD3", .$col)) |
#            (grepl("_25", .$row) & grepl("CSP-specific", .$row)) | (grepl("_25", .$col) & grepl("CSP-specific", .$col))) %>%
#   mutate(FDR = p.adjust(p, method = "BH")) %>%
#   filter(FDR<0.05) %>%
#   arrange(FDR)

flat_cormat_innate_BTMs_filtered <- flat_cormat_innate_BTMs %>%
  filter(grepl("tte", .$row) | grepl("tte", .$col)) %>%
  mutate(FDR = p.adjust(p, method = "BH")) %>%
  filter(p<0.01) %>%
  arrange(FDR)

select_features <- unique(c(flat_cormat_innate_BTMs_filtered$row, flat_cormat_innate_BTMs_filtered$column))
```

```{r plot networks innate_BTMs}
library(corrr)

cor_r_innate_BTMs_filtered <- cor_res_innate_BTMs$r[select_features,select_features]
cor_n_innate_BTMs_filtered <- cor_res_innate_BTMs$n[select_features,select_features]
cor_p_innate_BTMs_filtered <- cor_res_innate_BTMs$p[select_features,select_features]
cor_fdr_res_innate_BTMs_filtered <- psych::corr.p(cor_r_innate_BTMs_filtered,
                                           cor_n_innate_BTMs_filtered,
                                           adjust="fdr",alpha=.2,minlength=5,ci=TRUE)

rownames(cor_r_innate_BTMs_filtered) <- gsub("FACS_|_innate_BTM|_highBTM|_monaco|CytokineObsConc_", "", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("FACS_|_innate_BTM|_highBTM|_monaco|CytokineObsConc_", "", colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- gsub("PfSPZ", "Pf", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("PfSPZ", "Pf", colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- gsub("_of_live_PBMCs", "", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("_of_live_PBMCs", "", colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- gsub("_of_live_leukocytes", "", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("_of_live_leukocytes", "", colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- gsub("25", "post-vax", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("25", "post-vax", colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- gsub("_0", " baseline", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("_0", " baseline", colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- gsub("_", " ", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("_", " ", colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- gsub("tte.mal.atp.6", "time to parasitemia", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("tte.mal.atp.6", "time to parasitemia", colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- gsub("log2FC CSPAb", "ΔCSP-specific IgG", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("log2FC CSPAb", "ΔCSP-specific IgG", rownames(cor_r_innate_BTMs_filtered), colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- gsub("pfcsp post", "CSP-specific IgG post-vax", rownames(cor_r_innate_BTMs_filtered))
colnames(cor_r_innate_BTMs_filtered) <- gsub("pfcsp post", "CSP-specific IgG post-vax", colnames(cor_r_innate_BTMs_filtered))
rownames(cor_r_innate_BTMs_filtered) <- plyr::mapvalues(rownames(cor_r_innate_BTMs_filtered),
                                                        from = tmod$gs$ID, to = tmod$gs$Title, warn_missing =F) #replace BTM IDs with titles
colnames(cor_r_innate_BTMs_filtered) <- plyr::mapvalues(colnames(cor_r_innate_BTMs_filtered),
                                                        from = tmod$gs$ID, to = tmod$gs$Title, warn_missing =F) #replace BTM IDs with titles

my_innate_BTM_corplot <- ggcorrplot(corr = cor_r_innate_BTMs_filtered,
                               hc.order = TRUE,
                               hc.method = "complete",
                               outline.col = "white",
                               ggtheme = ggplot2::theme_minimal(),
                               colors = c("blue", "white", "red"),
                               type = "upper",
                               sig.level = 0.05,
                               p.mat = cor_p_innate_BTMs_filtered,
                               insig = "blank",
                               lab = "true",
                               tl.cex = 10,
                               lab_size = 2)

my_innate_BTM_network_plot <- network_plot(cor_r_innate_BTMs_filtered, 
                                legend = "full",
                                colors = c("blue", "white", "red"),
                                min_cor = .2,
                                curved = TRUE)

```

```{r plot corrmat innate_BTM, fig.align='center', fig.height=6, fig.width=6}
my_innate_BTM_corplot

#ggsave(filename = paste0(resdir, "all_timepoints_corplot_cognitive_cytokines_spearman_fdr10_with_maternal_education.pdf"), plot = baseline_corplot, device = "pdf", width = 8, height = 8)
```

```{r plot innate_BTM network plot, fig.align='center', fig.height=4, fig.width=8}
my_innate_BTM_network_plot
#ggsave(filename = paste0(resdir, "network_plot_mincor20_spearman.pdf"), plot = my_network_plot, device = "pdf", width = 8, height = 3)
```

